### 2019-11-28 latex/sections/abstract/english.tex
latex add

this is my abstract written in the language of English I am testing thought

alas  I knew

ok so now I'm trying something new

asdf

test

### 2019-11-28 latex/sections/abstract/icelandic.tex
latex add

this is my abstract written in the language of íslensku þæööasdf

### 2019-11-28 latex/sections/discussion.tex
latex add

\chapter{Discussion}

\ifdraft{Here I will discuss the myriad of fields research like this can help
for instance for health organizations, patients, taxpayers, and pharma}
hey

\section{Summary}

\ifdraft{summarize the work}
hey

\section{Conclusion}

\ifdraft{conclude the work, discuss the significance of the work}
hey

### 2019-11-28 latex/sections/glossary.tex
latex add

%% $URL$
%% $Id$
%% Glossary term definitions
%% Note, you will need to reference them in the text after defining them

\newglossaryentry{glossary}{name=glossary, description={a place to put all the terms that confuse people}}
\newglossaryentry{ohm}{name=ohm,symbol={\ensuremath{\Omega}}, description=unit of electrical resistance}  % Do not break this with newlines!
\newglossaryentry{dissertation}{name=dissertation, description={A large pile of paper that professors use to prop up shelves.}}
\newacronym{rfid}{RFID}{Radio Frequency IDentification}

\newacronym{led}{LED}{light-emitting diodes}

\newacronym{auv}{AUV}{Autonomous Undersea Vehicle}
%% Reference manual: http://ctan.uib.no/macros/latex/contrib/glossaries/glossaries-user.pdf

%% You can also define most entries near where they are used, but this may cause problems.  See manual chapter 4.8
%% Be careful about putting entries on Chapter, Section, etc.  See manual Chapter 6.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "DEGREE-NAME-YEAR"
%%% End:

### 2019-11-28 latex/sections/introduction.tex
latex add

\chapter{Introduction\label{cha:introduction}}
%% \ifdraft only shows the text in the first argument if you are in draft mode.
%% These directions will disappear in other modes.
\ifdraft{State the objectives of the exercise. Ask yourself:
  \underline{Why} did I design/create the item? What did I aim to
  achieve? What is the problem I am trying to solve?  How is my
  solution interesting or novel?}{}

In the year $2017$ the approximate amount of adults with diabetes was 425-million people \cite{Intro_FactsandFigures}, this number is 
projected to rise to $629$-million people by the year $2045$. The rate of people that are diagnosed with diabetes 
is 1 in every 2. Secondly of all people with diabetes 1 in every 3 develop diabetic retinopaty (DR)\cite{Intro_DiabetesRate}.

\textit{Diabetic retinopathy} is a symptom of both type 1 and type 2 diabetes and is now the leading cause of blindness amongst adults today. DR is 
diagnosed by two major technologies, firstly there is the fundus imaging method, where a photograph is taken of the fundus (back) 
of the patients eye, an example of this can be seen in Figure \ref{fig:fundus-image1}. Secondly ophthalmologists examine 
Optical Choerence Tomography (OCT) scans of human retina, an example of this can be seen in Figure \ref{fig:oct-retina1}. 
The work in this paper does not focus on the latter. The diagnosis of DR is costly and as the statistic indicate 
the amount of work is going to increase.

\begin{figure}[h]
  \centering
  \includegraphics[width=.3\linewidth]{../graphics/fundus_image1.jpg}
  \caption{Example of Fundus image}
  \label{fig:fundus-image1}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=.5\linewidth]{../graphics/oct_fundus.png}
  \caption{Example of OCT on Retina}
  \label{fig:oct-retina1}
\end{figure}

Patients that have DR are grouped into four different stages\cite{Intro_FourStages}:

1. \textit{Mild Nonproliferative Retinopathy} - This is the earliest stage of diabetic retinopathy, and it’s characterized by balloon-like swelling in the retina’s blood vessels. These are called microaneurysms, and these vessels can leak into the eye.

2. \textit{Moderate Nonproliferative Retinopathy} - During this stage, the blood vessels nourishing the retina swell and may even become blocked. This can contribute to diabetic macular edema (DME) which is a build-up of fluid in the macula region of the retina.

3. \textit{Severe Nonproliferative Retinopathy} -  At this stage, an increasing number of blood vessels nourishing the eye have become blocked. As a result, the retina is signaled to grow new blood vessels.

4. \textit{Proliferative Diabetic Retinopathy} - This is the final stage of diabetic retinopathy. New blood vessels proliferate, growing inside the retina and into the vitreous gel, which is the fluid that fills the eye. Because these blood vessels are delicate, they may begin to leak and bleed. As a result, scar tissue may form, causing retinal detachment, the pulling away of the retina from underlying tissue. Retinal detachment may cause spotty vision, flashes of lights, or severe vision loss.

Generally without intervention patients advance from one stage to the next eventually leading 
to permanent vision impairment, however with intervention by an ophthalmologist this advancement 
can be prevented \cite{Intro_StageAdvancement}.

This paper focuses on applying machine learning methods to patient information and Fundus images of those 
patients to assess the likelyhood (risk) of those patients DR developing into SDR.

In this paper we will apply machine learning methods to patients information and fundus images of patients 
that has been collected a period of time to examine a patiens likelyhood to advance from one stage 
of DR to the next, attempting to predict the timeline the patient has to seek medical help.

\section{Machine learning methods in Ophthalmology}

Applying machine learning to fundus images and patient information has been done before very successfully 
researchers\cite{Intro_googleDeepmind}, this work generally focuses on applying image recognition on 
the images to identify which stage of DR the eye has reached.

\section{Background}
\ifdraft{Provide background about the subject matter (e.g. How was morse code
developed?  How is it used today?). 
This is a place where there are usually many citations.
It is suspicious when there is not. 
Include the purpose of the different equipment and your design intent. 
Include references to relevant scientific/technical work and books.}

Applying statistics to medical data in order to better understand a patients evolution from DR to SDR 
has been done previously by a company called RetinaRisk, a significant contributor to this paper, but 
RetinaRisk's solution does not apply machine learning but a specialists field knowledge and a handcrafted 
formula. RetinaRisks formula has been used to significantly reduce costs to health organizations as 
when patients following their recommendations for doctor checkups come in less often but at time that 
are more critical. im just writing something so that it looks like I'm writing something

### 2019-11-28 latex/sections/managerial.tex
latex add

%% Some elements have different names on the RU Masters rules
%% They will be annotated with RUM: "name"
\frontmatter{} % setup formatting at beginning

%\frontcover{}%%If you want to see what it looks like with the printed cover
%% TODO:  link to fill-in PDF file on RU website

\frontrequiredpages{}%% the various signaturepages and abstract
%%% WARNING:  if you get an error on the previous line, it is probably because
%%% you put a bad macro or something strange in a title, author, or abstract.

\ifdraft{\coverchapter{Important!!!  Read the Instructions!!!} If you
  have not already done so, \LaTeX{} the \path{instructions.tex} to
  learn how to setup your document and use some of the features.  You
  can see a (somewhat recent) rendered PDF of the instructions included in this folder at \path{instructions-publish.pdf}.
  There is also more information on working with \LaTeX{} at
  \url{http://samvinna.ru.is/project/htgaru/how-to-get-around-projects-publish.pdf}.
  This includes common problems and fixes.

  This page will disappear in anything other than draft mode.}{}

%% Dedication is optional, comment out if it is absent
\begin{dedications}
  I dedicate this to my spouse/child/pet/power animal.
\end{dedications}

\enableindents{}% turn on/off paragraph indents
\coverchapter{Acknowledgements} 
\begin{quotation}
So long, and thanks for all the fish.
\end{quotation}
\vspace{\baselineskip}

\draftnote{Acknowledgements are optional; comment this chapter out if they are absent
  Note that it is important to acknowledge any funding that helped in the work}

This work was funded by \the\year~RANNIS grant ``Survey of man-eating Minke whales'' 1415550.
Additional equipment was generously donated by the Icelandic Tourism Board.

\coverchapter{Preface}

This dissertation is original work by the author, Sigurður Helgason.

\draftnote{The preface is an optional element
  explaining a little who performed what work.  See
  \url{https://www.grad.ubc.ca/sites/default/files/materials/thesis_sample_prefaces.pdf}
  for suggestions.

  List of publications as part of the preface is
  optional unless elements of the work have already been published.
  It should be a comprehensive list of all publications in which
  material in the thesis has appeared, preferably with references to
  sections as appropriate.  This is also a good place to state
  contribution of student and contribution of others to the work
  represented in the thesis.}

%\coverchapter{Publications}
%% RUM: Not mentioned, this was found in the CS thesis template.  
%% Maybe more applicable to PhD dissertations?
%%% Probably a duplication from before Preface became standard.

\starttables{}% setup formatting
%% TOC, list of figures and list of tables are required
\tableofcontents{}\clearpage%%RUM: "Table of contents"
\listoffigures{}\clearpage%%RUM: "List of figures"
\listoftables{}\clearpage%%RUM: "List of tables"

%\coverchapter{List of drawings and enclosed material}
%RUM: "List of drawings and enclosed material, e.g. CD(as appropriate)"

\listoffixmes{}
% if using fixme package, lists what needs to be done

%% The list of abbreviations is an example of a special list
%% Other lists may be added, such as lists of algorithms, symbols, theorems, etc.
%% IN CS PhD, this is sometimes centered.
\coverchapter{List of Abbreviations}%%RUM: Not mentioned
\begin{tabular}{ll}
MSc &Masters of Science\\
DR & Diabetic Retinopathy\\
ML & Machine Learning\\
AI & Artificial Intelligence\\
\end{tabular}

% \coverchapter{List of Symbols}%%RUM: Not mentioned
% \begin{tabular}{lll}
% Symbol &Description &Value/Units\\
% $E$ &Energy &\si{\joule}\\
% $m$ &Mass &\si{gram}\\
% $c$ &Speed of Light &\SI{2.99E8}{\meter\per\second\square}\\
% \end{tabular}

### 2019-11-28 latex/sections/methods.tex
latex add

\chapter{Methods}

%\section{Machine learning methods}

This section discusses the various machine learning methods utilized throughout this project 
and discusses their applicability.

%\subsection{Deep neural network}

I'm a little teapot

%\subsection{Data joined network}

I'm a smaller teapot

### 2019-11-28 latex/sections/results.tex
latex add

\chapter{Results}

\ifdraft{In this section you discuss any issues that came up while developing
the system.  If you found something particularly interesting,
difficult, or an important learning experience, put it here.  This is
also a good place to put additional figures and data.}
hello world

### 2020-01-26 latex/sections/background.tex
updating skeleton

\section{Background}

\subsection{Reinforcement Learning}

In 2017 a Google based artificial intelligence research company published a paper called
`Mastering Chess and Shogi by Self-Play with a General Reinforcement Learning Algorithm`.
The paper described a Model based Reinforment learning agent and an accompanying Residual
Deep Neural Network architecture that was able to master the game of chess without being
given any idea of how to play the game of chess, and others. The architecture used in this
paper is heavily inspired by that architecture.

\subsection{Model Interpretability}

In 2018 the paper 'Interpretability Beyond Feature Attribution: Quantitative Testing with Concept Activation Vectors (TCAV)'
was published where an approach to model interpretability was introduced which is able to
distinguish

The deep neural network architecture that is used in this paper is based on the AlphaZero paper
published by DeepMind,

### 2020-01-26 latex/sections/introduction.tex
updating skeleton

@@
 
-In the year $2017$ the approximate amount of adults with diabetes was 425-million people \cite{Intro_FactsandFigures}, this number is 
-projected to rise to $629$-million people by the year $2045$. The rate of people that are diagnosed with diabetes 
-is 1 in every 2. Secondly of all people with diabetes 1 in every 3 develop diabetic retinopaty (DR)\cite{Intro_DiabetesRate}.
+In the world of deep learning is exciting, we have models that are able to examine
+images and very reliably be able to identify it's content. Deep learning models have
+beaten Opthalmologists in identifing diabetic retinopathy, they've identified
+cancer cells where others have not. Deep learning models have even predicted
+the likelyhood a person will die in the following year. These models have done
+these things extremely accurately, cheap, and fast. These successes allow us to
+dream of a future where all things are eventually possible with the help of
+technology.
 
-\textit{Diabetic retinopathy} is a symptom of both type 1 and type 2 diabetes and is now the leading cause of blindness amongst adults today. DR is 
-diagnosed by two major technologies, firstly there is the fundus imaging method, where a photograph is taken of the fundus (back) 
-of the patients eye, an example of this can be seen in Figure \ref{fig:fundus-image1}. Secondly ophthalmologists examine 
-Optical Choerence Tomography (OCT) scans of human retina, an example of this can be seen in Figure \ref{fig:oct-retina1}. 
-The work in this paper does not focus on the latter. The diagnosis of DR is costly and as the statistic indicate 
-the amount of work is going to increase.
+The field of examining deep learning models to gain a richer understanding
+in how they work, and what makes them so much better than humans at various tasks
+is still new. This is the field of Explainable Artificial Intelligence (XAI).
+If we wish to continue using artificial intelligence (AI) and machine learning (ML),
+to improve our lives we must examine how they work, this is not only to improve
+ourselves but these explanations are a legal requirement now in many areas, namely, legal, and
+medical.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.3\linewidth]{../graphics/fundus_image1.jpg}
-  \caption{Example of Fundus image}
-  \label{fig:fundus-image1}
-\end{figure}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.5\linewidth]{../graphics/oct_fundus.png}
-  \caption{Example of OCT on Retina}
-  \label{fig:oct-retina1}
-\end{figure}
+This thesis takes a look at how we can examine an artificial neural network (ANN)
+to understand which higher level concepts (HLC) it deems important for a given state
+within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
+complicated like Chess or Go. If an ANN values a HLC greatly which we as humans don't find
+very important, and yet the ANN is able to defeat us there is reason to think
+that HLC should be of more importance to us.
 
-Patients that have DR are grouped into four different stages\cite{Intro_FourStages}:
+The chess engine DeepBlue was the first chess engine to win a chess match agains a
+chess grand master. It's power in the world of chess was a marvel. DeepBlue used tree-
+search to evaluate it's position and possible moves, doing an immense amount of calcluations
+in order to arrive at the best move it could given the time constraints. The tree search
+approach is dependent on knowing which board state is good and what board state is bad,
+This can be easy when the boardstate is such that you're guaranteed to have been checkmated
+in one move, and it's incredibly hard when you boardstate is such that both you and your
+opponent have 16 pieces and are only a handfull of moves in. Generally the boardstate
+is a function of the pieces you have, how safe they are, and how opportunistic you are to
+capture your opponents pieces. These values are not obvious and require a domain expert
+to describe the situations and give the values for these cases.
 
-1.
-\textit{Mild Nonproliferative Retinopathy} - This is the earliest stage of diabetic retinopathy, and it’s characterized by balloon-like swelling in the retina’s blood vessels.
-These are called microaneurysms, and these vessels can leak into the eye.
+The process of evaluating chess positions did not change for many years. With the revival
+of deep neural networks this changed. A neural network is a network of connected nodes (neurons)
+that each how a specific weight. These networks are fed input which is passed through the
+network and it is updated by the weights of the neurons, the input is morphed into
+a more comprehensible value and these values are then used to generate some output.
+The output can be a binary classification for instance input=Image output=\{cat, dog\}.
+Processing the pixels and altering them to numerical values to find the output cat or dog
+is not intuitive and understanding of the values of the neurons in the network
+is not considered fruitful.
 
-2.
-\textit{Moderate Nonproliferative Retinopathy} - During this stage, the blood vessels nourishing the retina swell and may even become blocked.
-This can contribute to diabetic macular edema (DME) which is a build-up of fluid in the macula region of the retina.
+This approach of
+chess engines is still popular today but not the reigning champion anymore. In todays
+best chess engines deep neural networks are used to
 
-3.
-\textit{Severe Nonproliferative Retinopathy} -  At this stage, an increasing number of blood vessels nourishing the eye have become blocked.
-As a result, the retina is signaled to grow new blood vessels.
+Deep neural network machine learning models have an amazing impact on modern society today,
+we use them to know which items to purchase, to know which people we should love, and we
+use them to drive us to work. These tasks given to deep learning models are all niceties
+for us to enjoy. However, as our dependance on them increases and their astounding success
+in any and all tasks continues, we will want to utilize them for all things. This leads
+to a utopia where decision making is removed from the hands of humans. The question we
+must ask ourselves is when do we want to know why a decision was taken, is it only when
+the decision taken is not preferrable to us. For instance, if a deep neural network model
+used by your lending authority decides you are not worty of a loan. These questions should
+have answers but as it stands today no such answer is possible.
 
-4.
-\textit{Proliferative Diabetic Retinopathy} - This is the final stage of diabetic retinopathy.
-New blood vessels proliferate, growing inside the retina and into the vitreous gel, which is the fluid that fills the eye.
-Because these blood vessels are delicate, they may begin to leak and bleed.
-As a result, scar tissue may form, causing retinal detachment, the pulling away of the retina from underlying tissue.
-Retinal detachment may cause spotty vision, flashes of lights, or severe vision loss.
+We can use methods like saliency maps to map a deep learning model emphasis on pixels to
+hopefully understand what portions the model is focusing on. This method however, tells us
+nothing about what
 
-Generally without intervention patients advance from one stage to the next eventually leading 
-to permanent vision impairment, however with intervention by an ophthalmologist this advancement 
-can be prevented \cite{Intro_StageAdvancement}.
-
-This paper focuses on applying machine learning methods to patient information and Fundus images of those 
-patients to assess the likelyhood (risk) of those patients DR developing into SDR.
-
-In this paper we will apply machine learning methods to patients information and fundus images of patients 
-that has been collected a period of time to examine a patiens likelyhood to advance from one stage 
-of DR to the next, attempting to predict the timeline the patient has to seek medical help.
-
-\section{Machine learning methods in Ophthalmology}
-
-Applying machine learning to fundus images and patient information has been done before very successfully 
-researchers\cite{Intro_googleDeepmind}, this work generally focuses on applying image recognition on 
-the images to identify which stage of DR the eye has reached.
-
-\section{Background}
-\ifdraft{Provide background about the subject matter (e.g. How was morse code
-developed?  How is it used today?). 
-This is a place where there are usually many citations.
-It is suspicious when there is not. 
-Include the purpose of the different equipment and your design intent. 
-Include references to relevant scientific/technical work and books.}
-
-Applying statistics to medical data in order to better understand a patients evolution from DR to SDR 
-has been done previously by a company called RetinaRisk, a significant contributor to this paper, but 
-RetinaRisk's solution does not apply machine learning but a specialists field knowledge and a handcrafted 
-formula. RetinaRisks formula has been used to significantly reduce costs to health organizations as 
-when patients following their recommendations for doctor checkups come in less often but at time that 
-are more critical. im just writing something so that it looks like I'm writing something
+results that are nothing but amazing.

### 2020-01-26 latex/sections/managerial.tex
updating skeleton

@@
 \begin{dedications}
-  I dedicate this to my spouse/child/pet/power animal.
+  I dedicate this thesis to my family who while not understanding most of what I do
+  always support me. The friends that still want to talk to me even though my
+  schedule has rendered me unmeetupwithable. Lastly but certainly not least,
+  Hulda Lilja who always easead my mind during my spurts of imposter syndrome, and fear of
+  failure.
+
+  Don't Panic.
 \end{dedications}
@@
 MSc &Masters of Science\\
-DR & Diabetic Retinopathy\\
 ML & Machine Learning\\
@@
 \end{tabular}
-
-% \coverchapter{List of Symbols}%%RUM: Not mentioned
-% \begin{tabular}{lll}
-% Symbol &Description &Value/Units\\
-% $E$ &Energy &\si{\joule}\\
-% $m$ &Mass &\si{gram}\\
-% $c$ &Speed of Light &\SI{2.99E8}{\meter\per\second\square}\\
-% \end{tabular}

### 2020-01-26 Stockfish/src/evaluate.cpp

@@ -747,9 +747,10 @@ Value Evaluation<T>::value() {
   // Probe the material hash table
   me = Material::probe(pos);
 
+  // skipping transposition table check ---- siggi
   // If we have a specialized evaluation function for the current material
   // configuration, call it and return.
-  if (me->specialized_eval_exists()) return me->evaluate(pos);
+  // if (me->specialized_eval_exists()) return me->evaluate(pos);
 
   // Initialize score by reading the incrementally updated scores included in
   // the position object (material + piece square tables) and the material
@@ -762,13 +763,20 @@ Value Evaluation<T>::value() {
 
   // Early exit if score is high
   Value v = (mg_value(score) + eg_value(score)) / 2;
-  if (abs(v) > LazyThreshold + pos.non_pawn_material() / 64)
-    return pos.side_to_move() == WHITE ? v : -v;
+
+  // skipping lazy stop ---- siggi
+  // if (abs(v) > LazyThreshold + pos.non_pawn_material() / 64)
+  // return pos.side_to_move() == WHITE ? v : -v;
 
   // Main evaluation begins here
 
   initialize<WHITE>();
   initialize<BLACK>();
+
+  Score piecesW = pieces<WHITE, KNIGHT>() - pieces<BLACK, KNIGHT>() +
+                  pieces<WHITE, BISHOP>() - pieces<BLACK, BISHOP>() +
+                  pieces<WHITE, ROOK>() - pieces<BLACK, ROOK>() +
+                  pieces<WHITE, QUEEN>() - pieces<BLACK, QUEEN>();
 
   // Pieces should be evaluated first (populate attack tables)
   score += pieces<WHITE, KNIGHT>() - pieces<BLACK, KNIGHT>() +
@@ -791,6 +799,23 @@ Value Evaluation<T>::value() {
 
   v /= PHASE_MIDGAME;
 
+  auto kingwmg = mg_value(king<WHITE>());
+  auto kingbmg = mg_value(king<BLACK>());
+  auto kingweg = eg_value(king<WHITE>());
+  auto kingbeg = eg_value(king<BLACK>());
+  auto pawnwmg = mg_value(pe->pawn_score(WHITE));
+  auto pawnbmg = mg_value(pe->pawn_score(BLACK));
+  auto pawnweg = eg_value(pe->pawn_score(WHITE));
+  auto pawnbeg = eg_value(pe->pawn_score(BLACK));
+
+  // Here I should start writing to csv all the values
+
[…]

### 2020-01-26 siggichesszero/breakthrough.py

@@ -1,0 +1,130 @@
+import numpy as np
+from mcts import Node
+from copy import deepcopy
+import random
+
+BREAKTHROUGHWHITE = 1
+BREAKTHROUGHBLACK = 2
+BREAKTHROUGHEMPTY = 0
+
+class BTBoard(Node):
+    def __init__(self, board_list, player_turn):
+        self.board = board_list
+        self.rows  = len(board_list)
+        self.cols  = len(board_list[0])
+        self.turn  = player_turn
+        self.terminal = self.is_terminal_helper()
+        self.children = self.get_children_helper()
+
+    def print_board(self):
+        print("BOARD STATE")
+        print("Turn: ", "White" if self.turn == BREAKTHROUGHWHITE else "Black")
+        for row in self.board:
+            print("".join([str[c] for c in row]))
+        print("==================")
+
+    def copy_board(self):
+        return deepcopy(self.board)
+
+    def find_random_child(self):
+        return random.Choice(self.children)
+
+    def gen_move(self,fpos,tpos):
+        fposy, fposx = fpos
+        tposy, tposx = tpos
+        new_board = self.copy_board()
+        new_board[tposy][tposx] = new_board[fposy][fposx]
+        new_board[fposy][fposx] = BREAKTHROUGHEMPTY
+        next_player = BREAKTHROUGHWHITE if self.turn == BREAKTHROUGHBLACK else BREAKTHROUGHBLACK
+        b_new_board = BTBoard(new_board, next_player)
+        return b_new_board
+
+    def get_children(self) -> set:
+        return self.children
+
+    def get_dirs(self, pos) -> list:
+        "pos is coordinates in list [y,x]"
+        dirs = []
+        do_left = pos[1] > 0
+        do_right = pos[1] < (self.cols - 1)
+        print("getdirs with ",pos)
+        ymod = -1 if self.turn == BREAKTHROUGHWHITE else 1 # y axis mod
+        enemy = BREAKTHROUGHWHITE if self.turn == BREAKTHROUGHBLACK else BREAKTHROUGHBLACK
+        if do_left and self.board[pos[0]+ymod][pos[1]-1] == enemy:
+            dirs.append([pos[0]+ymod,pos[1]-1])
+        if do_right and self.board[pos[0]+ymod][pos[1]+1] == enemy:
+            dirs.append([pos[0]+ymod,pos[1]+1])
[…]

### 2020-01-26 siggichesszero/game.py

@@ -1,0 +1,46 @@
+from mcts import MCTS
+from playchess import PlayChess
+import timeit
+from multiprocessing import Pool
+
+def play_game(white_think, black_think):
+  print("playgame started with {} {}".format(white_think, black_think))
+  whiteplayer = MCTS()
+  blackplayer = MCTS(negamaxing=True)
+  board = PlayChess()
+  while True:
+    for _ in range(white_think):
+        whiteplayer.do_rollout(board)
+    board = whiteplayer.choose_child(board)
+    # print("===================")
+    # board.print_board()
+    # print("===================")
+    if board.is_terminal():
+      break
+
+    for _ in range(black_think):
+        blackplayer.do_rollout(board)
+    board = blackplayer.choose_child(board)
+    print("===================")
+    board.print_board()
+    print("===================")
+    if board.terminal:
+        break
+  return board.board.result(claim_draw=True)
+
+test_times = 5
+with open("results.out",'w') as f:
+  for j in range(24,25):
+    for k in range(1,10):
+      f.write(f"Testing with white think {j} and black think {k}\n")
+      time = 0
+      print(f"Testing with white think {j} and black think {k}")
+      with Pool(test_times) as p:
+        print(p.starmap(play_game, [(j,k), (j,k), (j,k), (j,k), (j,k)]))
+      f.write(f"game length average {time}\n")
+
+
+
+
+
+

### 2020-01-26 siggichesszero/mcts.py

@@ -1,0 +1,105 @@
+from collections import defaultdict
+from mctsnode import Node
+import math
+
+class MCTS:
+  """
+    'Monte Carlo Tree Search
+
+    do random moves to average the score for a given node, this is good when it's
+    hard to actually evaluate a position on a board given just the position,
+    then we just simulate a bunch of moves from it and average the results
+
+    variables of MCTS:
+      Q total reward from nodes
+      N total visit count for nodes
+      children array of children nodes for a node
+      exploration weight constant
+
+    notes:
+      if a node is in the children dict it is expanded
+
+  """
+  def __init__(self, negamaxing=False, exploration_weight=1):
+    self.Q = defaultdict(int)
+    self.N = defaultdict(int)
+    self.children = dict()
+    self.exploration_weight = exploration_weight
+    self.negamaxing = negamaxing
+
+  def mcts_score(self, node) -> float:
+    if self.N[node] == 0:
+      # unknown nodes are not considered
+      return float("-inf")
+    "Average Score - is the sum of its rewards divided by the amount of visits"
+    return self.Q[node] / self.N[node]
+
+  def choose_child(self, node: Node) -> Node:
+    "Choose the best child of a node"
+    assert(not node.is_terminal(), "Calling choose child on a terminal node, is not allowed")
+
+    if node not in self.children:
+      return node.find_random_child()
+
+    return max(self.children[node], key=self.mcts_score)
+
+  def do_rollout(self, node: Node) -> None:
+    """
+      Rollout of nodes. We expand a node and do random moves to estimate the
+      actual undelying reward from doing the move
+    """
+    path = self.select(node)
+    leaf = path[-1]
+    self.expand(leaf)
+    reward = self.mcts_simulate(leaf)
+    self.backpropagate(path, reward)
+
+  def select(self, curr_node: Node) -> list:
+    """
+      function should not be called directly
+      it's called during rollout.
+
+      Function finds an unexpanded descendant of *node*
+    """
+    path = []
+    while True:
+      path.append(curr_node)
+      if curr_node not in self.children or curr_node.is_terminal():
+        return path
+      unexplored = self.children[curr_node] - self.children.keys()
+      if unexplored:
+        end = unexplored.pop()
+        path.append(end)
+        return path
+      curr_node = self.uct_select(curr_node) # randomly select a child
+    raise RuntimeError("Select didn't find a path")
+
[…]

### 2020-01-26 siggichesszero/playchess.py

@@ -1,0 +1,68 @@
+import chess
+import copy
+from mctsnode import Node
+import random
+
+class PlayChess(Node):
+
+  rewards_dict = {
+    '1/2-1/2': 0,
+    '1-0': 1,
+    '0-1': -1
+  }
+
+  def __init__(self, fen=""):
+    if fen:
+      self.board = chess.Board(fen)
+    else:
+      self.board = chess.Board()
+    self.white_turn = self.board.turn
+    self.terminal = self.board.is_game_over(claim_draw=True)
+    self.moves = list(self.board.legal_moves)
+
+  def is_terminal(self) -> bool:
+    return self.terminal
+
+  def reward(self) -> int:
+    assert(not self.is_terminal(), "calling reward on non terminal node is not allowed")
+    return self.rewards_dict[self.board.result(claim_draw=True)]
+
+  def find_random_child(self):
+    if self.is_terminal():
+      return None
+    self.board.push(random.choice(self.moves))
+    new_board = PlayChess(self.board.fen())
+    self.board.pop()
+    return new_board
+
+  def get_children(self) -> set:
+    children = set()
+    for m in self.moves:
+      self.board.push(m)
+      children.add(PlayChess(self.board.fen()))
+      self.board.pop()
+    return children
+
+  def make_move(self, move: chess.Move):
+    self.board.push(move)
+    new_board = copy.deepcopy(self.board)
+    self.board.pop()
+    return PlayChess(new_board.fen())
+
+  def print_board(self) -> None:
+    print(self.board)
+
+
+  """
+    The magic functions utilized just abuse the string magic functions
+    assuming they work perfectly... it should be fine.
+  """
+  def __str__(self):
+    return self.board.fen()
+
+  def __hash__(self):
+    return str.__hash__(self.board.fen())
+
+  def __eq__(self, other):
+    return str(self) == str(other)
+

### 2020-01-26 siggichesszero/test.py

@@ -1,0 +1,300 @@
+import torch
+import torch.nn as nn
+import torchvision.transforms as transforms
+import torchvision.datasets as dsets
+
+class ConvBlock(nn.Module):
+  def __init__(self):
+    super(ConvBlock, self).__init__()
+    self.conv1      = nn.Conv2d(in_channels=3, out_channels=128, kernel_size=3, stride=1, padding=1, bias=False)
+    self.batchnorm1 = nn.BatchNorm2d(128)
+    self.relu1      = nn.ReLU()
+
+  def forward(self,input):
+    output = self.conv1(input)
+    output = self.batchnorm1(output)
+    return self.relu1(output)
+
+class ResBlock(nn.Module):
+  def __init__(self):
+    super(ResBlock, self).__init__()
+    self.conv1 = nn.Conv2d(
+      in_channels=128,
+      out_channels=128,
+      kernel_size=3,
+      stride=1,
+      bias=False
+    )
+    self.batchnorm1 = nn.BatchNorm2d(128)
+    self.relu1      = nn.ReLU()
+    self.conv2 = nn.Conv2d(
+      in_channels=128,
+      out_channels=128,
+      kernel_size=3,
+      stride=1,
+      bias=False
+    )
+    self.batchnorm2 = nn.BatchNorm2d(128)
+    self.relu2      = nn.ReLU()
+
+  def forward(self, input):
+    residual = input
+    output = self.conv1(input)
+    output = self.batchnorm1(output)
+    output = self.relu1(output)
+    output = self.conv2(input)
+    output = self.batchnorm2(output)
+    # add the residual before reluing
+    output += residual
+    output = self.relu2(output)
+    return output
+
+class OutBlock(nn.Module):
+  def __init__(self):
+    super(OutBlock, self).__init__()
+    # Value head
+    self.conv_v = nn.Conv2d(
+      in_channels=128,
+      out_channels=3,
+      kernel_size=1
+    )
+    self.batchnorm_v = nn.BatchNorm2d(3)
+    self.relu_v_1 = nn.ReLU()
+    self.full_conn_1_v = nn.Linear(128, 32)
+    self.relu_v_2 = nn.ReLU()
+    self.full_conn_2_v = nn.Linear(32,1)
+    self.tahn_v = nn.Tanh()
+
+    # Policy head
+    self.conv_p = nn.Conv2d(
+      in_channels=128,
[…]

### 2020-01-26 siggichesszero/test_chess.py

@@ -1,0 +1,40 @@
+import chess
+import queue
+import copy
+
+a = chess.Board("3k4/8/8/8/8/8/8/K7 b - - 0 1")
+
+print(a.is_check())
+print(a.is_checkmate())
+print(a.is_game_over(claim_draw=True))
+print(a.is_variant_draw())
+print(a.is_variant_win())
+print(a.is_variant_loss())
+print(a.result())
+exit()
+
+frontier = queue.Queue()
+frontier.put((a,0))
+while not frontier.empty():
+  board, depth = frontier.get()
+  if depth > 2:
+    continue
+  if board.is_game_over():
+    print("board is game over ")
+    print(board)
+    break
+  print("examining board")
+  print(board)
+  print("=======================")
+  for m in board.generate_legal_moves():
+    board.push(m)
+    frontier.put((copy.deepcopy(board),depth+1))
+    board.pop()
+
+"""
+for m in a.generate_legal_moves():
+  a.push(m)
+  print(a)
+  a.pop()
+  print("================")
+  """

### 2020-01-26 siggichesszero/test_init.py

@@ -1,0 +1,16 @@
+import chess
+import time
+
+times = []
+for i in range(1,3010, 100):
+    print("initializing {} many chessboards".format(i))
+    start = time.time()
+    for j in range(1,i):
+        chess.Board().is_game_over(claim_draw=False)
+    times.append(time.time() - start)
+
+import matplotlib.pyplot as plt
+print(times)
+plt.plot(times)
+plt.show()
+

### 2020-01-27 hlc_evaluator/monte_carlo/mcts.py
hlc evaluator added

@@ -1,0 +1,107 @@
+from collections import defaultdict
+from monte_carlo.mctsnode import Node
+import math
+import chess
+
+
+class MCTS:
+    """
+      'Monte Carlo Tree Search
+
+      do random moves to average the score for a given node, this is good when it's
+      hard to actually evaluate a position on a board given just the position,
+      then we just simulate a bunch of moves from it and average the results
+
+      variables of MCTS:
+        Q total reward from nodes
+        N total visit count for nodes
+        children array of children nodes for a node
+        exploration weight constant
+
+      notes:
+        if a node is in the children dict it is expanded
+
+    """
+
+    def __init__(self, negamaxing=False, exploration_weight=1):
+        self.Q = defaultdict(int)
+        self.N = defaultdict(int)
+        self.children = dict()
+        self.exploration_weight = exploration_weight
+        self.negamaxing = negamaxing
+
+    def mcts_score(self, node) -> float:
+        if self.N[node] == 0:
+            # unknown nodes are not considered
+            return float("-inf")
+        "Average Score - is the sum of its rewards divided by the amount of visits"
+        return self.Q[node] / self.N[node]
+
+    def choose_child(self, node: Node) -> Node:
+        "Choose the best child of a node"
+        assert not node.is_terminal(), "Calling choose child on a terminal node, is not allowed"
+
+        if node not in self.children:
+            return node.find_random_child()
+
+        return max(self.children[node], key=self.mcts_score)
+
+    def do_rollout(self, node: Node) -> None:
+        """
+          Rollout of nodes. We expand a node and do random moves to estimate the
+          actual undelying reward from doing the move
+        """
+        path = self.select(node)
+        leaf = path[-1]
+        self.expand(leaf)
+        reward = self.mcts_simulate(leaf)
+        self.backpropagate(path, reward)
+
+    def select(self, curr_node: Node) -> list:
+        """
+          function should not be called directly
+          it's called during rollout.
+
+          Function finds an unexpanded descendant of *node*
+        """
+        path = []
+        while True:
+            path.append(curr_node)
+            if curr_node not in self.children or curr_node.is_terminal():
+                return path
+            unexplored = self.children[curr_node] - self.children.keys()
+            if unexplored:
+                end = unexplored.pop()
+                path.append(end)
+                return path
+            curr_node = self.uct_select(curr_node)  # randomly select a child
+
[…]

### 2020-01-27 hlc_evaluator/monte_carlo/mctsnode.py

@@ -1,0 +1,41 @@
+from abc import ABC, abstractmethod
+class Node(ABC):
+  """
+    Generic class for states in a game to do MCTS on
+  """
+
+  @abstractmethod
+  def get_children(self) -> set:
+    "Returns all children for a state (after applying an action)"
+    return set()
+
+  @abstractmethod
+  def find_random_child(self) -> object:
+    "Returns a random child for this state"
+    return None
+
+  @abstractmethod
+  def is_terminal(self) -> bool:
+    "Returns true if this state is a terminal state"
+    return True
+
+  @abstractmethod
+  def reward(self) -> int:
+    """returns the reward for the current node (should be a terminal node
+    otherwise methods other than mcts would be better)"""
+    return 0
+
+  @abstractmethod
+  def __str__(self) -> str:
+    "A node should have a string representation (it's used in error messages)"
+    return ""
+
+  @abstractmethod
+  def __hash__(self) -> int:
+    "a node has to be hashable to do efficient lookup"
+    return 1
+
+  @abstractmethod
+  def __eq__(self, other) -> bool:
+    "We need comparison of nodes"
+    return True

### 2020-01-27 latex/sections/background.tex
latex work

@@
 \section{Background}
+
+\subsection{Explainable Artificial Intelligence}
+
+Within XAI many methods have been developed to try to evaluate
+ANN's, these methods are often referred as model interpretability methods.
+In the field image recognition there has been a lot of work examining which
+pixels of an image the model deems important. Arguably the most popular
+method for this is Saliency Maps (CITE ME). There the pixel values the model
+deems important are colored in s.t. a human can examine the image and get a
+sense of what portions of the image are important to the model, an example of
+a classification of a dog can be seen in Figure \ref{fig:dog_saliency}.
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=.5\textwidth]{graphics/dog_saliency}
+  \caption{Example of a models saliency map for an image of a dog}
+  \label{fig:dog_saliency}
+\end{figure}
 

### 2020-01-27 latex/sections/introduction.tex
latex work

@@
 cancer cells where others have not. Deep learning models have even predicted
-the likelyhood a person will die in the following year. These models have done
-these things extremely accurately, cheap, and fast. These successes allow us to
-dream of a future where all things are eventually possible with the help of
-technology.
+the likelyhood a person will die in the following year with good accuracy (CITE).
+These models have done these things extremely accurately, cheap, and fast.
 
@@
 ourselves but these explanations are a legal requirement now in many areas, namely, legal, and
-medical.
+medical. Recently, XAI has generally been focused on Model Interpretability, in
+particular focus on gaining insights on the concepts learned by Deep Neural Networks.
 
@@
 that HLC should be of more importance to us.
+
+The method of examining HLC's within games has generally been done by examining the
+current state of the game by evaluating them w.r.t. a heuristic. A heuristic within
+games are a evaluation of the end cost for a state given just the state, for example,
+the amount of pieces left within a game of chess. Intuitively, the amount of pieces
+left is a good estimate for a state in chess, this is a higher level concept we
+use to evaluate a chess position. The piece amount is much lower level than other
+concepts we use, grand master chess players evaluate a position w.r.t. states
+where the king is safe from attack, or the structure of the pawn positions.
+The idea is to examine a neural networks evaluation of a state regarding
+those higher level concepts, if human players evaluate the king safety of a state
+low but the neural network highly values it, and the neural network plays better
+than the player. There is reason for us to consider a new way to play.
 

### 2020-01-27 hlc_evaluator/game_environments/breakthrough/breakthrough.py
selfplay refactor

@@ -1,4 +1,4 @@
-import numpy as np
+#import numpy as np
 from monte_carlo.mctsnode import Node
 from copy import deepcopy
 import random
@@ -20,14 +20,14 @@ class BTBoard(Node):
         print("BOARD STATE")
         print("Turn: ", "White" if self.turn == BREAKTHROUGHWHITE else "Black")
         for row in self.board:
-            print("".join([str[c] for c in row]))
+            print("".join([str(c) for c in row]))
         print("==================")
 
     def copy_board(self):
         return deepcopy(self.board)
 
     def find_random_child(self):
-        return random.Choice(self.children)
+        return random.choice(self.children)
 
     def gen_move(self, fpos, tpos):
         fposy, fposx = fpos

### 2020-01-27 hlc_evaluator/game_environments/play_chess/playchess.py

@@ -11,7 +11,7 @@ class PlayChess(Node):
         '0-1': -1
     }
 
-    def __init__(self, fen=""):
+    def __init__(self, fen="", player_turn=None):
         if fen:
             self.board = chess.Board(fen)
         else:

### 2020-01-27 hlc_evaluator/selfplay.py

@@ -2,14 +2,37 @@ from monte_carlo.mcts import MCTS
 from game_environments.play_chess.playchess import PlayChess
 from game_environments.breakthrough.breakthrough import BTBoard
 
-# Change GAME variable to change selfplayed game
-GAME = BTBoard
+from collections import namedtuple
+
+game_setup = namedtuple("gamesetup", "game_class initial_state")
+
+BREAKTHROUGH_INITIAL_STATE = [
+    [2, 2, 2],
+    [0, 0, 0],
+    [1, 1, 1]
+]
+
+CHESS_INITIAL_STATE = "rnbqkbnr/pppppppp/8/8/8/8/PPPPPPPP/RNBQKBNR w KQkq - 0 1"
+
+GAME = {
+    "breakthrough": game_setup(BTBoard, BREAKTHROUGH_INITIAL_STATE),
+    "chess": game_setup(PlayChess, CHESS_INITIAL_STATE),
+}
 
 def play_game(white_think, black_think):
+    """
+    Function that plays a game between two players
+    using monte carlo tree search, white_think and
+    black_think reference the amount of mcts rollout
+    each player does
+    """
     print("playgame started with {} {}".format(white_think, black_think))
     whiteplayer = MCTS()
     blackplayer = MCTS(negamaxing=True)
-    board = GAME()
+
+    selected_game = GAME["breakthrough"]
+
+    board = selected_game.game_class(selected_game.initial_state, 1)
     while True:
         for _ in range(white_think):
             whiteplayer.do_rollout(board)
@@ -28,13 +51,8 @@ def play_game(white_think, black_think):
         print("===================")
         if board.terminal:
             break
-    return board.board.result(claim_draw=True)
+    return board.reward()
 
-for j in range(24,25):
-    for k in range(1,10):
-        print(play_game(j,k))
-
-
-
-
-
+for j in range(24, 25):
+    for k in range(1, 10):
+        print(play_game(j, k))

### 2020-03-11 latex/sections/introduction.tex
latex iterations

@@
 
-This thesis takes a look at how we can examine an artificial neural network (ANN)
+In this we thesis takes a look at how we can examine an artificial neural network (ANN)
 to understand which higher level concepts (HLC) it deems important for a given state
 within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
-complicated like Chess or Go. If an ANN values a HLC greatly which we as humans don't find
-very important, and yet the ANN is able to defeat us there is reason to think
-that HLC should be of more importance to us.
+complicated like Chess or Go. Some of the research questions are what if an ANN values a HLC greatly
+which we as humans don't find particularily important, and yet the ANN is able to defeat
+us what does that mean?
 
 The method of examining HLC's within games has generally been done by examining the
-current state of the game by evaluating them w.r.t. a heuristic. A heuristic within
+current state of the game by evaluating them with respect to a heuristic. A heuristic within
 games are a evaluation of the end cost for a state given just the state, for example,
@@
 left is a good estimate for a state in chess, this is a higher level concept we
-use to evaluate a chess position. The piece amount is much lower level than other
-concepts we use, grand master chess players evaluate a position w.r.t. states
+use to evaluate a chess position. The piece amount can be considered as a lower-level concept
+than other concepts we use. For instance, grand master chess players evaluate a position w.r.t. states
 where the king is safe from attack, or the structure of the pawn positions.
 The idea is to examine a neural networks evaluation of a state regarding
-those higher level concepts, if human players evaluate the king safety of a state
+those higher-level concepts, if human players evaluate the king safety of a state
 low but the neural network highly values it, and the neural network plays better
-than the player. There is reason for us to consider a new way to play.
+than the player. There could be an avenue for us to improve our game by considering king safety more thoroughly.
 
-The chess engine DeepBlue was the first chess engine to win a chess match agains a
-chess grand master. It's power in the world of chess was a marvel. DeepBlue used tree-
-search to evaluate it's position and possible moves, doing an immense amount of calcluations
-in order to arrive at the best move it could given the time constraints. The tree search
-approach is dependent on knowing which board state is good and what board state is bad,
-This can be easy when the boardstate is such that you're guaranteed to have been checkmated
-in one move, and it's incredibly hard when you boardstate is such that both you and your
-opponent have 16 pieces and are only a handfull of moves in. Generally the boardstate
-is a function of the pieces you have, how safe they are, and how opportunistic you are to
-capture your opponents pieces. These values are not obvious and require a domain expert
-to describe the situations and give the values for these cases.
+In this thesis we take an example game of Breakthrough, train a neural network to play the game
+using only self-play. That is, the neural network is trained only by playing against
+itself with no outside input other than the rules of the game itself. And we examine
+the neural networks against popular Higher-level concepts (heuristics) that we use
+to play the game.
 
-The process of evaluating chess positions did not change for many years. With the revival
-of deep neural networks this changed. A neural network is a network of connected nodes (neurons)
-that each how a specific weight. These networks are fed input which is passed through the
-network and it is updated by the weights of the neurons, the input is morphed into
-a more comprehensible value and these values are then used to generate some output.
-The output can be a binary classification for instance input=Image output=\{cat, dog\}.
-Processing the pixels and altering them to numerical values to find the output cat or dog
-is not intuitive and understanding of the values of the neurons in the network
-is not considered fruitful.
+The neural network architecture is based of the popular AlphaZero neural network
+used to play Chess, Shogi, and Go on a superhuman level. (REMOVE ME) Unfortunately due to lack
+of thousands of TPU's this research can't go as far as to examine those more complicated
+games. (REMOVE ME)
 
-This approach of
-chess engines is still popular today but not the reigning champion anymore. In todays
-best chess engines deep neural networks are used to
+% The chess engine DeepBlue was the first chess engine to win a chess match agains a
+% chess grand master. It's power in the world of chess was a marvel. DeepBlue used tree-
+% search to evaluate it's position and possible moves, doing an immense amount of calcluations
+% in order to arrive at the best move it could given the time constraints. The tree search
+% approach is dependent on knowing which board state is good and what board state is bad,
+% This can be easy when the boardstate is such that you're guaranteed to have been checkmated
+% in one move, and it's incredibly hard when you boardstate is such that both you and your
+% opponent have 16 pieces and are only a handfull of moves in. Generally the boardstate
+% is a function of the pieces you have, how safe they are, and how opportunistic you are to
+% capture your opponents pieces. These values are not obvious and require a domain expert
+% to describe the situations and give the values for these cases.
 
-Deep neural network machine learning models have an amazing impact on modern society today,
-we use them to know which items to purchase, to know which people we should love, and we
-use them to drive us to work. These tasks given to deep learning models are all niceties
-for us to enjoy. However, as our dependance on them increases and their astounding success
-in any and all tasks continues, we will want to utilize them for all things. This leads
-to a utopia where decision making is removed from the hands of humans. The question we
-must ask ourselves is when do we want to know why a decision was taken, is it only when
-the decision taken is not preferrable to us. For instance, if a deep neural network model
-used by your lending authority decides you are not worty of a loan. These questions should
-have answers but as it stands today no such answer is possible.
+% The process of evaluating chess positions did not change for many years. With the revival
+% of deep neural networks this changed. A neural network is a network of connected nodes (neurons)
+% that each how a specific weight. These networks are fed input which is passed through the
+% network and it is updated by the weights of the neurons, the input is morphed into
+% a more comprehensible value and these values are then used to generate some output.
+% The output can be a binary classification for instance input=Image output=\{cat, dog\}.
+% Processing the pixels and altering them to numerical values to find the output cat or dog
+% is not intuitive and understanding of the values of the neurons in the network
+% is not considered fruitful.
 
-We can use methods like saliency maps to map a deep learning model emphasis on pixels to
-hopefully understand what portions the model is focusing on. This method however, tells us
-nothing about what
+% This approach of
+% chess engines is still popular today but not the reigning champion anymore. In todays
+% best chess engines deep neural networks are used to
 
-results that are nothing but amazing.
+% Deep neural network machine learning models have an amazing impact on modern society today,
+% we use them to know which items to purchase, to know which people we should love, and we
+% use them to drive us to work. These tasks given to deep learning models are all niceties
+% for us to enjoy. However, as our dependance on them increases and their astounding success
+% in any and all tasks continues, we will want to utilize them for all things. This leads
+% to a utopia where decision making is removed from the hands of humans. The question we
+% must ask ourselves is when do we want to know why a decision was taken, is it only when
+% the decision taken is not preferrable to us. For instance, if a deep neural network model
+% used by your lending authority decides you are not worty of a loan. These questions should
+% have answers but as it stands today no such answer is possible.
+
+% We can use methods like saliency maps to map a deep learning model emphasis on pixels to
+% hopefully understand what portions the model is focusing on. This method however, tells us
+% nothing about what
+
+% results that are nothing but amazing.

### 2020-03-11 hlc_evaluator/trainer.py

@@ -31,7 +31,7 @@ initial_state = selected_game.initial_state()
   Config varibles
 """
 EPISODE_AMOUNT = 5
-NEURAL_NETWORK_THINK = 25
+NEURAL_NETWORK_THINK = 50
 TEMP_THRESHOLD = 6
 TRAINING_ITERS = 4
 VERIFICATION_GAMES = 100

### 2020-03-11 hlc_evaluator/verifier.py

@@ -20,7 +20,7 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
   neural_network_1 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
   neural_network_2 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
 
-  neural_network_1.loadmodel(first_network_path, first_network_name)
+  neural_network_2.loadmodel(first_network_path, first_network_name)
 
   initial_node = Node(state_example.initial_state(), "START")
   first_win = 0

### 2020-03-24 hlc_evaluator/game_environments/breakthrough/breakthrough.py
minor changes

@@ -108,7 +108,7 @@ class BTBoard(GameNode):
                 elif self.board[y,x] == config.BLACK:
                     enc_board[y,x,1] = True
         if self.player == config.WHITE:
-            enc_board[:,:,2] = 1
+            enc_board[:,:,2] = True
         return enc_board
 
     def print_board(self):

### 2020-03-24 hlc_evaluator/neural_networks/breakthrough/breakthrough_nn.py

@@ -99,7 +99,7 @@ class BreakthroughNN(NNBase):
 
       self.scheduler.step()
       losses.append(total_loss)
-    print(f"[breakthrough_nn.py train()] total losses for training were {losses}")
+    # print(f"[breakthrough_nn.py train()] total losses for training were {losses}")
 
   def savemodel(self, path, filename):
     filepath = os.path.join(path,filename)

### 2020-03-24 hlc_evaluator/trainer.py

@@ -31,12 +31,12 @@ initial_state = selected_game.initial_state()
   Config varibles
 """
 EPISODE_AMOUNT = 5
-NEURAL_NETWORK_THINK = 25
-TEMP_THRESHOLD = 6
-TRAINING_ITERS = 4
+NEURAL_NETWORK_THINK = 50
+TEMP_THRESHOLD = 250
+TRAINING_ITERS = 20
 VERIFICATION_GAMES = 100
 
-def generate_dataset(primary_nn: BreakthroughNN, game_example : GameNode, saved_monte_tree=None, verbose=False):
+def generate_dataset(primary_nn: BreakthroughNN, game_example : GameNode, generation: int , verbose=False):
   initial_node = Node(game_example.initial_state(), "START")
   monte_tree = MCTS(initial_node, primary_nn)
   dataset = []
@@ -70,10 +70,10 @@ def generate_dataset(primary_nn: BreakthroughNN, game_example : GameNode, saved_
 
   return dataset
 
-def train_model(play_iterations, neural_network: BreakthroughNN, state_example: GameNode):
+def train_model(play_iterations, neural_network: BreakthroughNN, state_example: GameNode, generation: int):
   for _ in tqdm(range(play_iterations)):
-    dataset = generate_dataset(neural_network, state_example)
-    print("Dataset to train on has reward:",dataset[-1][2])
+    dataset = generate_dataset(neural_network, state_example, generation)
+    # print("Dataset to train on has reward:",dataset[-1][2])
     neural_network.train(dataset)
 
 
@@ -161,23 +161,21 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
             second_win += 1
           break
     print("[trainer.py] Episode record was White: {} | Black: {} | Tie: {}".format(first_win, second_win, VERIFICATION_GAMES - (first_win+second_win)))
-    if first_win/VERIFICATION_GAMES > 0.55:
-      print("[trainer.py] First network wins > 55%, saving first")
+    if first_win > second_win:
+      print("[trainer.py] First network wins, saving first")
       neural_network_2.loadmodel(first_network_path, first_network_name)
       neural_network_1.savemodel("./trained_models", "best_network.tar")
-    elif second_win/VERIFICATION_GAMES > 0.55:
-      print("[trainer.py] Second network wins > 55%, saving second")
+      print("[trainer.py] STARTING TRAINING")
+      train_model(TRAINING_ITERS, neural_network_1, state_example, generation)
+    else:
[…]

### 2020-03-24 hlc_evaluator/verifier.py

@@ -1,6 +1,6 @@
 from monte_carlo.mcts import MCTS
 from monte_carlo.mctsnode import Node
-from game_environments.play_chess.playchess import PlayChess
+# from game_environments.play_chess.playchess import PlayChess
 from game_environments.gamenode import GameNode
 from game_environments.breakthrough.breakthrough import BTBoard, config as BTconfig
 from neural_networks.breakthrough.breakthrough_nn import BreakthroughNN
@@ -20,19 +20,29 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
   neural_network_1 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
   neural_network_2 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
 
-  neural_network_1.loadmodel(first_network_path, first_network_name)
+  neural_network_2.loadmodel(first_network_path, first_network_name)
+
+  memo_nn1 = {}
+  memo_nn2 = {}
 
   initial_node = Node(state_example.initial_state(), "START")
   first_win = 0
   second_win = 0
+
   while True:
 
 
-    for _ in tqdm(range(10)):
+    for _ in tqdm(range(10000)):
       curr_node = initial_node
       while True:
         # First NN moves
-        pi,_ = neural_network_1.predict(curr_node.gamestate)
+
+        if curr_node in memo_nn1:
+          pi,val = memo_nn1[curr_node]
+        else:
+          pi,val = neural_network_1.safe_predict(curr_node.gamestate)
+          memo_nn1[curr_node] = (pi,val)
+
         pi = pi.detach().cpu().numpy().reshape(-1)
         if not curr_node.is_expanded():
           curr_node.expand()
@@ -56,7 +66,12 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
           break
 
         # Second NN moves (is playing black)
-        pi,_ = neural_network_2.predict(curr_node.gamestate)
+        if curr_node in memo_nn2:
+          pi,val = memo_nn2[curr_node]
+        else:
+          pi,val = neural_network_2.safe_predict(curr_node.gamestate)
+          memo_nn2[curr_node] = (pi,val)
+
         pi = pi.detach().cpu().numpy().reshape(-1)
         if not curr_node.is_expanded():
           curr_node.expand()

### 2020-03-24 latex/sections/background.tex

@@
 
-\subsection{Explainable Artificial Intelligence}
+\subsection{Explainable Artificial Intelligence (XAI)}
 
+The field of XAI research is still very far behind it's counterpart AI research,
+within XAI two fields are concidered the largest that is Model Interpretability
+and Model Explainability.
+
+\subsubsection{Model Interpretability}
+
+\subsubsubsection{Saliency Maps}
 Within XAI many methods have been developed to try to evaluate
@@
 
-\subsection{Reinforcement Learning}
+While this method is understandable in the context of image recognition it
+lacks severely when your input is not an image.
 
-In 2017 a Google based artificial intelligence research company published a paper called
-`Mastering Chess and Shogi by Self-Play with a General Reinforcement Learning Algorithm`.
-The paper described a Model based Reinforment learning agent and an accompanying Residual
-Deep Neural Network architecture that was able to master the game of chess without being
-given any idea of how to play the game of chess, and others. The architecture used in this
-paper is heavily inspired by that architecture.
+\subsubsubsection{Shapley Values}
 
-\subsection{Model Interpretability}
+Methods for explaining models that aren't image recognition models include
+shapley values. There the input is examined against it's output, then iteratively
+input values are selected to be fixed. Then the other input values are varied and
+an average change in prediction is calculated. With this the shapley value can be
+estimated for the fixed input value. This is done to examine which input values have
+the strongest link to the output value. Shapley values on a dataset can give insights
+on which input values the model deems important.
 
-In 2018 the paper 'Interpretability Beyond Feature Attribution: Quantitative Testing with Concept Activation Vectors (TCAV)'
-was published where an approach to model interpretability was introduced which is able to
-distinguish
+\subsubsubsection{Concept Activation Vectors}
 
-The deep neural network architecture that is used in this paper is based on the AlphaZero paper
-published by DeepMind,
+A recent paper by Been Kim Et. al (CITATION NEEDED), shows a method for examining
+a neural network giving a much more human insight into a prediction. Using Concept
+Activation Vectors (CAV) a directional derivative for a given input can be examined
+with respect to some HLC's. For example, when a human looks at an image of an animal
+and is supposed to decide whether the image is of a horse or a zebra, an intuitive
+approach would be to check whether the animal has stipes, or is both white and black.
+
+CAV's in neural networks are just binary classifiers of the same dataset used to train
+a neural network. The
+
+\subsubsection{Model Explainability}
+
+Model explainability within the context of neural networks isn't possible today. Model
+explainability referrs to firstly considering some input and output from a model. Then
+afterwards the model is examined to determine exactly what led to the predicted output.
+This concept is simple when we're working with Decision Trees. A decision tree is a tree
+whose nodes are representative of an input value and at every node a branch is selected
+based on the value of the input value. It is therefore easy to see how to examine the tree
+to explain the output. We just follow the branches in the tree. That being said the branches
+are created by algorithms like ID3 which construct branches based on the initial dataset used
+to construct the tree, but again ID3 follows simple statistics and can be explained properly.
+
+When we talk about neural networks this process is much more difficult, the underlying nodes
+are generally in the thousands, the different layers of the neural network varies in the operations
+it applies to the input value and such while travelling through the neural network the modified
+value becomes far removed from the initial input value to the eyes of the reader. That being said,
+while the possibility of completely monitoring the training process and completely monitoring the
+evaluation process is truly possible it is not feasible. And secondly the process of seeing an
+input and it's corresponding output will not be of any value if one were to consider the process
+of prediction.
+
+\subsection{Machine Learning}
+
+\subsubsection{Reinforcement Learning}
+
+Reinforcement Learning is a subfield of Machine Learning where a model learns from experience.
+That is, the notion of input/output values changes, the model uses itself to generate input
+values by acting in an environment. The environment then returns some result, that could be
+losing in a game, making a correct prediction, or any number of outcomes. The result is then
+propagated through the model moving it's parameters in the direction of the recieved output
+from the environment.
+
+Many algorithms are popular in reinforcement learning, for instance Q-learning(CITE ME), CARLA (CITE ME),
+Monte Carlo Tree Search, and many others. The work in this paper relies heavily on Monte Carlo Tree Search,
+and will be the only one thoroughly discussed.
+
+\subsubsection{Monte Carlo Tree Search}
+
+In the algorithm Monte Carlo Tree Search there is an agent exists
+within some environment. Where each node in the environment represents a state-action pair of the environment, by
+state-action pair what is meant it is some state and the action that brought the agent to
+that state. This pair should be unique within the environment. We will discuss these concepts within game-environments
+and therefore we might say player instead of agent and game instead of environment.
+
+Monte Carlo Tree Search is a method of exploring an enviroment in a randomized manner (Monte Carlo is the term implying
+randomness). In MCTS there are four stages. Selection, Expansion, Simulation, and Backpropagation. They happen sequentially
+and repeatedly. MCTS is initialized with a tree consisting of the unexpanded initial state of the environment.
+
+In MCTS there is a tree representing the game-environment. This tree consists of nodes $n_i$ where $i$ represents the point in time
+of the node, for example $N_0$ in chess is the initial position and $N_x$ is some position in the middle of the game and $N_e$ is one of
+the states representing a position where there is either a draw or one player has won the match. Each of the nodes has $4$ values,
+$s$, $a$, $Q$, and $N$. These values represent these items, $s$ is state of the environment, $a$ is the action that brought
+the previous node $n_(i-1)$ to node $n_i$, $Q$ is the average value of the node (value meaning the outcome of the game),
+and $N$ the amount of times the MCTS algorithm has visited this node. The values $s$ and $a$ uniquely identify a position
+in the environment and are often called state-action pairs.
+
+The MCTS algorithms four phases
+\begin{enumerate}
+  \item Selection
+  \item Expansion
+  \item Simulation/Rollout
+  \item Backpropagation
+\end{enumerate}
+
+\begin{equation} \label{UCT_formula}
+  \text{Child UCT value} = \frac{Q_{(s',a')}}{N_{(s',a')}} + c_{uct} * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
+\end{equation}
+
+\subsubsection*{Selection}
+During the selection phase a node $(s,a)$ within the tree which has not yet been expanded is found.
+This process uses Upper Confidence Bound on Trees (UCT) to find that node $(s,a)$, the formula is described in Equation \ref{UCT_formula}. For a parent node $(s,a)$
+(initially the root of the tree) we select the child with the highest UCT value. Repeatedly until an unexpanded
+node is found. This process is done to balance the amount of exploration vs exploitation of nodes in the
+tree.
+
+\subsubsection*{Expansion}
+Then the expansion phase expands the node generating all of $(s,a)$'s children $(s',a')$ are generated
+and connected to $(s,a)$. This is done by applying all actions $a'$ to $(s,a)$. These children are the
+product of applying each action $a'$ to $s$ in the parent node.
+
+\subsubsection*{Rollout}
+Next during rollout actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated
+to go to $(s'',a'')$, until a terminal node within the environment is reached. By terminal we mean a
+state in which the game is finished. A terminal node in MCTS can generally return any value, but in the
+context of this paper we only return (+1 white wins, -1 black wins, 0 draw).
+
+\subsubsection*{Backpropagation}
+The result from the terminal node is then propagated up through the path taken by selection $(s,a)$ up
+to the root of the tree, updating the $Q_{(s,a)}$ values of each node $(s,a)$.
+
+When training a neural network the UCT formula is modified slightly to prefer selecting nodes
+that the neural network values highly by introducing a second scalar to the formula $f(s,a) = (p,v)$,
+the resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT. Secondly, the backpropagation process
+is modified to instead of doing rollout/simulation to recieve a reward the predicted value $v$ from
+the neural network is used instead.
+
+\begin{equation} \label{PUCT_formula}
+  \text{Child PUCT value} = \frac{Q_{(s',a')}}{N_{(s',a')}} + c_{uct} * P((s,a)) * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
+\end{equation}

### 2020-03-24 latex/sections/methods.tex
updating skeleton

@@
 \chapter{Methods}
+
+To train neural networks to play breakthrough we employ a process of self play, where two neural networks are first
+randomly initialized, then they play against each other. They play for $1000$ games, and the neural network that wins
+lets call that one more games is called $m_1$ and the other $M_2$. $M_1$'s weights are then copied to $M_2$. Then $M_1$,
+plays against itself for $20$ iterations recording the outcome of the game. The outcome is compiled to a dataset containing
+four values ($b$, $p$, $v$, $r$). $b$ stands for a gamestate within it's simulation while playing against itself, $p$ stands
+for the policy returned by the neural network by
+
+\section{Monte Carlo Tree Search}
 

### 2020-07-11 latex/sections/background.tex
wrote to background and intro

@@
-\section{Background}
+\chapter{Background\label{cha:background}}
 
-\subsection{Explainable Artificial Intelligence (XAI)}
+In this chapter we discuss the algorithms we decided to use, what some other similar 
+methods exsist as well as the field in general.
 
-The field of XAI research is still very far behind it's counterpart AI research,
-within XAI two fields are concidered the largest that is Model Interpretability
-and Model Explainability.
+\section{Environments}
+\label{sec:environments}
 
-\subsubsection{Model Interpretability}
+When doing research on Artificial Intelligence it is important to select 
+an environment that is a suitable abstraction for the task at hand. Environments vary 
+significantly, and can be identified by their characteristic. The characteristics 
+that are generally used to describe environments can be seen in Table \ref{tab:env_characteristics}. 
+CITE AIMABOOK
 
-\subsubsubsection{Saliency Maps}
-Within XAI many methods have been developed to try to evaluate
-ANN's, these methods are often referred as model interpretability methods.
-In the field image recognition there has been a lot of work examining which
-pixels of an image the model deems important. Arguably the most popular
-method for this is Saliency Maps (CITE ME). There the pixel values the model
-deems important are colored in s.t. a human can examine the image and get a
-sense of what portions of the image are important to the model, an example of
-a classification of a dog can be seen in Figure \ref{fig:dog_saliency}.
+\begin{table}[h]
+  \centering
+  \begin{tabular}{|c|c|p{6cm}|}
+    \hline
+    \textbf{characteristic} & \textbf{Values} & \textbf{Description}\\
+    \hline
+    Observable & Fully, Partially & How much of the environment can your agent percieve.\\
+    \hline
+    Agents & Single, Multi & Are there multiple agents playing in the environment.\\
+    \hline
+    Deterministic & Deterministic, Stochastic & Do the actions your agent do deterministicly impact the environment\\
+    \hline
+    Episodic & Sequential, Episodic & Are actions episodic or sequential\\
+    \hline
+    Static & Static, Semi-Static, Dynamic & ehh  \\
+    \hline
+    Discrete & Discrete, Continious & Is your environment discrete w.r.t actions (does it end) \\
+    \hline
+  \end{tabular}
+  \caption{Charcteristics of environments}
+  \label{tab:env_characteristics}
+\end{table}
+
+Categorizing environments like this gives you the power to find an environment in which 
+a method works and know it can be applied to different environments with the same characteristics.
+Talk about \textbf{agents} in the context of environments as the entities that act within the 
+environment. A game player is an agent as well as the neural network model that exists within 
+a car that is able to drive in the real world.
+
+\section{Game Environment}
+
+With classical Artificial Intelligence Game Environments are commonly used to validate a method, 
+game environments can be games like Tic-Tac-Toe, Breakthrough, and driving simulators. Game environments 
+are a suitable place to apply AI as they serve an important function as an abstraction of the real 
+world, for instance, a self-driving car agent who is verified to avoid driving into walls in a 
+simulation is possibly safer than one who is not.
+
+\section{Breakthrough}
+
+The game breakthrough is a simplified version of chess, the game is set up on an $MxN$ board with 
+cells like in chess, and each player starts with two rows of pawns at either end. The objective of the 
+game is to move one of the pawn pieces to the opposite end of the board. A player wins if 
+either they have reached the opposite end of the board, or has captured all of his opponents pawns.
+The pawns differ from chess pawns in such a way that they can not move two squares on the first move and,
+they can move diagonally as well as forward, this leads to the game being impossible to draw as 
+pieces are always able to move. An example of an initial board in breakthrough can be seen in Figure \ref{fig:initbtboard}
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{graphics/dog_saliency}
-  \caption{Example of a models saliency map for an image of a dog}
-  \label{fig:dog_saliency}
+
+  \breakthrough{8/8/pppppp2/pppppp2/8/8/PPPPPP2/PPPPPP2 w - - 0 1}
+
+  \caption{Example breakthrough board}
+  \label{fig:initbtboard}
 \end{figure}
 
-While this method is understandable in the context of image recognition it
-lacks severely when your input is not an image.
+\begin{table}[h]
+  \centering
+  \begin{tabular}{|c|c|}\hline
+    \textbf{Characteristic} & \textbf{Value}\\\hline
+    Observable & Fully \\
+    Agents & Multi \\
+    Deterministic & Deterministic\\
+    Episodic & Sequential \\
+    Static & Static \\
+    Discrete & Discrete \\\hline
+  \end{tabular}
+  \caption{Categorization of Breakthrough}
+  \label{tab:breakthrough_cat}
+\end{table}
 
-\subsubsubsection{Shapley Values}
+Categorizing Breakthrough with the characteristics described in Section \ref{sec:environments}. We end up with the description 
+shown in Table \ref{tab:breakthrough_cat}. This description is identical to that of Tic-Tac-Toe, and Chess.
+This description is the most common in board games where two player compete.
 
-Methods for explaining models that aren't image recognition models include
-shapley values. There the input is examined against it's output, then iteratively
-input values are selected to be fixed. Then the other input values are varied and
-an average change in prediction is calculated. With this the shapley value can be
-estimated for the fixed input value. This is done to examine which input values have
-the strongest link to the output value. Shapley values on a dataset can give insights
-on which input values the model deems important.
+\subsection{Heuristics of Breakthrough}
 
-\subsubsubsection{Concept Activation Vectors}
+To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance 
+a very simple heuristic would be the amount of pieces each player has left. This heuristics gives us some 
+insight into how well the game is progressing, but obviously there are cases where this doesn't tell us 
+much, as in cases when your opponent has a single piece left that is on the row immediately before the 
+row needed for him to win. No matter how many pieces you have left, this state is bad for you if you're not 
+able to capture that piece. An example of such a state can be seen in Figure \ref{fig:bt_h1_bad}.
 
-A recent paper by Been Kim Et. al (CITATION NEEDED), shows a method for examining
-a neural network giving a much more human insight into a prediction. Using Concept
-Activation Vectors (CAV) a directional derivative for a given input can be examined
-with respect to some HLC's. For example, when a human looks at an image of an animal
-and is supposed to decide whether the image is of a horse or a zebra, an intuitive
-approach would be to check whether the animal has stipes, or is both white and black.
+\begin{figure}[]
+  \centering
+  \breakthrough{8/8/pppp4/pp3P2/2p5/8/8/PPP5 w - - 0 1}
+  \caption{Breakthrough board where Heuristic 1 doesn't work well}
+  \label{fig:bt_h1_bad}
+\end{figure}
 
-CAV's in neural networks are just binary classifiers of the same dataset used to train
-a neural network. The
+A different heuristic would be the distance of your most advanced pawn minus your opponents most 
+advanced pawn, this heuristic could give you insight into how close you are to winning the game or 
+how close your opponent is. As a higher level concept we can call this concept your aggressiveness, 
+as it closely resembles how aggressive you are for going in for the win. Generally in Breakthrough 
+it is favourable to move your whole team as a unit and play more defensively, so this heuristic 
+is probably not one that will lead to the most wins.
 
-\subsubsection{Model Explainability}
+\section{State-Space Search}
 
-Model explainability within the context of neural networks isn't possible today. Model
-explainability referrs to firstly considering some input and output from a model. Then
-afterwards the model is examined to determine exactly what led to the predicted output.
-This concept is simple when we're working with Decision Trees. A decision tree is a tree
-whose nodes are representative of an input value and at every node a branch is selected
-based on the value of the input value. It is therefore easy to see how to examine the tree
-to explain the output. We just follow the branches in the tree. That being said the branches
-are created by algorithms like ID3 which construct branches based on the initial dataset used
-to construct the tree, but again ID3 follows simple statistics and can be explained properly.
+Traditionally methods for playing games were searching through the environment using a heuristic to
+guide the search. A simple way of doing a heuristic based search would be to give all states a $0$ 
+as a score and give states where you win a positive score. We say that those search algorithm 
+is not guided, and the algorithm will probably have to evaluate every state in the state-space. 
+This method of searching is generally extremenly inefficient as the state-spaces of game environments 
+are generally extremely large, and often infinite. For instance an estimate of the state-space 
+for the game breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width 
+of the board. The $2^{N*4}$ represent each piece either being alive or captured. So for a small board
+$5x4$ the estimated state-space is $1310720$ many states.
 
-When we talk about neural networks this process is much more difficult, the underlying nodes
-are generally in the thousands, the different layers of the neural network varies in the operations
-it applies to the input value and such while travelling through the neural network the modified
-value becomes far removed from the initial input value to the eyes of the reader. That being said,
-while the possibility of completely monitoring the training process and completely monitoring the
-evaluation process is truly possible it is not feasible. And secondly the process of seeing an
-input and it's corresponding output will not be of any value if one were to consider the process
-of prediction.
+This is why a good heuristic is very valuable because we can disregard all states $S'$ that 
+result from doing a move $a$ in state $S$ as they will only lead to worse outcomes.
 
-\subsection{Machine Learning}
+The algorithms that are used in traditional state-space search are for instance Depth-first search (DFS),
+Breadth-first Search (BFS), Alpha* Search (A-Star), and Monte-Carlo Tree Search (MCTS).
 
-\subsubsection{Reinforcement Learning}
+More modernly these search methods have been amplified by Machine Learning, in such a way that 
+we do not need to figure out a good heuristic for a given state, but rather, we apply a machine 
+learning model to learn a function that takes in a state and returns an evaluation of that state.
+This can lead to a significant time reduction as we do not need to simulate a whole game from a 
+state to recieve its evaluation we rather receive the evaluation from the model.
 
-Reinforcement Learning is a subfield of Machine Learning where a model learns from experience.
-That is, the notion of input/output values changes, the model uses itself to generate input
-values by acting in an environment. The environment then returns some result, that could be
-losing in a game, making a correct prediction, or any number of outcomes. The result is then
-propagated through the model moving it's parameters in the direction of the recieved output
-from the environment.
+\subsection{Monte-Carlo Tree Search}
 
-Many algorithms are popular in reinforcement learning, for instance Q-learning(CITE ME), CARLA (CITE ME),
-Monte Carlo Tree Search, and many others. The work in this paper relies heavily on Monte Carlo Tree Search,
-and will be the only one thoroughly discussed.
+\label{sec:mcts}
 
-\subsubsection{Monte Carlo Tree Search}
-
-In the algorithm Monte Carlo Tree Search there is an agent exists
+In the algorithm Monte-Carlo Tree Search there is an agent exists
 within some environment. Where each node in the environment represents a state-action pair of the environment, by
@@
 
-Monte Carlo Tree Search is a method of exploring an enviroment in a randomized manner (Monte Carlo is the term implying
+MCTS is a method of exploring an enviroment in a randomized manner (Monte Carlo is the term implying
 randomness). In MCTS there are four stages. Selection, Expansion, Simulation, and Backpropagation. They happen sequentially
@@
 
-\subsubsection*{Selection}
+\subsection*{Selection}
 During the selection phase a node $(s,a)$ within the tree which has not yet been expanded is found.
@@
 
-\subsubsection*{Expansion}
+\subsection*{Expansion}
 Then the expansion phase expands the node generating all of $(s,a)$'s children $(s',a')$ are generated
@@
 
-\subsubsection*{Rollout}
+\subsection*{Rollout}
 Next during rollout actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated
@@
 
-\subsubsection*{Backpropagation}
+\subsection*{Backpropagation}
 The result from the terminal node is then propagated up through the path taken by selection $(s,a)$ up
@@
 \end{equation}
+
+\section{Machine Learning}
+
+Machine learning (ML) focuses on the using of data and a corresponding outcome w.r.t that data to 
+extrapolate some underlying function of that data. This some task like predicting the 
+rise and fall of some stock, what the weather will be in a week, and whether an image 
+is an image of a dog or a cat. Many datastructures and algorithms are used to achieve this 
+goal, but recently the field of ANN/DNN's has been the standard for achieveing the best
+results.
+
+\subsection{Reinforcement Learning}
+
+Reinforcement Learning is a subfield of Machine Learning where a model learns from experience.
+That is, the notion of input/output values changes, the model uses itself to generate input
+values by acting in an environment. The environment then returns some result, that could be
+losing in a game, making a correct prediction, or any number of outcomes. The result is then
+propagated through the model allowing it to improve with this new information.
+% When doing moving it's parameters in the direction of the recieved output
+% from the environment.
+
+Many algorithms are popular in reinforcement learning, for instance Q-learning(CITE ME), CARLA (CITE ME), 
+and many others.
+
+\subsection{Neural Networks}
+
+Neural networks (NN) are popular methods within a subfield of ML which is called Deep Learning (DL). 
+NN's are created to resemble how the human brain functions. In the brain we have neurons which 
+when they get a signal they apply some function to them and if the resulting signal is high 
+enough, they fire to the next neuron. This is how it is done in the neural network model as well.
+There we have neurons that when they get some input, generally a vector of numbers. The neuron 
+takes the sum of that vector, weights the sum by a constant, then applies a non-linear activation 
+function to it. The result of doing this is then passed on to the next neuron. Until a final layer 
+of neurons is reached. At that point we have a value that the neural network corresponds to the 
+input value. This value can be a binary classification (cat or dog image), a regression value (the 
+value of a property), or any number of outputs. It can then be said that a neural network is doing 
+a function approximation of an input to some value. $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.
+
+Neural networks are machine learning models that take in as input anything that is numerical 
+and they apply differentiatable function to that input s.t. they end with some output. This 
+output is then compared to the expected output the difference between these outputs is called 
+a loss. A loss is backpropagated through the neural network, and each function is derived in 
+order to find its slope. We can then modify the weights of the neural network in the direction 
+of the correct output. Leading to a function approximation of this function
+$f_{neuralnetwork}(X_{input}) = O_{expectedoutput}$
+
+\section{Explainable Artificial Intelligence (XAI)}
+
+The field of XAI research is still very far behind it's counterpart AI research,
+within XAI two fields are concidered the largest that is Model Interpretability
+and Model Explainability.
+
+\subsection{Model Interpretability}
+
+\subsubsection{Saliency Maps}
+Within XAI many methods have been developed to try to evaluate
+ANN's, these methods are often referred as model interpretability methods.
+In the field image recognition there has been a lot of work examining which
+pixels of an image the model deems important. Arguably the most popular
+method for this is Saliency Maps (CITE ME). There the pixel values the model
+deems important are colored in s.t. a human can examine the image and get a
+sense of what portions of the image are important to the model, an example of
+a classification of a dog can be seen in Figure \ref{fig:dog_saliency}.
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=.5\textwidth]{graphics/dog_saliency}
+  \caption{Example of a models saliency map for an image of a dog}
+  \label{fig:dog_saliency}
+\end{figure}
+
+While this method is understandable in the context of image recognition it
+lacks severely when your input is not an image.
+
+\subsubsection{Shapley Values}
+
+Methods for explaining models that aren't image recognition models include
+shapley values. There the input is examined against it's output, then iteratively
+input values are selected to be fixed. Then the other input values are varied and
+an average change in prediction is calculated. With this the shapley value can be
+estimated for the fixed input value. This is done to examine which input values have
+the strongest link to the output value. Shapley values on a dataset can give insights
+on which input values the model deems important.
+
+\subsubsection{Concept Activation Vectors}
+
+A recent paper by Been Kim Et. al (CITATION NEEDED), shows a method for examining
+a neural network giving a much more human insight into a prediction. Using Concept
+Activation Vectors (CAV) a directional derivative for a given input can be examined
+with respect to some HLC's. For example, when a human looks at an image of an animal
+and is supposed to decide whether the image is of a horse or a zebra, an intuitive
+approach would be to check whether the animal has stipes, or is both white and black.
+That method of determining if a horse is a zebra could then be called a higher-level
+concept, and if we're able to gather if a nerual network uses this strategy for prediction 
+we have a deeper understanding of its underlying structure. Leading to an explanation of 
+the result.
+
+CAV's in neural networks are just binary classifiers of the same dataset used to train
+a neural network. The
+
+\subsection{Model Explainability}
+
+Model explainability within the context of neural networks isn't possible today. Model
+explainability referrs to firstly considering some input and output from a model. Then
+afterwards the model is examined to determine exactly what led to the predicted output.
+This concept is simple when we're working with Decision Trees. A decision tree is a tree
+whose nodes are representative of an input value and at every node a branch is selected
+based on the value of the input value. It is therefore easy to see how to examine the tree
+to explain the output. We just follow the branches in the tree. That being said the branches
+are created by algorithms like ID3 which construct branches based on the initial dataset used
+to construct the tree, but again ID3 follows simple statistics and can be explained properly.
+
+When we talk about neural networks this process is much more difficult, the underlying nodes
+are generally in the thousands, the different layers of the neural network varies in the operations
+it applies to the input value and such while travelling through the neural network the modified
+value becomes far removed from the initial input value to the eyes of the reader. That being said,
+while the possibility of completely monitoring the training process and completely monitoring the
+evaluation process is truly possible it is not feasible. And secondly the process of seeing an
+input and it's corresponding output will not be of any value if one were to consider the process
+of prediction.

### 2020-07-11 latex/sections/introduction.tex
wrote to background and intro

@@
 
-In the world of deep learning is exciting, we have models that are able to examine
+The world of deep learning (DL) is exciting, we have models that are able to examine
 images and very reliably be able to identify it's content. Deep learning models have
@@
 
+DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search, 
+was used to defeat the standing world champion NAME in the incredibly expansive game of Go. This was done in the year 
+2017, the researchers used reinforcement learning (RL), where the agent purely played against itself. 
+These result imply that given enough time to learn the computer agent is able to
+gain a deep insight on how the game should be played.
+
 The field of examining deep learning models to gain a richer understanding
-in how they work, and what makes them so much better than humans at various tasks
+of how they work, and what makes them so much better than humans at various tasks
 is still new. This is the field of Explainable Artificial Intelligence (XAI).
@@
 to improve our lives we must examine how they work, this is not only to improve
-ourselves but these explanations are a legal requirement now in many areas, namely, legal, and
+the models themselves but could also augment our ability in many cases.
+Furthermore, these explanations are a legal requirement now in many areas, namely, legal, and
 medical. Recently, XAI has generally been focused on Model Interpretability, in
@@
 within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
-complicated like Chess or Go. Some of the research questions are what if an ANN values a HLC greatly
-which we as humans don't find particularily important, and yet the ANN is able to defeat
-us what does that mean?
+complicated like Chess or Go.
 
@@
 current state of the game by evaluating them with respect to a heuristic. A heuristic within
-games are a evaluation of the end cost for a state given just the state, for example,
+games are an evaluation of the end cost for a state given just the state, for example,
 the amount of pieces left within a game of chess. Intuitively, the amount of pieces
@@
 
+We define the research questions
+
+\begin{enumerate}
+  \item Given a NN that plays Breakthrough very well, can we evaluate if that NN recognizes the 
+  HLCs that human players use.
+  \item Does a NN that trains itself using self-play learn these HLCs over time, and does it start to emphasize 
+  the HLCs that are generally taught later to human players, more as it trains itself more.
+  \item Does a NN that plays better than some humans focus on HLCs that are not generally 
+  considered by those human players and it.
+\end{enumerate}
+
+\subsection{Summary of the thesis}
+
 In this thesis we take an example game of Breakthrough, train a neural network to play the game
@@
 to play the game.
-
-The neural network architecture is based of the popular AlphaZero neural network
-used to play Chess, Shogi, and Go on a superhuman level. (REMOVE ME) Unfortunately due to lack
-of thousands of TPU's this research can't go as far as to examine those more complicated
-games. (REMOVE ME)
 

### 2020-07-11 latex/sections/methods.tex
wrote to background and intro

--- 2019-11-28 latex/sections/methods.tex
@@
-\chapter{Methods}
+\chapter{Methods\label{cha:methods}}
 
-%\section{Machine learning methods}
+In this chapter we train a neural network to play the game breakthrough described in 
+Chapter \ref{cha:background}. The training process uses MCTS described in Section \ref{sec:mcts} 
+to guide it's training.
 
-This section discusses the various machine learning methods utilized throughout this project 
-and discusses their applicability.
+The training algorithm is a reinforcement learning based self-play algorithm based on the work 
+done in AlphaZero (CITE ME).
 
-%\subsection{Deep neural network}
+\subsection{Neural network Architecture}
 
-I'm a little teapot
+The neural network architecture we opted to use was a single convolutional layer with a ReLU 
+activation function, followed by 5 residual layers, and lastly a split policy/value head.
+A figure describing the architecture can be seen in Figure \ref{fig:nnarch}.
 
-%\subsection{Data joined network}
+\begin{figure}[]
+    \centering
 
-I'm a smaller teapot
+    \includegraphics[width=0.65\textwidth]{graphics/test}
+
+    \caption{Neural Network Architecture}
+    \label{fig:nnarch}
+\end{figure}
+
+\subsection{Self-play}
+
+To train the neural network to play Breakthrough we initialize two neural networks $N_1$ and $N_2$ 
+with random weights. Then we let $N_1$ play against itself using MCTS to direct its training. 
+While playing against itself the neural network gathers data for it to train with.
+
+The data collected is $(\pi, \tau, p, v)$, where $\pi$ is the action probabilities provided by 
+MCTS for $N$ iterations, $\tau$ is the reward from playing the game, $p$ is the policy vector
+from the neural network, and $v$ is the predicted reward of the game by the neural network.
+
+\subsection{Loss Function}
+
+The data collected is backpropagated through the NN moving the weights to the direction of this 
+loss function $l = p * \pi + (\tau - v)^2$

### 2020-07-17 hlc_evaluator/verifier.py
verifier stuff

@@ -20,7 +20,7 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
   neural_network_1 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
   neural_network_2 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
 
-  test_as_white = True
+  test_as_white = False
 
   if test_as_white:
     neural_network_1.loadmodel(first_network_path, first_network_name)
@@ -30,7 +30,7 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
   memo_nn1 = {}
   memo_nn2 = {}
 
-  NN_THINK = 50
+  NN_THINK = 1
 
   initial_node = Node(state_example.initial_state(), "START")
   first_win = 0
@@ -47,12 +47,20 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
 
         # WHITE MOVES
         pi = monte_tree_1.get_policy(curr_node, NN_THINK, neural_network_1)
+        
+        if not curr_node.is_expanded():
+          curr_node.expand()
+
+        pi,v = neural_network_1.safe_predict(curr_node.gamestate)
+        pi = pi.view(-1)
 
         for i,child in enumerate(curr_node.children):
           if not child:
             pi[i] = 0
-        pi = pi / sum(pi)
+        pi = pi / sum(pi).item()
 
+        print("policy:",pi)
+        print(sum(pi))
         curr_node = np.random.choice(curr_node.children, p=pi)
 
         if curr_node.gamestate.is_terminal():
@@ -62,11 +70,17 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
 
         # BLACK MOVES
         pi = monte_tree_2.get_policy(curr_node, NN_THINK, neural_network_2)
+        if not curr_node.is_expanded():
+          curr_node.expand()
+        pi,v = neural_network_2.safe_predict(curr_node.gamestate)
+        pi = pi.view(-1)
 
         for i,child in enumerate(curr_node.children):
           if not child:
             pi[i] = 0
-        pi = pi / sum(pi)
+        pi = pi / sum(pi).item()
+        print("policy:",pi)
+        print(sum(pi).item())
 
         curr_node = np.random.choice(curr_node.children, p=pi)
 
@@ -87,4 +101,4 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
[…]

### 2020-07-17 latex/sections/managerial.tex

@@
 AI & Artificial Intelligence\\
+ANN & Artificial Neural Network\\
+DNN & Deep Neural Network\\
+MCTS & Monte-Carlo Tree Search\\
+CAV & Concept Activation Vector\\
+
 \end{tabular}

### 2020-07-17 latex/sections/methods.tex
starting some methods writing

@@
 
+\subsection{Policy head}
+
+The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move in the space (diagonally both directions, forward and backwards).
+The vector is then masked s.t.
+the values representing moves that can not be played on the board are given the value of $0$.
+
 \subsection{Self-play}
@@
 
-The data collected is backpropagated through the NN moving the weights to the direction of this 
+The data collected is back-propagated through the NN moving the weights to the direction of this 
 loss function $l = p * \pi + (\tau - v)^2$

### 2020-07-17 latex/sections/background.tex

@@
 In this chapter we discuss the algorithms we decided to use, what some other similar 
-methods exsist as well as the field in general.
+methods exist as well as the field in general.
 
@@
     \hline
-    Discrete & Discrete, Continious & Is your environment discrete w.r.t actions (does it end) \\
+    Discrete & Discrete, Continuous & Is your environment discrete w.r.t actions (does it end) \\
     \hline
   \end{tabular}
-  \caption{Charcteristics of environments}
+  \caption{Characteristics of environments}
   \label{tab:env_characteristics}
@@
 as it closely resembles how aggressive you are for going in for the win. Generally in Breakthrough 
-it is favourable to move your whole team as a unit and play more defensively, so this heuristic 
+it is favorable to move your whole team as a unit and play more defensively, so this heuristic 
 is probably not one that will lead to the most wins.
@@
 is not guided, and the algorithm will probably have to evaluate every state in the state-space. 
-This method of searching is generally extremenly inefficient as the state-spaces of game environments 
+This method of searching is generally extremely inefficient as the state-spaces of game environments 
 are generally extremely large, and often infinite. For instance an estimate of the state-space 
@@
 This can lead to a significant time reduction as we do not need to simulate a whole game from a 
-state to recieve its evaluation we rather receive the evaluation from the model.
+state to receive its evaluation we rather receive the evaluation from the model.
 
@@
 
-MCTS is a method of exploring an enviroment in a randomized manner (Monte Carlo is the term implying
-randomness). In MCTS there are four stages. Selection, Expansion, Simulation, and Backpropagation. They happen sequentially
+MCTS is a method of exploring an environment in a randomized manner (Monte Carlo is the term implying
+randomness). In MCTS there are four stages. Selection, Expansion, Simulation, and Back-propagation. They happen sequentially
 and repeatedly. MCTS is initialized with a tree consisting of the unexpanded initial state of the environment.
@@
 
-Machine learning (ML) focuses on the using of data and a corresponding outcome w.r.t that data to 
-extrapolate some underlying function of that data. This some task like predicting the 
-rise and fall of some stock, what the weather will be in a week, and whether an image 
-is an image of a dog or a cat. Many datastructures and algorithms are used to achieve this 
-goal, but recently the field of ANN/DNN's has been the standard for achieveing the best
-results.
+Machine Learning (ML) is a research field in which machines apply statistical functions on data to achieve the output that is \textit{correct}, by correct we mean the corresponding result which we expect.
+Generally this a repetitive process where we look at examples of the data, and the algorithm progressively gets closer to the underlying function of the data it is fed.
+This process is therefore similar to trial and error for humans.
+ML is a sub-field of Artificial Intelligence.
+ML algorithms come in the form of two classes \textit{Classification}, where the algorithm should find a class representing the data its given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.
+
+Typically ML can be viewed in three different groups, Supervised Learning, Reinforcement Learning (RL), and Unsupervised Learning.
+Where in Supervised Learning, the algorithm is given data examples and their corresponding outcome.
+For example, a supervised learning algorithm could be provided with data regarding the weather and the corresponding temperature, the algorithm should then find a pattern within the weather data and find the continuous function represented by the data.
+This would then be an example of a regression task.
+Flipping thing example around, if the algorithm would just be provided the temperature and it should tell us whether it is sunny outside or not, that would be a classification task.
+
+In RL the algorithm is given only some input, and then the algorithm tries some outcome generally actions in some environment.
+Then over an episode~\footnote{a series of outcomes / a timespan} once the episode is finished some reward is given.
+The algorithm will then learn whether the actions were good actions from the reward.
+Examples of this are agents playing a game like Flappy Bird, where the data they are given is the state of the game, and they try to either jump or not jump.
+
+%Machine learning (ML) focuses on the using of data and a corresponding outcome w.r.t that data to 
+%extrapolate some underlying function of that data. Examples of how we use machine learning is for instance the ability to predicting the 
+%rise and fall of some stock, predicting what the weather will be in a week, and whether an image 
+%is an image of a dog or a cat. Many data structures and algorithms are used to achieve this 
+%goal, but recently the field of ANN/DNN's has been the standard for achieving the best
+%results.
+
+%As a general notion we can split machine learning into three different sects those are Supervised Learning, Reinforcement Learning, and Unsupervised Learning.
+
+\subsection{Supervised Learning}
+
+In supervised learning, the data and their corresponding outcomes are already available to us. Examples of the data we can use in supervised learning would be labeled images.
 
@@
 
-Reinforcement Learning is a subfield of Machine Learning where a model learns from experience.
-That is, the notion of input/output values changes, the model uses itself to generate input
-values by acting in an environment. The environment then returns some result, that could be
-losing in a game, making a correct prediction, or any number of outcomes. The result is then
-propagated through the model allowing it to improve with this new information.
+Reinforcement Learning is where a model learns from experience.
+That is, the notion of input/output values changes, the model uses itself to generate input values by acting in an environment.
+The environment then returns some result, that could be losing in a game, making a correct prediction, or any number of outcomes.
+The result is then propagated through the model allowing it to improve with this new information.
 % When doing moving it's parameters in the direction of the recieved output
@@
 
-Neural networks (NN) are popular methods within a subfield of ML which is called Deep Learning (DL). 
+Neural networks (NN) are popular methods within a sub-field of ML which is called Deep Learning (DL). 
 NN's are created to resemble how the human brain functions. In the brain we have neurons which 

### 2020-07-17 latex/sections/methods.tex

@@
 The data collected is $(\pi, \tau, p, v)$, where $\pi$ is the action probabilities provided by 
-MCTS for $N$ iterations, $\tau$ is the reward from playing the game, $p$ is the policy vector
+MCTS for $N$ iterations, $\tau$ is the reward for the whole episode, $1$ if white wins,
+$-1$ if black wins, and $0$ on a tied game, $p$ is the policy vector
 from the neural network, and $v$ is the predicted reward of the game by the neural network.
@@
 
-The data collected is back-propagated through the NN moving the weights to the direction of this 
-loss function $l = p * \pi + (\tau - v)^2$
+The data collected is backpropagated through the NN moving the weights to the direction of this 
+loss function $l = (p * \pi) + (\tau - v)^2$ for each state the NN encountered during self-play.

### 2020-07-17 latex/sections/introduction.tex
latex work

@@
 DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search, 
-was used to defeat the standing world champion NAME in the incredibly expansive game of Go. This was done in the year 
-2017, the researchers used reinforcement learning (RL), where the agent purely played against itself. 
+was used to defeat the standing world champion Lee Sedol in the incredibly expansive game of Go. This was done in the year 
+2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself. 
 These result imply that given enough time to learn the computer agent is able to
@@
 The method of examining HLC's within games has generally been done by examining the
-current state of the game by evaluating them with respect to a heuristic. A heuristic within
-games are an evaluation of the end cost for a state given just the state, for example,
+current state of the game by evaluating them with respect to some heuristic. A heuristic within
+games are evaluations of the end reward for a state given only the state, for example,
 the amount of pieces left within a game of chess. Intuitively, the amount of pieces
@@
 where the king is safe from attack, or the structure of the pawn positions.
-The idea is to examine a neural networks evaluation of a state regarding
+This paper examines a neural networks evaluation of a state regarding
 those higher-level concepts, if human players evaluate the king safety of a state
@@
 to play the game.
-
-% The chess engine DeepBlue was the first chess engine to win a chess match agains a
-% chess grand master. It's power in the world of chess was a marvel. DeepBlue used tree-
-% search to evaluate it's position and possible moves, doing an immense amount of calcluations
-% in order to arrive at the best move it could given the time constraints. The tree search
-% approach is dependent on knowing which board state is good and what board state is bad,
-% This can be easy when the boardstate is such that you're guaranteed to have been checkmated
-% in one move, and it's incredibly hard when you boardstate is such that both you and your
-% opponent have 16 pieces and are only a handfull of moves in. Generally the boardstate
-% is a function of the pieces you have, how safe they are, and how opportunistic you are to
-% capture your opponents pieces. These values are not obvious and require a domain expert
-% to describe the situations and give the values for these cases.
-
-% The process of evaluating chess positions did not change for many years. With the revival
-% of deep neural networks this changed. A neural network is a network of connected nodes (neurons)
-% that each how a specific weight. These networks are fed input which is passed through the
-% network and it is updated by the weights of the neurons, the input is morphed into
-% a more comprehensible value and these values are then used to generate some output.
-% The output can be a binary classification for instance input=Image output=\{cat, dog\}.
-% Processing the pixels and altering them to numerical values to find the output cat or dog
-% is not intuitive and understanding of the values of the neurons in the network
-% is not considered fruitful.
-
-% This approach of
-% chess engines is still popular today but not the reigning champion anymore. In todays
-% best chess engines deep neural networks are used to
-
-% Deep neural network machine learning models have an amazing impact on modern society today,
-% we use them to know which items to purchase, to know which people we should love, and we
-% use them to drive us to work. These tasks given to deep learning models are all niceties
-% for us to enjoy. However, as our dependance on them increases and their astounding success
-% in any and all tasks continues, we will want to utilize them for all things. This leads
-% to a utopia where decision making is removed from the hands of humans. The question we
-% must ask ourselves is when do we want to know why a decision was taken, is it only when
-% the decision taken is not preferrable to us. For instance, if a deep neural network model
-% used by your lending authority decides you are not worty of a loan. These questions should
-% have answers but as it stands today no such answer is possible.
-
-% We can use methods like saliency maps to map a deep learning model emphasis on pixels to
-% hopefully understand what portions the model is focusing on. This method however, tells us
-% nothing about what
-
-% results that are nothing but amazing.

### 2020-07-17 latex/sections/methods.tex
latex work

@@
 
-\subsection{Policy head}
-
-The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move in the space (diagonally both directions, forward and backwards).
-The vector is then masked s.t.
-the values representing moves that can not be played on the board are given the value of $0$.
-
 \subsection{Self-play}
@@
 \subsection{Loss Function}
-
 The data collected is backpropagated through the NN moving the weights to the direction of this 

### 2020-07-20 hlc_evaluator/static/js/jquery.js
adding web frontend

@@ -1,0 +1,2 @@
+/*! jQuery v3.4.1 | (c) JS Foundation and other contributors | jquery.org/license */
+!function(e,t){"use strict";"object"==typeof module&&"object"==typeof module.exports?module.exports=e.document?t(e,!0):function(e){if(!e.document)throw new Error("jQuery requires a window with a document");return t(e)}:t(e)}("undefined"!=typeof window?window:this,function(C,e){"use strict";var t=[],E=C.document,r=Object.getPrototypeOf,s=t.slice,g=t.concat,u=t.push,i=t.indexOf,n={},o=n.toString,v=n.hasOwnProperty,a=v.toString,l=a.call(Object),y={},m=function(e){return"function"==typeof e&&"number"!=typeof e.nodeType},x=function(e){return null!=e&&e===e.window},c={type:!0,src:!0,nonce:!0,noModule:!0};function b(e,t,n){var r,i,o=(n=n||E).createElement("script");if(o.text=e,t)for(r in c)(i=t[r]||t.getAttribute&&t.getAttribute(r))&&o.setAttribute(r,i);n.head.appendChild(o).parentNode.removeChild(o)}function w(e){return null==e?e+"":"object"==typeof e||"function"==typeof e?n[o.call(e)]||"object":typeof e}var f="3.4.1",k=function(e,t){return new k.fn.init(e,t)},p=/^[\s\uFEFF\xA0]+|[\s\uFEFF\xA0]+$/g;function d(e){var t=!!e&&"length"in e&&e.length,n=w(e);return!m(e)&&!x(e)&&("array"===n||0===t||"number"==typeof t&&0<t&&t-1 in e)}k.fn=k.prototype={jquery:f,constructor:k,length:0,toArray:function(){return s.call(this)},get:function(e){return null==e?s.call(this):e<0?this[e+this.length]:this[e]},pushStack:function(e){var t=k.merge(this.constructor(),e);return t.prevObject=this,t},each:function(e){return k.each(this,e)},map:function(n){return this.pushStack(k.map(this,function(e,t){return n.call(e,t,e)}))},slice:function(){return this.pushStack(s.apply(this,arguments))},first:function(){return 
[…]

### 2020-08-22 latex/sections/background.tex
latexing

@@
 
-In this chapter we discuss the algorithms we decided to use, what some other similar 
+In this chapter we discuss traditional artificial intelligence in the context of search the algorithms we used, what some other similar 
 methods exist as well as the field in general.
@@
 Categorizing Breakthrough with the characteristics described in Section \ref{sec:environments}. We end up with the description 
-shown in Table \ref{tab:breakthrough_cat}. This description is identical to that of Tic-Tac-Toe, and Chess.
-This description is the most common in board games where two player compete.
+shown in Table \ref{tab:breakthrough_cat}. These characteristics are identical to that of Tic-Tac-Toe, and Chess.
+This categorization is the most common in board games where two player compete.
 
@@
 as it closely resembles how aggressive you are for going in for the win. Generally in Breakthrough 
-it is favorable to move your whole team as a unit and play more defensively, so this heuristic 
+it is favorable to move your whole team as a unit and play more defensively (CITE UNITY HEURISTIC), so this heuristic 
 is probably not one that will lead to the most wins.
@@
 This method of searching is generally extremely inefficient as the state-spaces of game environments 
-are generally extremely large, and often infinite. For instance an estimate of the state-space 
+are often extremely large, and sometimes infinite. For instance an estimate of the state-space 
 for the game breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width 
@@
 The algorithms that are used in traditional state-space search are for instance Depth-first search (DFS),
-Breadth-first Search (BFS), Alpha* Search (A-Star), and Monte-Carlo Tree Search (MCTS).
+Breadth-first Search (BFS), Alpha-Beta Pruning Search (AB-Search), and Monte-Carlo Tree Search (MCTS).
 

### 2020-08-22 latex/sections/introduction.tex
latexing

@@
 DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search, 
-was used to defeat the standing world champion Lee Sedol in the incredibly expansive game of Go. This was done in the year 
+was used to defeat the standing world champion Lee Sedol (CITE ALPHAGO ZERO) in the incredibly expansive game of Go. This was done in the year 
 2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself. 

### 2020-08-22 latex/sections/methods.tex
latexing

--- 2020-07-17 latex/sections/methods.tex
@@
 
-In this chapter we train a neural network to play the game breakthrough described in 
-Chapter \ref{cha:background}. The training process uses MCTS described in Section \ref{sec:mcts} 
-to guide it's training.
+In this chapter we describe our implementation of the game Breakthrough and the methods for training a neural network to play the game Breakthrough, described in Chapter \ref{cha:background}.
+The training process uses MCTS described in Section \ref{sec:mcts} to guide its training.
+We outline the architecture used in the neural network.
 
-The training algorithm is a reinforcement learning based self-play algorithm based on the work 
-done in AlphaZero (CITE ME).
+The training algorithm is a reinforcement learning self-play algorithm based on previous work by DeepMind (CITE ME).
 
-\subsection{Neural network Architecture}
+\section{Implementation}
 
-The neural network architecture we opted to use was a single convolutional layer with a ReLU 
-activation function, followed by 5 residual layers, and lastly a split policy/value head.
-A figure describing the architecture can be seen in Figure \ref{fig:nnarch}.
+In this section we will take a brief look at the implementation of the game Breakthrough, how we model its state representation, and the pseudocode for training.
+
+\subsection{Breakthrough}
+
+In order for a neural network to be able to use input, that input needs to be numeric.
+We model the board in gamethrough with a single $NxM$ array with three values $\{-1,0,1\}$.
+Where each $x,y$ index on the board represents what value is on the board at the time, $-1$ represnts a white piece, $0$ represents an empty square, and $1$ represents a black piece.
+
+A action vector is coupled with each state, the action vector is a $3$-dimensional matrix where the lengths of the dimensions are $row x columns x 6$ and each index of the vector $X,Y,Z$ represents an action moving from cell $x,y$ moving to the direction $z$.
+The directions are as follows $0$ represents moving upwards and to the left diagonally on the board, $1$ upwards, and $2$ upwards and to the right diagonally, $4$ downwards and to the left diagonally, $5$ downwards, and lastly $6$ downwards and to the right diagonally.
+
+\subsection{Selfplay}
+
+\section{Neural Network Architecture}
+
+The neural network architecture we opted to use was a single convolutional layer with a ReLU activation function, followed by 5 residual layers each containing two layers of convolution, batch normalization and a ReLu activation.
+Lastly a split policy/value head, where the output of the last residual layer is split into two different outputs.
+The policy head are two layers: first a final convolutional layer, and last a fully connected layer with a \textit{log softmax} layer.
+The value head, consists of three layers: first a convolutional layer, then a fully connected layer, and lastly a fully connected layer with a \textit{tanh} activation function.
+The Figure \ref{fig:nnarch}.
+depicts the architecture.
 
@@
 
-\subsection{Policy head}
+\subsection{Convolutional Layers}
 
-The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move in the space (diagonally both directions, forward and backwards).
+It is important to understand why we use convolutional layers when dealing with board games as convolution is often associated with image.
+This is because a convolutional layer is focused on merging multiple input parameters to a single neuron, making it an input parameter for the next layer representing the locality around the center point of the original input.
+Playing the game of Breakthrough, a piece in only able to capture pieces in its immediate vicinity.
+This lead to the selection of a convolutional kernel with the size of $3x3$ and a stride of $1$.
+
+The knowledge of each piece moves isn't enough to represent the whole gamestate within the neural network.
+This is why we use multiple residual layers with convolution.
+These residual layers stack convolutions on each other making the final convolution represent the locality of all the other localities, allowing the neural network to have a representation of the whole gamestate in its parameters.
+
+The selection of parameters was done for these reasons as well as to most closely resemble the architecture described in AlphaZero.
+
+\subsection{Residual Layers}
+
+The residual layers as a concept remember the output of the previous layer and add the results of the residual layer and the previous layer together.
+On a higher level this leads to the internal representation of the neural network to maintain the whole game boards in its representation, s.t.
+two areas that are far away from each other maintain the same level of locality as two that are close.
+
+Why we use a residual layer for applying a convolution layer multiple times is to gain locality of the whole game board.
+A residual layer applies a function like this $res(x) = x + l(x)$ where $l(x)$ is the function of the layer, commonly a convolutional layer.
+
+\subsection{Policy Head}
+
+The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move (straight, and diagonally both left and right for the first player, and backwards for the second player).
 The vector is then masked s.t.
 the values representing moves that can not be played on the board are given the value of $0$.
+
+\subsection{Value Head}
+
+The value head of the neural network returns a single value representing the predicted value of the neural network.
+This predicted value is trained to be the end value of the game after taking the predicted move.
 
@@
 MCTS for $N$ iterations, $\tau$ is the reward for the whole episode, $1$ if white wins,
-$-1$ if black wins, and $0$ on a tied game, $p$ is the policy vector
+$-1$ if black wins, $p$ is the policy vector
 from the neural network, and $v$ is the predicted reward of the game by the neural network.
+
+This data collection trains the neural network in such a way that it is a function approximation for a single state to return the values that would have been returned from applying MCTS for $N$ iterations.
 
@@
 loss function $l = (p * \pi) + (\tau - v)^2$ for each state the NN encountered during self-play.
+
+\section{Explainable state representations}
+
+In this section we describe the
+
+\subsection{Testing With Concept Activation Vectors}
+
+Once the neural network has learned to play the game of Breakthrough, we examine its internal state w.r.t HLC's that we understand.
+The first examined HLC is numbers advantage, that is, the number of pieces that a player has over his opponent.
+The higher level idea to a human player would be that they are in a better position since they have more pieces.
+
+To examine the higher level component we take a look at the internal state of the neural network itself.
+To do this we take a state, we run the state through the neural network, and while the state is propagating through the network we select a layer to split the network, we call that layer $l_{split}$.
+At that point in time the state is a mutated vector $l_{split}(l_{split-1}(...l_0(state)))$ of the initial state.
+We save that vector as a point in the N-dimensional space that it represents, and once we've gathered enough points in that N-dimensional space we're able to train a linear classifier using the HLC as a label.
+We train a Stochastic Gradient Descent classifier, from the SKLearn library, to construct a hyperplane that splits the space into two binary states, one that contains the HLC and another that doesn't.
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/linear_separation}
+    \caption{Trained linear classifier on a 2-dimensional space}
+    \label{fig:scattersplit}
+\end{figure}
+An example of how this would look with only two dimensions is shown in Figure \ref{fig:scattersplit} Once we've successfully trained this linear classifier we can run another state $s$ through the neural network.
+Once the prediction has finished, we view the selected action $a$ of $s$ from the policy vector $p$ from the neural network.
+We view the loss of that selected action $a$.
+Then we apply the backpropagation algorithm up to that same layer $l_{split}$ and evaluate the gradient there.
+If that gradient moves to the direction of the HLC we say that the state includes the HLC, and doesn't if it move away from the HLC.
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/linear_separation_with_direction}
+    \caption{Trained linear classifier on a 2-dimensional space with arrow representing gradient}
+    \label{fig:scattersplitarrow}
+\end{figure}
+We show an example in Figure \ref{fig:scattersplitarrow}.
+where the arrow in the image represents the direction the gradient is moving, if it moves toward the HLC the state $s$ includes the HLC otherwise it does not.
+Using this method we evaluate the internal representation of the state within the neural network, and are able to see whether it recognizes the HLC.

### 2020-08-22 latex/sections/results.tex
latexing

@@
 also a good place to put additional figures and data.}
-hello world
+
+In this section we discuss the results of considering HLC within the game Breakthrough, using a Concept Activation vector for evaluating how the neural network recognizes the HLC's.
+We first take a look at the changes in ephasis of the neural network during training.
+The main point of interest there being whether the nerual network notices simple HLCs early then stops taking them into consideration as the network improves.
+
+\section{Evaluating the improvements of the neural network over generations}
+
+To test which HLC the neural network places its emphasis on during training we trained a neural network for $200$ iterations, taking snapshots of the network every $10$ generations.
+Then we had the neural network play against itself for $100$ games collecting the states it encountered during play.
+These states were then examined by a concept activation vector representing these HLCs.
+Firstly there is the numbers HLC, where the number of pawn the player has minus the number of pawns the opponent had, the breakpoint we selected was $2$ meaning that if you have $2$ more pawns then your opponent, you're in a position where a HLC called numbers advantage is present.
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/number_pawns_trend}
+    \caption{Percentage of selected states containing the HLC numbers advantage}
+    \label{fig:numberadvantage}
+\end{figure}
+
+The Figure \ref{fig:numberadvantage}.
+shows that over the course of training the numbers advantage HLC is only ever a slight factor in the selection of states and we can say that the neural network doesn't really consider number advantage in its selection process.
+
+The second HLC we examined was aggressiveness, that is a state in which you most advanced pawn is $2$ or more squares further than the opponents most advanced pawn.
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/most_advanced_trend.png}
+    \caption{Percentage of selected states containing the HLC aggressiveness}
+    \label{fig:aggressiveness}
+\end{figure}
+
+From Figure \ref{fig:aggressiveness}.
+we can see that the aggressiveness HLC is a growing factor over as the nerual network is trained.
+Generally when your opponent is not skilled this strategy is considered a good one.
+
+The last HLC that we examined was unity, the unity HLC represents the absolute average distance of your pawns from the center row of your pawns.
+This HLC is calculated as the row of your furthest pawn from the starting row $r_{far}$, the row of your nearest pawn from the starting row $r_{near}$.
+Finding the middle row is then $\frac{r_{far} + r_{near}}{2}$, we then take the absolute of the average distance from the pawns to that row.
+To find the point at which this value relates to a state with the HLC unity, we sampled a myriad of states and decided on the value of $0.35$.
+The value of $0.35$ generally allows your states to have two rows that have the majority of the pawns and one or two pawns one row away from the group.
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/unity_trend.png}
+    \caption{Percentage of selected states containing the HLC aggressiveness}
+    \label{fig:unity}
+\end{figure}
+
+The literature of breakthrough implies that being patient and waiting until your opponent makes a mistake is generally the preferred strategy for winning, and the concept activation vector for Unity is generally triggered, and increases with the amount of generations the neural network is trained for.
+These results imply that the neural network does infact recognize successful strategies, and focuses its training in the direction of the successful strategies.

### 2020-08-22 hlc_evaluator/search_algorithms/alphabeta.py
data and alphabeta

@@ -1,0 +1,70 @@
+
+"""
+function negamax(node, depth, α, β, color) is
+    if depth = 0 or node is a terminal node then
+        return color × the heuristic value of node
+
+    childNodes := generateMoves(node)
+    childNodes := orderMoves(childNodes)
+    value := −∞
+    foreach child in childNodes do
+        value := max(value, −negamax(child, depth − 1, −β, −α, −color))
+        α := max(α, value)
+        if α ≥ β then
+            break (* cut-off *)
+    return value
+"""
+
+from game_environments.gamenode import GameNode
+from game_environments.breakthrough.breakthrough import config
+
+"""
+all_heuristics = [
+    "player_piece_amount",
+    "piece_difference",
+    "furthest_piece",
+    "furthest_piece_difference"
+]
+"""
+
+selected_heuristic = None
+
+def alphabeta(node, depth, alpha, beta):
+    global selected_heuristic
+    if depth == 0 or node.is_terminal():
+        # special case as my gamenodes always give their heuristics as positive from their viewpoint
+        heuristic = node.get_heuristics(selected_heuristic)
+        if node.player == config.BLACK:
+            heuristic *= -1
+        return heuristic
+
+    child_nodes = [node.execute_move(move) for move in node.legal_moves]
+    val = -float("inf")
+
+    for node in child_nodes:
+        val = max(val, -alphabeta(node, depth - 1, -beta, -alpha))
+        alpha = max(alpha,val)
+        if alpha >= beta:
+            break
+    return val
+
+def alpha_beta_search(node,depth,heuristic):
+    global selected_heuristic
+    selected_heuristic = heuristic
+    children = [node.execute_move(move) for move in node.legal_moves]
+    is_maxing = node.player == config.WHITE
+    best_val = -float("inf") if is_maxing else float("inf")
+    best_child = None
+    for child in children:
+        child_val = alphabeta(child,depth-1,-float("inf"),float("inf"))
+        if is_maxing and child_val > best_val:
+            best_val = child_val
+            best_child = child
+        elif not is_maxing and child_val < best_val:
+            best_val = child_val
+            best_child = child
+    return best_child
+        
+
+
+

### 2020-11-10 latex/sections/background.tex
iteration

@@
 
-In this chapter we discuss traditional artificial intelligence in the context of search the algorithms we used, what some other similar 
-methods exist as well as the field in general.
+In this chapter, we discuss traditional artificial intelligence in the context of search the algorithms we used, what some other similar methods exist as well as the field in general.
+We allocate a large portion of the section to Breakthrough as that will be the testbed for future sections
 
@@
 
-When doing research on Artificial Intelligence it is important to select 
-an environment that is a suitable abstraction for the task at hand. Environments vary 
-significantly, and can be identified by their characteristic. The characteristics 
-that are generally used to describe environments can be seen in Table \ref{tab:env_characteristics}. 
-CITE AIMABOOK
+When researching Artificial Intelligence it is important to select an environment that is a suitable abstraction for the task at hand.
+Environments vary significantly and can be identified by their characteristic.
+The characteristics that are generally used to describe environments can be seen in Table~\ref{tab:env_characteristics}.
+\cite{Russell:AIModern}
 
@@
     \hline
-    \textbf{characteristic} & \textbf{Values} & \textbf{Description}\\
+    \textbf{characteristic} & \textbf{Values}              & \textbf{Description}                                                \\
     \hline
-    Observable & Fully, Partially & How much of the environment can your agent percieve.\\
+    Observable              & Fully, Partially             & How much of the environment can your agent percieve.                \\
     \hline
-    Agents & Single, Multi & Are there multiple agents playing in the environment.\\
+    Agents                  & Single, Multi                & Are there multiple agents playing in the environment.               \\
     \hline
-    Deterministic & Deterministic, Stochastic & Do the actions your agent do deterministicly impact the environment\\
+    Deterministic           & Deterministic, Stochastic    & Do the actions your agent do deterministicly impact the environment \\
     \hline
-    Episodic & Sequential, Episodic & Are actions episodic or sequential\\
+    Episodic                & Sequential, Episodic         & Are actions episodic or sequential                                  \\
     \hline
@@
 
-Categorizing environments like this gives you the power to find an environment in which 
-a method works and know it can be applied to different environments with the same characteristics.
-Talk about \textbf{agents} in the context of environments as the entities that act within the 
-environment. A game player is an agent as well as the neural network model that exists within 
-a car that is able to drive in the real world.
+Categorizing environments like this gives you the power to find an environment in which a method works and know it can be applied to different environments with the same characteristics.
+Talk about agents in the context of environments as the entities that act within the environment.
+A game player is an agent as well as the neural network model that exists within a car that can drive in the real world.
 
@@
 
-With classical Artificial Intelligence Game Environments are commonly used to validate a method, 
-game environments can be games like Tic-Tac-Toe, Breakthrough, and driving simulators. Game environments 
-are a suitable place to apply AI as they serve an important function as an abstraction of the real 
-world, for instance, a self-driving car agent who is verified to avoid driving into walls in a 
-simulation is possibly safer than one who is not.
+With classical Artificial Intelligence Game Environments are commonly used to validate a method, game environments can be games like Tic-Tac-Toe, Breakthrough, and driving simulators.
+Game environments are a suitable place to apply AI as they serve an important function as an abstraction of the real world, for instance, a self-driving car agent who is verified to avoid driving into walls in a simulation is possibly safer than one who is not.
 
@@
 
-The game breakthrough is a simplified version of chess, the game is set up on an $MxN$ board with 
-cells like in chess, and each player starts with two rows of pawns at either end. The objective of the 
-game is to move one of the pawn pieces to the opposite end of the board. A player wins if 
-either they have reached the opposite end of the board, or has captured all of his opponents pawns.
-The pawns differ from chess pawns in such a way that they can not move two squares on the first move and,
-they can move diagonally as well as forward, this leads to the game being impossible to draw as 
-pieces are always able to move. An example of an initial board in breakthrough can be seen in Figure \ref{fig:initbtboard}
+The game breakthrough is a simplified version of chess, the game is set up on a $MxN$ board with cells like in chess, and each player starts with two rows of pawns at either end.
+The objective of the game is to move one of the pawn pieces to the opposite end of the board.
+A player wins if either they have reached the opposite end of the board or have captured all of his opponents' pawns.
+The pawns differ from chess pawns in such a way that they can not move two squares on the first move and, they can move diagonally as well as forward, this leads to the game being impossible to draw as pieces are always able to move.
+An example of an initial board in breakthrough can be seen in Figure~\ref{fig:initbtboard}
 
@@
   \begin{tabular}{|c|c|}\hline
-    \textbf{Characteristic} & \textbf{Value}\\\hline
+    \textbf{Characteristic} & \textbf{Value} \\\hline
     Observable              & Fully          \\
     Agents                  & Multi          \\
-    Deterministic & Deterministic\\
+    Deterministic           & Deterministic  \\
     Episodic                & Sequential     \\
@@
 
+Categorizing Breakthrough with the characteristics described in Section~\ref{sec:environments}.
+We end up with the description shown in Table~\ref{tab:breakthrough_cat}.
+These characteristics are identical to that of Tic-Tac-Toe, and Chess.
+This categorization is the most common in board games where two-player compete.
+
 \subsection{Heuristics of Breakthrough}
 
-To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance 
-a very simple heuristic would be the amount of pieces each player has left. This heuristics gives us some 
-insight into how well the game is progressing, but obviously there are cases where this doesn't tell us 
-much, as in cases when your opponent has a single piece left that is on the row immediately before the 
-row needed for him to win. No matter how many pieces you have left, this state is bad for you if you're not 
-able to capture that piece. An example of such a state can be seen in Figure \ref{fig:bt_h1_bad}.
+To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance a very simple heuristic would be the number of pieces each player has left.
+This heuristics gives us some insight into how well the game is progressing, but obviously, there are cases where this doesn't tell us much, as in cases when your opponent has a single piece left that is on the row immediately before the row needed for him to win.
+No matter how many pieces you have left, this state is bad for you if you're not able to capture that piece.
+An example of such a state can be seen in Figure~\ref{fig:bt_h1_bad}.
 
@@
 
-A different heuristic would be the distance of your most advanced pawn minus your opponents most 
-advanced pawn, this heuristic could give you insight into how close you are to winning the game or 
-how close your opponent is. As a higher level concept we can call this concept your aggressiveness, 
-as it closely resembles how aggressive you are for going in for the win. Generally in Breakthrough 
-it is favorable to move your whole team as a unit and play more defensively (CITE UNITY HEURISTIC), so this heuristic 
-is probably not one that will lead to the most wins.
+A different heuristic would be the distance of your most advanced pawn minus your opponents most advanced pawn, this heuristic could give you insight into how close you are to winning the game or how close your opponent is.
+As a higher-level concept, we can call this concept your aggressiveness, as it closely resembles how aggressive you are going for in for the win.
+Generally, in Breakthrough, it is favorable to move your whole team as a unit and play more defensively, so this heuristic is probably not one that will lead to the most wins.
 
@@
 
-Traditionally methods for playing games were searching through the environment using a heuristic to
-guide the search. A simple way of doing a heuristic based search would be to give all states a $0$ 
-as a score and give states where you win a positive score. We say that those search algorithm 
-is not guided, and the algorithm will probably have to evaluate every state in the state-space. 
-This method of searching is generally extremely inefficient as the state-spaces of game environments 
-are often extremely large, and sometimes infinite. For instance an estimate of the state-space 
-for the game breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width 
-of the board. The $2^{N*4}$ represent each piece either being alive or captured. So for a small board
-$5x4$ the estimated state-space is $1310720$ many states.
+Traditionally methods for playing games were searching through the environment using a heuristic to guide the search.
+A simple way of doing a heuristic-based search would be to give all states a $0$ as a score and give states where you win a positive score.
+We say that that search algorithm is not guided, and the algorithm will probably have to evaluate every state in the state-space.
+This method of searching is generally extremely inefficient as the state-spaces of game environments are often extremely large, and sometimes infinite.
+For instance, an estimate of the state-space for breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width of the board.
+The $2^{N*4}$ represents each piece either being alive or captured.
+So for a small board, $5x4$ the upper-bound estimation of the state-space is $1310720$ states.
 
-This is why a good heuristic is very valuable because we can disregard all states $S'$ that 
-result from doing a move $a$ in state $S$ as they will only lead to worse outcomes.
+This is why a good heuristic is very valuable because we can disregard all states $S'$ that result from doing a move $a$ in state $S$ as they will only lead to worse outcomes.
 
-The algorithms that are used in traditional state-space search are for instance Depth-first search (DFS),
-Breadth-first Search (BFS), Alpha-Beta Pruning Search (AB-Search), and Monte-Carlo Tree Search (MCTS).
+The algorithms that are used in traditional state-space searches are for instance Depth-first search (DFS), Breadth-first Search (BFS), Alpha-Beta Pruning Search (AB-Search), and Monte-Carlo Tree Search (MCTS).
 
-More modernly these search methods have been amplified by Machine Learning, in such a way that 
-we do not need to figure out a good heuristic for a given state, but rather, we apply a machine 
-learning model to learn a function that takes in a state and returns an evaluation of that state.
-This can lead to a significant time reduction as we do not need to simulate a whole game from a 
-state to receive its evaluation we rather receive the evaluation from the model.
+More modernly, these search methods have been amplified by Machine Learning, in such a way that we do not need to figure out a good heuristic for a given state, but rather, we apply a machine learning model to learn a function that takes in a state and returns an evaluation of that state.
+This can lead to a significant time reduction as we do not need to simulate a whole game from a state to receive its evaluation we rather receive the evaluation from the model.
 
@@
 
-In the algorithm Monte-Carlo Tree Search there is an agent exists
-within some environment. Where each node in the environment represents a state-action pair of the environment, by
-state-action pair what is meant it is some state and the action that brought the agent to
-that state. This pair should be unique within the environment. We will discuss these concepts within game-environments
-and therefore we might say player instead of agent and game instead of environment.
+In the algorithm Monte-Carlo Tree Search, there is an agent within some environment.
+Where each node in the environment represents a state-action pair of the environment, by state-action pair what is meant it is some state and the action that brought the agent to that state.
+This pair should be unique within the environment.
+We will discuss these concepts within game-environments and therefore we might say player instead of agent and game instead of the environment.
 
@@
 
-In MCTS there is a tree representing the game-environment. This tree consists of nodes $n_i$ where $i$ represents the point in time
-of the node, for example $N_0$ in chess is the initial position and $N_x$ is some position in the middle of the game and $N_e$ is one of
-the states representing a position where there is either a draw or one player has won the match. Each of the nodes has $4$ values,
-$s$, $a$, $Q$, and $N$. These values represent these items, $s$ is state of the environment, $a$ is the action that brought
-the previous node $n_(i-1)$ to node $n_i$, $Q$ is the average value of the node (value meaning the outcome of the game),
-and $N$ the amount of times the MCTS algorithm has visited this node. The values $s$ and $a$ uniquely identify a position
-in the environment and are often called state-action pairs.
+MCTS is a method of exploring an environment in a randomized manner (Monte Carlo is the term implying randomness).
+In MCTS there are four stages.
+Selection, Expansion, Simulation, and Back-propagation.
+They happen sequentially and repeatedly.
+MCTS is initialized with a tree consisting of the unexpanded initial state of the environment.
+
+In MCTS there is a tree representing the game-environment.
+This tree consists of nodes $n_i$ where $i$ represents the point in time of the node, for example, $N_0$ in chess is the initial position and $N_x$ is some position in the middle of the game and $N_e$ is one of the states representing a position where there is either a draw or one player has won the match.
+Each of the nodes has $4$ values, $s$, $a$, $Q$, and $N$.
+These values represent these items, $s$ is the state of the environment, $a$ is the action that brought the previous node $n_(i-1)$ to node $n_i$, $Q$ is the average value of the node (value meaning the outcome of the game), and $N$ the number of times the MCTS algorithm has visited this node.
+The values $s$ and $a$ uniquely identify a position in the environment and are often called state-action pairs.
 
@@
 \subsection*{Selection}
-During the selection phase a node $(s,a)$ within the tree which has not yet been expanded is found.
+During the selection phase, a node $(s,a)$ within the tree which has not yet been expanded is found.
 This process uses Upper Confidence Bound on Trees (UCT) to find that node $(s,a)$, the formula is described in Equation \ref{UCT_formula}. For a parent node $(s,a)$
@@
 Next during rollout actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated
-to go to $(s'',a'')$, until a terminal node within the environment is reached. By terminal we mean a
+to go to $(s'',a'')$, until a terminal node within the environment is reached. By terminal, we mean a
 state in which the game is finished. A terminal node in MCTS can generally return any value, but in the
-context of this paper we only return (+1 white wins, -1 black wins, 0 draw).
+context of this paper, we only return (+1 white wins, -1 black wins, 0 draw).
 
@@
 the resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT. Secondly, the backpropagation process
-is modified to instead of doing rollout/simulation to recieve a reward the predicted value $v$ from
+is modified to instead of doing rollout/simulation to receive a reward the predicted value $v$ from
 the neural network is used instead.
@@
 
-Machine Learning (ML) is a research field in which machines apply statistical functions on data to achieve the output that is \textit{correct}, by correct we mean the corresponding result which we expect.
-Generally this a repetitive process where we look at examples of the data, and the algorithm progressively gets closer to the underlying function of the data it is fed.
+Machine Learning (ML) is a research field in which machines apply statistical functions on data to achieve a correct output, by \textit{correct} we mean the corresponding result which we expect.
+Generally, this a repetitive process where we look at examples of the data, and the algorithm progressively gets closer to the underlying function of the data it is fed.
 This process is therefore similar to trial and error for humans.
 ML is a sub-field of Artificial Intelligence.
-ML algorithms come in the form of two classes \textit{Classification}, where the algorithm should find a class representing the data its given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.
+ML algorithms come in the form of two classes \textit{Classification}, where the algorithm should find a class representing the data it is given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.
 
@@
 Neural networks (NN) are popular methods within a sub-field of ML which is called Deep Learning (DL).
-NN's are created to resemble how the human brain functions. In the brain we have neurons which 
-when they get a signal they apply some function to them and if the resulting signal is high 
-enough, they fire to the next neuron. This is how it is done in the neural network model as well.
-There we have neurons that when they get some input, generally a vector of numbers. The neuron 
-takes the sum of that vector, weights the sum by a constant, then applies a non-linear activation 
-function to it. The result of doing this is then passed on to the next neuron. Until a final layer 
-of neurons is reached. At that point we have a value that the neural network corresponds to the 
-input value. This value can be a binary classification (cat or dog image), a regression value (the 
-value of a property), or any number of outputs. It can then be said that a neural network is doing 
-a function approximation of an input to some value. $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.
+NN's are created to resemble how the human brain functions.
+In the brain, we have neurons which when they get a signal they apply some function to them and if the resulting signal is high enough, they fire to the next neuron.
+This is how it is done in the neural network model as well.
+There we have neurons that when they get some input, generally a vector of numbers.
+The neuron takes the sum of that vector, weighs the sum by a constant, then applies a non-linear activation function to it.
+The result of doing this is then passed on to the next neuron.
+Until a final layer of neurons is reached.
+At that point, we have a value that the neural network corresponds to the input value.
+This value can be a binary classification (cat or dog image), a regression value (the value of a property), or any number of outputs.
+It can then be said that a neural network is doing a function approximation of the input to some value.
+$f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.
 
-Neural networks are machine learning models that take in as input anything that is numerical 
-and they apply differentiatable function to that input s.t. they end with some output. This 
-output is then compared to the expected output the difference between these outputs is called 
-a loss. A loss is backpropagated through the neural network, and each function is derived in 
-order to find its slope. We can then modify the weights of the neural network in the direction 
-of the correct output. Leading to a function approximation of this function
-$f_{neuralnetwork}(X_{input}) = O_{expectedoutput}$
+Neural networks are machine learning models that take in as input anything numerical and they apply continuous differentiable functions to that input s.t.
+they end with some output.
+This output is then compared to the expected output the difference between these outputs is called a loss.
+A loss is backpropagated through the neural network, and each function is derived to find its slope.
+We can then modify the weights of the neural network in the direction of the correct output.
+Leading to a function approximation of this function $f_{neuralnetwork}(X_{input}) = O_{expectedoutput}$
 
@@
 
-The field of XAI research is still very far behind it's counterpart AI research,
-within XAI two fields are concidered the largest that is Model Interpretability
-and Model Explainability.
+The field of XAI research is still very far behind its counterpart AI research, within XAI two fields are the largest, those are Model Interpretability and Model Explainability.
 
@@
 pixels of an image the model deems important. Arguably the most popular
-method for this is Saliency Maps (CITE ME). There the pixel values the model
+method for this is Saliency Maps\cite{Koch:saliency}. There the pixel values the model
 deems important are colored in s.t. a human can examine the image and get a
@@
 Methods for explaining models that aren't image recognition models include
-shapley values. There the input is examined against it's output, then iteratively
+shapley values, from the Lundberg \& Lee\cite{LundbergL:shapley}. There the input is examined against it's output, then iteratively
 input values are selected to be fixed. Then the other input values are varied and
@@
 
-A recent paper by Been Kim Et. al (CITATION NEEDED), shows a method for examining
+A recent paper by Been Kim Et. al \cite{Keem:TCAV}, shows a method for examining
 a neural network giving a much more human insight into a prediction. Using Concept
@@
 and is supposed to decide whether the image is of a horse or a zebra, an intuitive
-approach would be to check whether the animal has stipes, or is both white and black.
+approach would be to check whether the animal has stipes, or the animal has both white and black colors.
 That method of determining if a horse is a zebra could then be called a higher-level
@@
 
-CAV's in neural networks are just binary classifiers of the same dataset used to train
-a neural network. The
+The construction of a CAV requires a method of labelling the values in your dataset with
+the corresponding concept in order to create a simple binary classifier on the data points.
 

### 2020-11-10 latex/sections/discussion.tex
iteration

@@
 for instance for health organizations, patients, taxpayers, and pharma}
-hey
+
+This section discusses the results and the future work that could span from this research. Additionally, we conclude the work.
 
@@
 
-\ifdraft{summarize the work}
-hey
+As seen in Chapter~\ref{cha:results}.
+the our trained agent does indeed quickly rise to using the higher level concepts.
+However, there is not a lot of difference between each heuristic, and not much variability in the heuristics once it reaches $\tilde50\%$.
+These results are not what we hoped we would see, as a concept like numbers advantage clearly isn't a great heuristic.
+A reasonable assumption would be that as we train the agent more it will start to use these strategies more, or to increase the size of the neural network to gain performance equal or better to that of a human player.
+
+There is a possibility that the described HLC's are not the HLC's that a neural network uses to play the game, and thus we don't see a connection between the network and the concepts.
+The amount of HLC however is infinite and they must be selected carefully, we selected only ones that are common beginner heuristics in board games.
+A possible avenue of research would be to examine a fully trained neural network that does compete in games at a super human level, a network like AlphaZeroGo and examine it's play style.
+From its play style it would be possible to gather which HLC's it focusess on, then retraining the model examining these HLC's over time to find new pedagogical method of teaching Go.
+
+Additionally applying this method to a greater set of environment would be an interesting field of research.
+We envision a self driving car agent being examined with respect to a higher level concept of aggressiveness, or a mortgage agent being examined with respect to a higher level concept of racism.
 
@@
 \ifdraft{conclude the work, discuss the significance of the work}
-hey
+
+The work done in this paper is only a first step in examining working agents in active environments with respect to human level concepts.
+If improved could have a great impact in our trust on neural network that improve our daily lives.
+While the testbed for the research was a discrete game environment, there is a clear way forward to a dynamic system with discrete human level concepts.
+Our results show that for an agent that clearly doesn't achieve human level performance it's actions are often selected with respect to human level concepts, up to 60\% of the actions.

### 2020-11-10 latex/sections/introduction.tex
iteration

--- 2020-07-17 latex/sections/introduction.tex
@@
 
-The world of deep learning (DL) is exciting, we have models that are able to examine
+The world of deep learning (DL) is exciting, we have models that can examine
 images and very reliably be able to identify it's content. Deep learning models have
-beaten Opthalmologists in identifing diabetic retinopathy, they've identified
+beaten Ophthalmologists in identifying diabetic retinopathy, they've identified
 cancer cells where others have not. Deep learning models have even predicted
-the likelyhood a person will die in the following year with good accuracy (CITE).
+the likelihood a person will die in the following year with good accuracy (CITE).
 These models have done these things extremely accurately, cheap, and fast.
 
-DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search, 
-was used to defeat the standing world champion Lee Sedol in the incredibly expansive game of Go. This was done in the year 
-2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself. 
-These result imply that given enough time to learn the computer agent is able to
-gain a deep insight on how the game should be played.
+DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search was used to defeat the standing world champion Lee Sedol (CITE ALPHAGO ZERO) in the incredibly expansive game of Go.
+This was done in the year 2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself.
+These results imply that given enough time to learn the computer agent can gain a deep insight into how the game should be played.
 
@@
 Furthermore, these explanations are a legal requirement now in many areas, namely, legal, and
-medical. Recently, XAI has generally been focused on Model Interpretability, in
-particular focus on gaining insights on the concepts learned by Deep Neural Networks.
+medical. Recently, XAI has generally been focused on Model Interpretability (MI), in
+particular, interpreting the model in such a way that we might predict what the model will
+do given a particular input. Another class of XAI would be Model Explainability (ME), where we can adequately explain how the internal mechanics of the model works.
+A simple example of when we're able to apply ME would be when we print out a decision tree
+that is generated from an ML algorithm, with the decision tree nodes and is conditions we can see exactly what the tree will do with a given input.
+While in MI we would have to predict the results, without going into details as to why.
 
-In this we thesis takes a look at how we can examine an artificial neural network (ANN)
-to understand which higher level concepts (HLC) it deems important for a given state
+In this thesis, we take a look at how we can examine an artificial neural network (ANN)
+to understand which higher-level concepts (HLC) it deems important for a given state
 within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
@@
 The method of examining HLC's within games has generally been done by examining the
-current state of the game by evaluating them with respect to some heuristic. A heuristic within
-games are evaluations of the end reward for a state given only the state, for example,
-the amount of pieces left within a game of chess. Intuitively, the amount of pieces
-left is a good estimate for a state in chess, this is a higher level concept we
+current state of the game by evaluating them concerning some heuristic. A heuristic within games are evaluations of the end reward for a state given only the state, for example,
+the number of pieces left within a game of chess. Intuitively, the amount of pieces
+left is a good estimate for a state in chess, this is a higher-level concept we
 use to evaluate a chess position. The piece amount can be considered as a lower-level concept
-than other concepts we use. For instance, grand master chess players evaluate a position w.r.t. states
-where the king is safe from attack, or the structure of the pawn positions.
-This paper examines a neural networks evaluation of a state regarding
-those higher-level concepts, if human players evaluate the king safety of a state
+than other concepts we use. For instance, grandmaster chess players evaluate a position w.r.t. states
+where the king is safe from attack or the structure of the pawn positions.
+This paper examines the evaluation of a neural network of a state regarding those higher-level concepts if human players evaluate the king's safety of a state
 low but the neural network highly values it, and the neural network plays better
@@
 
-In this thesis we take an example game of Breakthrough, train a neural network to play the game
-using only self-play. That is, the neural network is trained only by playing against
-itself with no outside input other than the rules of the game itself. And we examine
-the neural networks against popular Higher-level concepts (heuristics) that we use
-to play the game.
+In this thesis, we take an example game of Breakthrough, train a neural network to play the game using only self-play.
+That is, the neural network is trained only by playing against itself with no outside input other than the rules of the game itself.
+And we examine the neural networks against popular Higher-level concepts (heuristics) that we use to play the game.

### 2020-11-10 latex/sections/managerial.tex
latexing
iteration

--- 2020-01-26 latex/sections/managerial.tex
@@
   schedule has rendered me unmeetupwithable. Lastly but certainly not least,
-  Hulda Lilja who always easead my mind during my spurts of imposter syndrome, and fear of
-  failure.
 
@@
 AI & Artificial Intelligence\\
+ANN & Artificial Neural Network\\
+DNN & Deep Neural Network\\
+MCTS & Monte-Carlo Tree Search\\
+CAV & Concept Activation Vector\\
+DL & Deep Learning\\
+MI & Model Interpretability\\
+ME & Model Explainability\\
+
 \end{tabular}

### 2020-11-10 latex/sections/methods.tex
iteration

@@
 
-In this chapter we describe our implementation of the game Breakthrough and the methods for training a neural network to play the game Breakthrough, described in Chapter \ref{cha:background}.
+In this chapter, we describe our implementation of the game Breakthrough and the methods for training a neural network to play the game Breakthrough, described in Chapter \ref{cha:background}.
 The training process uses MCTS described in Section \ref{sec:mcts} to guide its training.
@@
 
-The training algorithm is a reinforcement learning self-play algorithm based on previous work by DeepMind (CITE ME).
+The training algorithm is a reinforcement learning self-play algorithm based on previous work by DeepMind~\cite{silver:alphagozero}.
 
-\section{Implementation}
+\section{Game Implementation}
 
-In this section we will take a brief look at the implementation of the game Breakthrough, how we model its state representation, and the pseudocode for training.
+In this section, we will take a brief look at the implementation of the game Breakthrough, how we model its state representation and the pseudocode for training.
 
@@
 
-In order for a neural network to be able to use input, that input needs to be numeric.
+For a neural network to be able to use input, that input needs to be numeric.
 We model the board in gamethrough with a single $NxM$ array with three values $\{-1,0,1\}$.
-Where each $x,y$ index on the board represents what value is on the board at the time, $-1$ represnts a white piece, $0$ represents an empty square, and $1$ represents a black piece.
+Where each $x,y$ index on the board represents what value is on the board at the time, $-1$ represents a white piece, $0$ an empty square, and $1$ a black piece.
 
-A action vector is coupled with each state, the action vector is a $3$-dimensional matrix where the lengths of the dimensions are $row x columns x 6$ and each index of the vector $X,Y,Z$ represents an action moving from cell $x,y$ moving to the direction $z$.
+An action vector is coupled with each state, the action vector is a $3$-dimensional matrix where the lengths of the dimensions are $N * M * 6$ and each index of the matrix represents an action moving from cell $x,y$ moving to the direction $z$.
 The directions are as follows $0$ represents moving upwards and to the left diagonally on the board, $1$ upwards, and $2$ upwards and to the right diagonally, $4$ downwards and to the left diagonally, $5$ downwards, and lastly $6$ downwards and to the right diagonally.
-
-\subsection{Selfplay}
 
@@
 
-The neural network architecture we opted to use was a single convolutional layer with a ReLU activation function, followed by 5 residual layers each containing two layers of convolution, batch normalization and a ReLu activation.
-Lastly a split policy/value head, where the output of the last residual layer is split into two different outputs.
-The policy head are two layers: first a final convolutional layer, and last a fully connected layer with a \textit{log softmax} layer.
-The value head, consists of three layers: first a convolutional layer, then a fully connected layer, and lastly a fully connected layer with a \textit{tanh} activation function.
+The neural network architecture we opted to use was a single convolutional layer with a ReLU activation function, followed by 5 residual layers each containing two layers of convolution, batch normalization, and a ReLu activation.
+Lastly, a split policy/value head, where the output of the last residual layer is split into two different outputs.
+The policy head are two layers: first a final convolutional layer, and then a fully connected layer with a \textit{log softmax} layer.
+The value head consists of three layers: first a convolutional layer, then a fully connected layer, and lastly a fully connected layer with a \textit{tanh} activation function.
 The Figure \ref{fig:nnarch}.
@@
 
-It is important to understand why we use convolutional layers when dealing with board games as convolution is often associated with image.
+It is important to understand why we use convolutional layers when dealing with board games as convolution is often associated with an image.
 This is because a convolutional layer is focused on merging multiple input parameters to a single neuron, making it an input parameter for the next layer representing the locality around the center point of the original input.
-Playing the game of Breakthrough, a piece in only able to capture pieces in its immediate vicinity.
-This lead to the selection of a convolutional kernel with the size of $3x3$ and a stride of $1$.
+Playing the game of Breakthrough, a piece is only able to capture pieces in its immediate vicinity.
+This lead to the selection of a convolutional kernel with a size of $3x3$ and a stride of $1$.
 
-The knowledge of each piece moves isn't enough to represent the whole gamestate within the neural network.
+The knowledge of each piece moves isn't enough to represent the whole game state within the neural network.
 This is why we use multiple residual layers with convolution.
-These residual layers stack convolutions on each other making the final convolution represent the locality of all the other localities, allowing the neural network to have a representation of the whole gamestate in its parameters.
+These residual layers stack convolutions on each other making the final convolution represent the locality of all the other localities, allowing the neural network to have a representation of the whole game state in its parameters.
 
@@
 The residual layers as a concept remember the output of the previous layer and add the results of the residual layer and the previous layer together.
-On a higher level this leads to the internal representation of the neural network to maintain the whole game boards in its representation, s.t.
+On a higher level, this leads to the internal representation of the neural network to maintain the whole game boards in its representation, s.t.
 two areas that are far away from each other maintain the same level of locality as two that are close.
@@
 
-The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move (straight, and diagonally both left and right for the first player, and backwards for the second player).
+The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move (straight, and diagonally both left and right for the first player, and backward for the second player).
 The vector is then masked s.t.
@@
 
+\subsection{Implementation}
+
+This subsection focuses on the implementation of the various functions required to train the neural network to play breakthrough.
+The description is mainly through pseudo code, and the code is awailable on \href{github.com/sigurdurhelga/msc-cheis}{Github} for examination.
+
 \subsection{Self-play}
 
-To train the neural network to play Breakthrough we initialize two neural networks $N_1$ and $N_2$ 
-with random weights. Then we let $N_1$ play against itself using MCTS to direct its training. 
-While playing against itself the neural network gathers data for it to train with.
+\begin{algorithm}[t]
+\caption{Neural network selfplay pseudocode}
+\label{alg:selfplay}
+\begin{algorithmic}[1]
+\STATE \textbf{neural network} nn1 = randomizedInitialNN()
+\STATE \textbf{neural network} nn2 = randomizedInitialNN()
+\WHILE{\TRUE}
+    \STATE{dataset = \textbf{generate\_dataset}(nn1)}
+    \STATE{nn1.train\_on\_examples(dataset)}
+    \STATE{win\_rate = \textbf{compete}(nn1,nn2)}
+    \IF{win\_rate > $0.5$}
+        \STATE{nn2 = nn1.copy()}
+    \ELSE
+        \STATE{nn1 = nn2.copy()}
+    \ENDIF
+    \STATE \textbf{save}(nn1)
+\ENDWHILE
+\end{algorithmic}
+\end{algorithm}
 
-The data collected is $(\pi, \tau, p, v)$, where $\pi$ is the action probabilities provided by 
-MCTS for $N$ iterations, $\tau$ is the reward for the whole episode, $1$ if white wins,
-$-1$ if black wins, $p$ is the policy vector
-from the neural network, and $v$ is the predicted reward of the game by the neural network.
+To train the neural network to play Breakthrough we initialize two neural networks nn1 and nn2 with random weights.
+Then we let nn1 play against itself using MCTS with PUCT to select moves.
+We collect data from this self-play to train on.
+The data collected is $(s, \pi, \tau, p, v)$, $s$ is the state, $\pi$ is the action probability vector provided by MCTS with PUCT, $\tau$ the end reward for the episode $1$ if white wins, $-1$ if black wins.
+$p$ is the action probability vector predicted by the neural network, and $v$ is the predicted reward of the game by the neural network.
 
-This data collection trains the neural network in such a way that it is a function approximation for a single state to return the values that would have been returned from applying MCTS for $N$ iterations.
+The pseudocode can be seen in Algorithm~\ref{alg:selfplay}.
+
+\subsection{Compete}
+
+\begin{algorithm}[t]
+\caption{Neural network compete pseudocode}
+\label{alg:competition}
+\begin{algorithmic}[1]
+\STATE{\textbf{Input:} neural network nn1, neural network nn2}
+\STATE{\textbf{Output:} win rate for white player}
+\STATE{white\_wins $= 0$}
+\STATE{game = \textbf{initial\_breakthrough}()}
+\FOR{1 \TO 100}
+    \WHILE{\textbf{not} game.is\_termina()}
+        \STATE{nn1.make\_move()}
+        \STATE{nn2.make\_move()}
+    \ENDWHILE
+    \IF{game.white\_wins()}
+        \STATE{white\_wins++}
+    \ENDIF
+\ENDFOR
+\STATE{\textbf{return} white\_wins/100}
+\end{algorithmic}
+\end{algorithm}
+
+The compete function differs from the self-play function in such a way that we select the moves by a single pass through the neural network thereby only evaluating the neural networks ability to predict best actions.
+The pseudocode is available in Algorithm~\ref{alg:competition}.
 
@@
 
-The data collected is backpropagated through the NN moving the weights to the direction of this 
-loss function $l = (p * \pi) + (\tau - v)^2$ for each state the NN encountered during self-play.
+The data collected is backpropagated through the NN moving the weights to the direction of this loss function $l = (p * \pi) + (\tau - v)^2$ for each state the NN encountered during self-play.
+Importantly the variable $p$ has masked illegal actions to $0$ to direct the NN to not learn on illegal moves.
 
@@
 
-In this section we describe the
+In this section we describe the process of trainin a Concept Activation Vector in order to linearly separate each state into points in a space with the concept and ones without the concept.
 
@@
 The first examined HLC is numbers advantage, that is, the number of pieces that a player has over his opponent.
-The higher level idea to a human player would be that they are in a better position since they have more pieces.
+The higher-level idea to a human player would be that they are in a better position since they have more pieces.
 
-To examine the higher level component we take a look at the internal state of the neural network itself.
+To examine the higher-level component we take a look at the internal state of the neural network itself.
 To do this we take a state, we run the state through the neural network, and while the state is propagating through the network we select a layer to split the network, we call that layer $l_{split}$.
@@
 We train a Stochastic Gradient Descent classifier, from the SKLearn library, to construct a hyperplane that splits the space into two binary states, one that contains the HLC and another that doesn't.
+
 \begin{figure}[]
@@
 \end{figure}
+
 An example of how this would look with only two dimensions is shown in Figure \ref{fig:scattersplit} Once we've successfully trained this linear classifier we can run another state $s$ through the neural network.
@@
 Then we apply the backpropagation algorithm up to that same layer $l_{split}$ and evaluate the gradient there.
-If that gradient moves to the direction of the HLC we say that the state includes the HLC, and doesn't if it move away from the HLC.
+If that gradient moves in the direction of the HLC we say that the state includes the HLC, and doesn't if it moves away from the HLC.
+
 \begin{figure}[]
@@
 \end{figure}
+
 We show an example in Figure \ref{fig:scattersplitarrow}.
 where the arrow in the image represents the direction the gradient is moving, if it moves toward the HLC the state $s$ includes the HLC otherwise it does not.
-Using this method we evaluate the internal representation of the state within the neural network, and are able to see whether it recognizes the HLC.
+Using this method we evaluate the internal representation of the state within the neural network and can see whether it recognizes the HLC.

### 2020-11-10 latex/sections/results.tex
iteration

--- 2019-11-28 latex/sections/results.tex
@@
 \chapter{Results}
+
+\label{cha:results}
 
@@
     also a good place to put additional figures and data.}
-hello world
+
+In this section we discuss the results of considering HLC within the game Breakthrough, using a Concept Activation vector for evaluating how the neural network recognizes the HLC's.
+We first evaluate the neural network against various agents to validate that it has learned how to play the game.
+Secondly, we take a look at the changes in the emphasis of the neural network during training.
+The main point of interest there being whether the neural network notices simple HLCs early then stops considering them as the network improves.
+
+\section{Evolution of the neural networks win rate}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.7\textwidth]{graphics/winratevsrandom.png}
+        \end{center}
+        \caption{Winrate vs random agent}
+        \label{fig:winratevsrandom}
+    \end{small}
+\end{figure}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.7\textwidth]{graphics/winratevsmcts.png}
+        \end{center}
+        \caption{Winrate vs monte carlo tree search agent}
+        \label{fig:winratevsmcts}
+    \end{small}
+\end{figure}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.7\textwidth]{graphics/winratevsneuralnetwork.png}
+        \end{center}
+        \caption{Winrate vs neural network trained for 150 generations}
+        \label{fig:winratevsneuralnetwork}
+    \end{small}
+\end{figure}
+
+Examining the neural network is only interesting if the neural network can effectively play the game.
+We first examine the history neural network playing against an agent that only takes random moves, on every $10$ generations.
+The win rate over generations can be seen in Figure \ref{fig:winratevsrandom}.
+It is clear that very early on in the training process, almost immediately after generation $10$ our agent performs nearly perfectly against the random agent.
+This implies that the agent does indeed learn some strategy in the game.
+The next agent we tried the neural network against, was an agent that does regular Monte-Carlo tree search on the game space with $100$ iterations of rollout for each move.
+A graph showing the win rate over generations can be seen in Figure \ref{fig:winratevsmcts}.
+Again, quickly our agent learns some strategy and can win often, but doesn't achieve a perfect win rate after $300$ generations.
+The last agent we tested against was the neural network itself, although only trained to $150$ generations.
+A graph showing the win rate can be seen in Figure \ref{fig:winratevsneuralnetwork}.
+As expected the agent performs poorly until it has trained for more than $150$ generations.
+
+These results imply that our agent certainly understands how to play the game of breakthrough to a certain level.
+However, it should be stated that as we the creators attempt to compete against the agent it generally loses in a way that we would hope the agent should be able to play against.
+It is likely that as we train the neural network more it will become better.
+
+\section{Evaluating the improvements of the neural network over generations}
+
+To test which HLC the neural network places its emphasis on during training we trained a neural network for $300$ generations, taking snapshots of the network every $10$ generations.
+Then we had the neural network play against itself for $100$ games collecting the states it encountered during play.
+These states were then examined by a concept activation vector representing these HLCs.
+
+We test four concepts \textit{Numbers Advantage}, \textit{Aggressiveness}, \textit{Unity}, and \textit{Lorentz-Horey}.
+
+\subsection{Description of concepts}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/dist_number_adv}
+        \end{center}
+    \end{small}
+    \caption{Distribution of piece amount difference}
+    \label{fig:dist_number_adv}
+\end{figure}
+
+The first concept we examine is Numbers Advantage which conveys the idea of having more pawns than your opponent.
+When faced with a board it is difficult to argue for whether having a single pawn up on your opponent constitues as Numbers Advantage or three.
+In order to an appropriate value for this case we used the neural network that had been trained for $300$ generations to generate $20000$ unique states.
+The distribution of the difference of pawns for each player can be seen in Figure~\ref{fig:dist_numbers_adv}.
+From examining the distribution and examining samples we selected the breakpoint of $2$, meaning that if you have $2$ pieces on your opponent you are in a state that has this concept.
+From the $20000$ states only $5.5\%$ of states have this property.
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/dist_aggressiveness}
+        \end{center}
+        \caption{Distribution of difference of furthest pawns}
+        \label{fig:dist_aggressiveness}
+    \end{small}
+\end{figure}
+
+The next concept is Aggressiveness, which describes how much closer your furthest piece is to the opponents edge than your opponents furthest piece to your edge.
+This is then the minimum amount of how many moves it will take you to win the game.
+A distribution of these values can be seen in Figure~\ref{fig:dist_aggressiveness}.
+We examined the distribution, and sampled states from various breaking points and decided on the value of $\geq 1$.
+That is if the value of a state's furthest piece difference is greater than or equal to $1$ that state has the concept of aggressiveness.
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/dist_unity}
+        \end{center}
+        \caption{Difference of distance from center}
+        \label{fig:dist_unity}
+    \end{small}
+\end{figure}
+
+\subsection{Numbers Advantage}
+
+Firstly there is the numbers advantage HLC, where the number of pawns the player has minus the number of pawns the opponent had, the breakpoint we selected was $2$ meaning that if you have $2$ more pawns then your opponent, you're in a position where a HLC called numbers advantage is present.
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/number_pawns_trend}
+    \caption{Percentage of selected states containing the HLC numbers advantage}
+    \label{fig:numberadvantage}
+\end{figure}
+
+The Figure \ref{fig:numberadvantage}.
+shows that throughout training the numbers advantage HLC is only ever a slight factor in the selection of states and we can say that the neural network doesn't consider number advantage in its selection process.
+
+The second HLC we examined was aggressiveness, which is a state in which your most advanced pawn is $2$ or more squares further than the opponents most advanced pawn.
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/most_advanced_trend.png}
+    \caption{Percentage of selected states containing the HLC aggressiveness}
+    \label{fig:aggressiveness}
+\end{figure}
+
+From Figure \ref{fig:aggressiveness}.
+we can see that the aggressiveness HLC is a growing factor over as the neural network is trained.
+Generally, when your opponent is not skilled this strategy is considered a good one.
+
+The third HLC that we examined was unity, the unity HLC represents the absolute average distance of your pawns from the center row of your pawns.
+This HLC is calculated as the row of your furthest pawn from the starting row $r_{far}$, the row of your nearest pawn from the starting row $r_{near}$.
+Finding the middle row is then $\frac{r_{far} + r_{near}}{2}$, we then take the absolute of the average distance from the pawns to that row.
+To find the point at which this value relates to a state with the HLC unity, we sampled a myriad of states and decided on the value of $0.35$.
+The value of $0.35$ generally allows your states to have two rows that have the majority of the pawns and one or two pawns one row away from the group.
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/unity_trend.png}
+    \caption{Percentage of selected states containing the HLC unity}
+    \label{fig:unity}
+\end{figure}
+
+The literature of breakthrough implies that being patient and waiting until your opponent makes a mistake is generally the preferred strategy for winning.
+the concept activation vector for Unity is not often triggered and is relatively stable at $50\%$ througout training.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|}
+        \hline
+        5  & 15 & 5  & 5  & 15 & 5  \\\hline
+        2  & 3  & 3  & 3  & 3  & 2  \\\hline
+        4  & 6  & 6  & 6  & 6  & 4  \\\hline
+        7  & 10 & 10 & 10 & 10 & 7  \\\hline
+        11 & 15 & 15 & 15 & 15 & 11 \\\hline
+        21 & 21 & 21 & 21 & 21 & 21 \\\hline
+    \end{tabular}
+    \caption{Lorentz-Horey cell values, from white players point of view}
+    \label{table:lorentzcell}
+\end{table}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.7\textwidth]{graphics/lorentz_horey_distribution.png}
+        \end{center}
+        \caption{Distribution of Lorentz Horey values}
+        \label{fig:lorentzdistribution}
+    \end{small}
+\end{figure}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/lorentz_horey_trend.png}
+        \end{center}
+        \caption{Percentage of selected states containing the HLC Lorentz-Horey}
+        \label{fig:lorentzheuristic}
+    \end{small}
+\end{figure}
+
+The last heuristic we examined was the Lorentz-Horey heuristic.
+This heuristic is a popular heuristic in breakthrough defined in the paper by Lorentz \& Horey~\cite{lorentz:heuristic}.
+For this heuristic, we give each cell on the board a point value.
+For a given player the heuristic is then the sum of the cell values where they have pawns.
+The cell values can be seen in Table \ref{table:lorentzcell}.
+the matrix shown is flipped in the black player's point of view.
+The values are selected in such a way that as your pawn approach the opponent's side their value increases, having pieces on the edge isn't optimal, as such a pawn will only be able to capture in one direction and can not escape capture as easily.
+To select the breakpoint for Lorentz-Horey heuristic we generated 20000 states and considered the values, the resulting distribution can be seen in Figure \ref{fig:lorentzdistribution}.
+This distribution leads us to select the value of $5$, as a breakpoint indicating that a state has the HLC of Lorentz-Horey.
+The evolution of the neural network using this HLC for selection can be seen in Figure \ref{fig:lorentzheuristic}
+
+Examining Figure~\ref{fig:lorentzheuristic}, we can see that the trend of using this HLC increases over time.
+As this heuristic is the one most often cited as a successful strategy in breakthrough this result is promising.
+
+\section{Examining state properties}
+
+To be able to view the results from the concept activation vectors it is important to know how many states we expect to have a concept so that we can know whether the model is only guessing at random or not.

### 2020-11-10 hlc_evaluator/neural_networks/breakthrough/neuralnetwork.py
minor changes

@@ -120,7 +120,7 @@ class BreakThroughAlphaZero(nn.Module):
     super(BreakThroughAlphaZero, self).__init__()
     self.convBlock = ConvBlock(game_width, game_height, conv_filters)
 
-    self.residual_layer_count = 10
+    self.residual_layer_count = 5
 
     for i in range(self.residual_layer_count):
       setattr(self, f"res_{i}", ResBlock(conv_filters, "res_layer_{}".format(i)))

### 2020-11-10 hlc_evaluator/search_algorithms/alphabeta.py

@@ -51,19 +51,22 @@ def alphabeta(node, depth, alpha, beta):
 def alpha_beta_search(node,depth,heuristic):
     global selected_heuristic
     selected_heuristic = heuristic
-    children = [node.execute_move(move) for move in node.legal_moves]
+    children = [(node.execute_move(move),move) for move in node.legal_moves]
     is_maxing = node.player == config.WHITE
     best_val = -float("inf") if is_maxing else float("inf")
     best_child = None
-    for child in children:
+    best_move = None
+    for child,move in children:
         child_val = alphabeta(child,depth-1,-float("inf"),float("inf"))
         if is_maxing and child_val > best_val:
             best_val = child_val
             best_child = child
+            best_move = move
         elif not is_maxing and child_val < best_val:
             best_val = child_val
             best_child = child
-    return best_child
+            best_move = move
+    return (best_child,best_move)
         
 
 

### 2020-11-10 hlc_evaluator/server.py

@@ -69,18 +69,21 @@ def post_move():
         return jsonify({"valid":False})
 
 neural_network = BreakthroughNN(working_board.gamestate.cols, working_board.gamestate.rows, working_board.gamestate.get_move_amount())
-neural_network.loadmodel('./trained_models', 'session_res5_gen300.tar')
+neural_network.loadmodel('./trained_models', 'session2_res5_gen300.tar')
 
 @app.route('/get_ai_move', methods=['GET'])
 def get_ai_move():
     global neural_network, working_board
     think = request.args.get('think', default=10, type=int)
     monte_tree = MCTS()
-    pi = monte_tree.get_policy(working_board, think, neural_network)
-
+    pi,v = neural_network.safe_predict(working_board.gamestate)
+    pi = pi.detach().cpu().numpy()[0]
+    if not working_board.is_expanded():
+        working_board.expand()
     for i,child in enumerate(working_board.children):
         if not child:
             pi[i] = 0
+
     pi = pi / sum(pi)
 
     working_board = np.random.choice(working_board.children, p=pi)

### 2020-11-10 hlc_evaluator/test_alphabeta.py
random ipynbs

@@ -1,0 +1,16 @@
+from game_environments.breakthrough.breakthrough import BTBoard
+from search_algorithms.alphabeta import alpha_beta_search
+import numpy as np
+
+state = BTBoard(np.zeros([6,6]), 1).initial_state()
+
+while not state.is_terminal():
+    state.print_board()
+    state,_ = alpha_beta_search(state, 5, "player_piece_amount")
+    if not state.is_terminal():
+        state.print_board()
+        state,_ = alpha_beta_search(state, 1, "player_piece_amount")
+
+state.print_board()
+    
+

### 2020-11-10 hlc_evaluator/verify_against_search.py

@@ -1,0 +1,102 @@
+from monte_carlo.mcts import MCTS
+from monte_carlo.mctsnode import Node
+# from game_environments.play_chess.playchess import PlayChess
+from game_environments.gamenode import GameNode
+from game_environments.breakthrough.breakthrough import BTBoard, config as BTconfig
+from neural_networks.breakthrough.breakthrough_nn import BreakthroughNN
+from search_algorithms.alphabeta import alpha_beta_search
+
+import torch
+
+import numpy as np
+from tqdm import tqdm
+
+GAME = {
+    "breakthrough": BTBoard(np.zeros([6,6]), 1),
+}
+
+selected_game = GAME["breakthrough"]
+initial_state = selected_game.initial_state()
+
+CUDA = torch.cuda.is_available()
+
+def selfplay(first_network_path, first_network_name, state_example):
+
+  neural_network_1 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
+
+  test_as_white = True
+
+  neural_network_1.loadmodel(first_network_path, first_network_name)
+  SEARCH_DEPTH = 3
+  initial_node = Node(state_example.initial_state(), "START")
+  first_win = 0
+  second_win = 0
+  total_games = 0
+  while True:
+    for _ in range(20):
+      curr_node = initial_node
+      monte_tree_1 = MCTS()
+      total_games += 1
+      while True:
+        if test_as_white:
+          if not curr_node.is_expanded():
+            curr_node.expand()
+          pi,v = neural_network_1.safe_predict(curr_node.gamestate)
+          if CUDA:
+            pi = pi.detach().cpu().numpy() 
+            pi = pi.reshape(-1)
+          else:
+            pi = pi.view(-1)
+
+          for i,child in enumerate(curr_node.children):
+            if not child:
+              pi[i] = 0
+          pi = pi / sum(pi).item()
+          curr_node = np.random.choice(curr_node.children, p=pi)
+
+          if curr_node.gamestate.is_terminal():
+            if curr_node.gamestate.reward() == 1:
+              first_win += 1
+            break
+        else:
+          next_node,move = alpha_beta_search(curr_node.gamestate, SEARCH_DEPTH, "player_piece_amount")
+          curr_node = Node(next_node, move)
+          if curr_node.gamestate.is_terminal():
+            if curr_node.gamestate.reward() == 1:
+              first_win += 1
+            break
+
+        if not test_as_white:
+          if not curr_node.is_expanded():
[…]

### 2020-11-10 hlc_evaluator/verify_generations_over_time.py

@@ -1,0 +1,129 @@
+from monte_carlo.mcts import MCTS
+from monte_carlo.mctsnode import Node
+# from game_environments.play_chess.playchess import PlayChess
+from game_environments.gamenode import GameNode
+from game_environments.breakthrough.breakthrough import BTBoard, config as BTconfig
+from neural_networks.breakthrough.breakthrough_nn import BreakthroughNN
+
+import torch
+
+import numpy as np
+from tqdm import tqdm
+
+GAME = {
+    "breakthrough": BTBoard(np.zeros([6,6]), 1),
+}
+
+selected_game = GAME["breakthrough"]
+initial_state = selected_game.initial_state()
+
+CUDA = torch.cuda.is_available()
+
+def selfplay(first_network_path, first_network_name, second_network_path, second_network_name, state_example):
+
+  neural_network_1 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
+  neural_network_2 = BreakthroughNN(state_example.cols, state_example.rows, state_example.get_move_amount())
+
+  test_as_white = True
+
+  if test_as_white:
+    neural_network_1.loadmodel(first_network_path, first_network_name)
+    neural_network_2.loadmodel(second_network_path, second_network_name)
+  else:
+    neural_network_2.loadmodel(first_network_path, first_network_name)
+    neural_network_1.loadmodel(second_network_path, second_network_name)
+
+  memo_nn1 = {}
+  memo_nn2 = {}
+
+  initial_node = Node(state_example.initial_state(), "START")
+  first_win = 0
+  second_win = 0
+  total_games = 0
+  while total_games < 200:
+
+    for _ in tqdm(range(5)):
+      curr_node = initial_node
+      monte_tree_1 = MCTS()
+      monte_tree_2 = MCTS()
+      total_games += 1
+      while True:
+
+        # WHITE MOVES
+        # pi = monte_tree_1.get_policy(curr_node, NN_THINK, neural_network_1,0)
+
+        
+        if not curr_node.is_expanded():
+          curr_node.expand()
+        # for i,child in enumerate(curr_node.children):
+          # if not child:
+            # pi[i] = 0
+        # pi = pi / sum(pi).item()
+
+        # """
+        pi,v = neural_network_1.safe_predict(curr_node.gamestate)
+        if CUDA:
+          pi = pi.detach().cpu().numpy() 
+          pi = pi.reshape(-1)
+        else:
+          pi = pi.view(-1)
+
[…]

### 2020-11-12 latex/sections/background.tex
latexing

@@
 In this chapter, we discuss traditional artificial intelligence in the context of search the algorithms we used, what some other similar methods exist as well as the field in general.
-We allocate a large portion of the section to Breakthrough as that will be the testbed for future sections
+We allocate a large portion of the section to Breakthrough as that will be the testbed for future sections.
+We discuss the different classes of machine learning as well as discuss neural networks in more depht.
 
@@
 The characteristics that are generally used to describe environments can be seen in Table~\ref{tab:env_characteristics}.
+This categorization of environments is described in the book Artificial Intelligence by Norvig \& Russell.
 \cite{Russell:AIModern}
 
-\begin{table}[h]
+\begin{table}[ht]
   \centering
@@
     \hline
-    Deterministic           & Deterministic, Stochastic    & Do the actions your agent do deterministicly impact the environment \\
+    Deterministic           & Deterministic, Stochastic    & Do the actions your agent do deterministicly impact the environment.                 \\
     \hline
-    Episodic                & Sequential, Episodic         & Are actions episodic or sequential                                  \\
+    Episodic                & Sequential, Episodic         & Are actions episodic or sequential.                                                  \\
     \hline
-    Static                  & Static, Semi-Static, Dynamic & ehh                                                                 \\
+    Static                  & Static, Semi-Static, Dynamic & Does the environment without agent input, or does it wait until agents take actions. \\
     \hline
-    Discrete                & Discrete, Continuous         & Is your environment discrete w.r.t actions (does it end)            \\
+    Discrete                & Discrete, Continuous         & Is your environment discrete w.r.t actions.                                          \\
     \hline
@@
 Categorizing environments like this gives you the power to find an environment in which a method works and know it can be applied to different environments with the same characteristics.
-Talk about agents in the context of environments as the entities that act within the environment.
-A game player is an agent as well as the neural network model that exists within a car that can drive in the real world.
+Additionally, it allowes us to talk about agents in the context of environments as the entities that act within the environment.
 
@@
 The game breakthrough is a simplified version of chess, the game is set up on a $MxN$ board with cells like in chess, and each player starts with two rows of pawns at either end.
-The objective of the game is to move one of the pawn pieces to the opposite end of the board.
+The objective of the game is for a player to move one of their pawns to the opposite end of the board.
 A player wins if either they have reached the opposite end of the board or have captured all of his opponents' pawns.
@@
 
-  \caption{Example breakthrough board}
+  \caption{Initial breakthrough board}
   \label{fig:initbtboard}
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \centering
@@
 
-Categorizing Breakthrough with the characteristics described in Section \ref{sec:environments}. We end up with the description
-shown in Table \ref{tab:breakthrough_cat}. These characteristics are identical to that of Tic-Tac-Toe, and Chess.
-This categorization is the most common in board games where two player compete.
-
 Categorizing Breakthrough with the characteristics described in Section~\ref{sec:environments}.
@@
 
-To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance a very simple heuristic would be the number of pieces each player has left.
+To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance a very simple heuristic would be a players Material Advantage.
+Material Advantage is the amount of pieces the player has minus the amount of pieces the opponent has.
 This heuristics gives us some insight into how well the game is progressing, but obviously, there are cases where this doesn't tell us much, as in cases when your opponent has a single piece left that is on the row immediately before the row needed for him to win.
@@
   \centering
-  \breakthrough{8/8/pppp4/pp3P2/2p5/8/8/PPP5 w - - 0 1}
-  \caption{Breakthrough board where Heuristic 1 doesn't work well}
+  \breakthrough{8/8/pppp4/pp3P2/2p5/8/8/8 w - - 0 1}
+  \caption{Breakthrough board where Material Advantage doesn't work well}
   \label{fig:bt_h1_bad}
@@
 As a higher-level concept, we can call this concept your aggressiveness, as it closely resembles how aggressive you are going for in for the win.
-Generally, in Breakthrough, it is favorable to move your whole team as a unit and play more defensively, so this heuristic is probably not one that will lead to the most wins.
+Generally, in Breakthrough, it is favorable to move your whole team as a unit and play more defensively, and additional heuristics for Breakthrough will be discussed in detail later in this research.
 
@@
 This method of searching is generally extremely inefficient as the state-spaces of game environments are often extremely large, and sometimes infinite.
-For instance, an estimate of the state-space for breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width of the board.
-The $2^{N*4}$ represents each piece either being alive or captured.
-So for a small board, $5x4$ the upper-bound estimation of the state-space is $1310720$ states.
+For instance, an upper-bound estimate of the state-space for breakthrough is $3^{(M-1)*(N-1)}+2*N$ where $M$ is the height of the board $N$ is the width of the board.
+The $3^{(M-1)*(N-1)}$ represents each position of the board having either a white, or black piece, or being empty.
+So for a small board, $5x4$ the upper-bound estimation of the state-space is $531,449$ states.
 
-This is why a good heuristic is very valuable because we can disregard all states $S'$ that result from doing a move $a$ in state $S$ as they will only lead to worse outcomes.
+This is why a good heuristic is very valuable because we can disregard all following states that result from doing a move some in a previous state as they will only lead to worse outcomes.
 
@@
 
-More modernly, these search methods have been amplified by Machine Learning, in such a way that we do not need to figure out a good heuristic for a given state, but rather, we apply a machine learning model to learn a function that takes in a state and returns an evaluation of that state.
-This can lead to a significant time reduction as we do not need to simulate a whole game from a state to receive its evaluation we rather receive the evaluation from the model.
+More modernly, these search methods have been amplified by Machine Learning, in such a way that we do not need to figure out a good heuristic for a given state, but rather, we apply a machine learning model to learn a function that takes in a state and returns an evaluation of that state.\cite{neuralnetworksgames:michulke} This can lead to a significant time reduction as we do not need to simulate a whole game from a state to receive its evaluation we rather receive the evaluation from the model.
 
@@
 This pair should be unique within the environment.
-We will discuss these concepts within game-environments and therefore we might say player instead of agent and game instead of the environment.
-
-MCTS is a method of exploring an environment in a randomized manner (Monte Carlo is the term implying
-randomness). In MCTS there are four stages. Selection, Expansion, Simulation, and Back-propagation. They happen sequentially
-and repeatedly. MCTS is initialized with a tree consisting of the unexpanded initial state of the environment.
 
@@
 Each of the nodes has $4$ values, $s$, $a$, $Q$, and $N$.
-These values represent these items, $s$ is the state of the environment, $a$ is the action that brought the previous node $n_(i-1)$ to node $n_i$, $Q$ is the average value of the node (value meaning the outcome of the game), and $N$ the number of times the MCTS algorithm has visited this node.
+These values represent these items, $s$ is the state of the environment, $a$ is the action that brought the previous node $n_{i-1}$ to node $n_i$, $Q$ is the average reward from running the MCTS algorithm from this node, and $N$ the number of times the MCTS algorithm has visited this node.
 The values $s$ and $a$ uniquely identify a position in the environment and are often called state-action pairs.
@@
 \subsection*{Expansion}
-Then the expansion phase expands the node generating all of $(s,a)$'s children $(s',a')$ are generated
-and connected to $(s,a)$. This is done by applying all actions $a'$ to $(s,a)$. These children are the
-product of applying each action $a'$ to $s$ in the parent node.
+Then the expansion phase expands the node generating all of $(s,a)$'s children, $(s',a')$ are generated by applying all actions $a'$ in $(s,a)$.
 
 \subsection*{Rollout}
-Next during rollout actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated
-to go to $(s'',a'')$, until a terminal node within the environment is reached. By terminal, we mean a
-state in which the game is finished. A terminal node in MCTS can generally return any value, but in the
-context of this paper, we only return (+1 white wins, -1 black wins, 0 draw).
+Next during rollout, actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated to go to $(s'',a'')$, until a terminal node within the environment is reached.
+By terminal, we mean a state in which the game is finished.
+A terminal node in MCTS can generally return any value, but in the context of this paper, we only return (+1 white wins, or -1 black wins).
 
@@
 When training a neural network the UCT formula is modified slightly to prefer selecting nodes
-that the neural network values highly by introducing a second scalar to the formula $f(s,a) = (p,v)$,
-the resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT. Secondly, the backpropagation process
-is modified to instead of doing rollout/simulation to receive a reward the predicted value $v$ from
-the neural network is used instead.
+that the neural network values highly by introducing a second scalar to the formula $f((s,a)) = (p,v)$, where $f$ is the neural network, $p$ is the policy vector returned by the neural network and $v$ is the predicted value from the neural network.
+The resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT.
+Secondly, the backpropagation process
+is modified to instead of doing rollout/simulation to receive a reward the predicted value $v$ from the neural network is used instead.
 
 \begin{equation} \label{PUCT_formula}
-  \text{Child PUCT value} = \frac{Q_{(s',a')}}{N_{(s',a')}} + c_{uct} * P((s,a)) * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
+  \text{Child PUCT value} = \frac{Q_{(s',a')}}{N_{(s',a')}} + c_{uct} * p_{(s,a)} * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
 \end{equation}
@@
 
+Model Interpretability is the more common approach of XAI, as it is
+
 \subsubsection{Saliency Maps}
+
 Within XAI many methods have been developed to try to evaluate
@@
 Methods for explaining models that aren't image recognition models include
-shapley values, from the Lundberg \& Lee\cite{LundbergL:shapley}. There the input is examined against it's output, then iteratively
+Shapley values, from the Lundberg \& Lee\cite{LundbergL:shapley}. There the input is examined against it's output, then iteratively
 input values are selected to be fixed. Then the other input values are varied and

### 2020-11-12 latex/sections/discussion.tex
latexing

--- 2019-11-28 latex/sections/discussion.tex
@@
 
-\ifdraft{Here I will discuss the myriad of fields research like this can help
-for instance for health organizations, patients, taxpayers, and pharma}
-hey
+This section discusses the results and the future work that could span from this research. Additionally, we conclude the work.
 
@@
 
-\ifdraft{summarize the work}
-hey
+As seen in Chapter \ref{cha:results}, our trained agent does indeed quickly rise to using the higher-level concepts.
+The results are promising, as we see a popular concepts like Lorentz-Horey rise in emphasis for the neural network, and a simple but effective concept like Material Advantage rise slowly but not exessively.
+We find the Unity concept to be disappointing as stated earlier in this paper maintaining closeness of your pawns tends to be preferential.
+And lastly, for a the concept Aggressiveness, the fact that it falls off early is what we expected.
+It would have been the preferred result if we saw a greater variability in the concepts, but this could be a result from not enough training iterations, or not a complex enough model.
+Importantly, these concepts are only a few possibilities of an infinite set of concepts, and the neural network could be learning a completely different concept then those that we tested.
+However, that was not the focus of this research, it was to examine if a neural network does move towards HLC's that we humans use in games.
+
+\section{Future Work}
+
+To iterate on this research, training a neural network with greater computing power will allow the researchers to hopefully achieve a super-human neural network in Breakthrough.
+This would give a greater confidence in the learned concepts, possibly allowing for pedagological research on the topic regarding which concepts are optimal to train people in the game environment.
+This could also arise from a larger neural network architecture.
+
+An avenue of research would be to use a trained AlphaZero model that plays chess to a super human ability, and be able to draw from a more diverse set of concepts.
+Furthermore, one could examine the play-style of the super-human model to extract concepts in order to construct new and improved heuristics for a non-neural network based agent.
+
+Additionally applying this method to a greater set of environment would be an interesting field of research.
+We envision a self driving car agent being examined with respect to a higher level concept of aggressiveness, or a mortgage agent being examined with respect to a higher level concept of gender bias.
 
@@
 
-\ifdraft{conclude the work, discuss the significance of the work}
-hey
+The work done in this paper is only a first step in examining working agents in active environments with respect to human level concepts.
+If improved could have a great impact in our trust on neural network that improve our daily lives.
+While the testbed for the research was a discrete game environment, there is a clear way forward to a dynamic system with discrete human level concepts.
+Our results show that for an agent that clearly doesn't achieve human level performance it's actions are often selected with respect to human level concepts, up to 60\% of the actions.

### 2020-11-12 latex/sections/introduction.tex
latexing

@@
 %% These directions will disappear in other modes.
-\ifdraft{State the objectives of the exercise. Ask yourself:
-  \underline{Why} did I design/create the item? What did I aim to
-  achieve? What is the problem I am trying to solve?  How is my
-  solution interesting or novel?}{}
 
@@
 beaten Ophthalmologists in identifying diabetic retinopathy, they've identified
-cancer cells where others have not. Deep learning models have even predicted
-the likelihood a person will die in the following year with good accuracy (CITE).
+cancer cells where others have not. Deep learning models are used for the bleeding
+edge of protein folding, to gain further understanding of the underlying structure of organisms, such as, viruses. \cite{deepmind:alphafold}
 These models have done these things extremely accurately, cheap, and fast.
@@
 the models themselves but could also augment our ability in many cases.
-Furthermore, these explanations are a legal requirement now in many areas, namely, legal, and
+Furthermore, these explanations are a legal requirement now in many areas\cite{legalexplanation:goodman}, namely, legal, and
 medical. Recently, XAI has generally been focused on Model Interpretability (MI), in
@@
 
-We define the research questions
+We define the research questions:
 
@@
         HLCs that human players use.
-  \item Does a NN that trains itself using self-play learn these HLCs over time, and does it start to emphasize 
-  the HLCs that are generally taught later to human players, more as it trains itself more.
-  \item Does a NN that plays better than some humans focus on HLCs that are not generally 
-  considered by those human players and it.
+  \item Does a NN that trains itself using self-play learn some HLCs over time, and does it start to emphasize
+        the HLCs that are generally considered better, more as it trains more, and oppositely does it start to
+        demphasize HLCs that are considered worse.
 \end{enumerate}

### 2020-11-12 latex/sections/methods.tex
latexing

@@
 
-In this chapter, we describe our implementation of the game Breakthrough and the methods for training a neural network to play the game Breakthrough, described in Chapter \ref{cha:background}.
-The training process uses MCTS described in Section \ref{sec:mcts} to guide its training.
+In this chapter, we describe our implementation of the game Breakthrough and the methods for training a neural network to play the game.
+The training process uses MCTS, as described in Section \ref{sec:mcts}, to guide its training.
 We outline the architecture used in the neural network.
@@
 
-In this section, we will take a brief look at the implementation of the game Breakthrough, how we model its state representation and the pseudocode for training.
+In this section, we will take a look at the implementation of the Breakthrough neural network, how we model its state representation and the pseudo-code for training.
 
-\subsection{Breakthrough}
+\subsection{Description of the input and output}
 
 For a neural network to be able to use input, that input needs to be numeric.
-We model the board in gamethrough with a single $NxM$ array with three values $\{-1,0,1\}$.
-Where each $x,y$ index on the board represents what value is on the board at the time, $-1$ represents a white piece, $0$ an empty square, and $1$ a black piece.
+We model the board with a single $NxMx3$ array with two values $\{0,1\}$, where the first two indices on the $z$ axis represent the players, and the last layer on the $z$ axis represents which players turn it is.
+The first $Z$ index represents whites pawns, that is, each $x,y$ position on the board where white has a pawn that cell $x,y,0$ has the value $1$ otherwise it has $0$, and similarily on the second layer for the black player.
+The last $z$ layer, contains all $1$'s if it is the white players turn otherwise $0$.
 
-An action vector is coupled with each state, the action vector is a $3$-dimensional matrix where the lengths of the dimensions are $N * M * 6$ and each index of the matrix represents an action moving from cell $x,y$ moving to the direction $z$.
+One of the outputs of the nerual network is the policy vector, the policy vector is a $3$-dimensional matrix where the lengths of the dimensions are $N * M * 6$ and each index of the matrix represents an action moving from cell $x,y$ moving to the direction $z$.
 The directions are as follows $0$ represents moving upwards and to the left diagonally on the board, $1$ upwards, and $2$ upwards and to the right diagonally, $4$ downwards and to the left diagonally, $5$ downwards, and lastly $6$ downwards and to the right diagonally.
+The cell values indicate the probability distribution of the neural network selecting that move.
+Importantly, before we select a move from the neural networks prediction we zero out the illegal moves, and renormalize the array such that it sums to $1$.
 
@@
 The value head consists of three layers: first a convolutional layer, then a fully connected layer, and lastly a fully connected layer with a \textit{tanh} activation function.
-The Figure \ref{fig:nnarch}.
+Figure~\ref{fig:nnarch}.
 depicts the architecture.
+Using this architecture we end up with $1737986$ trainable parameters.
 
@@
 
-It is important to understand why we use convolutional layers when dealing with board games as convolution is often associated with an image.
+It is important to understand why we use convolutional layers when dealing with board games as convolution is more typically associated with an image.
 This is because a convolutional layer is focused on merging multiple input parameters to a single neuron, making it an input parameter for the next layer representing the locality around the center point of the original input.
 Playing the game of Breakthrough, a piece is only able to capture pieces in its immediate vicinity.
-This lead to the selection of a convolutional kernel with a size of $3x3$ and a stride of $1$.
+This lead to the selection of a $3$-d convolutional kernel with a size of $3x3$ and a stride of $1$.
 
-The knowledge of each piece moves isn't enough to represent the whole game state within the neural network.
+We require a global view of the board.
 This is why we use multiple residual layers with convolution.
@@
 
-The residual layers as a concept remember the output of the previous layer and add the results of the residual layer and the previous layer together.
+The residual layers as a concept remember the output of the previous layer and add together the results of the residual layer and the previous layer.
 On a higher level, this leads to the internal representation of the neural network to maintain the whole game boards in its representation, s.t.
@@
 
-This subsection focuses on the implementation of the various functions required to train the neural network to play breakthrough.
-The description is mainly through pseudo code, and the code is awailable on \href{github.com/sigurdurhelga/msc-cheis}{Github} for examination.
+This subsection focuses on the implementation of the various functions required to train the neural network to play Breakthrough.
+The description is mainly through pseudo-code, and the code is available on GitHub for examination~\cite{siggi:github}.
 
@@
 \begin{algorithm}[t]
-\caption{Neural network selfplay pseudocode}
+    \caption{Neural network self-play pseudo-code}
     \label{alg:selfplay}
@@
 We collect data from this self-play to train on.
-The data collected is $(s, \pi, \tau, p, v)$, $s$ is the state, $\pi$ is the action probability vector provided by MCTS with PUCT, $\tau$ the end reward for the episode $1$ if white wins, $-1$ if black wins.
-$p$ is the action probability vector predicted by the neural network, and $v$ is the predicted reward of the game by the neural network.
+The data collected is $(s, \pi, \tau, p, v)$, where $s$ is the state, $\pi$ is the action policy vector provided by MCTS with PUCT, $\tau$ the end reward for the episode $1$ if white wins, $-1$ if black wins.
+$p$ is the policy vector predicted by the neural network, and $v$ is the predicted reward of the game by the neural network.
 
-The pseudocode can be seen in Algorithm~\ref{alg:selfplay}.
+The pseudo-code is shown as Algorithm~\ref{alg:selfplay}.
 
@@
 \begin{algorithm}[t]
-\caption{Neural network compete pseudocode}
+    \caption{Neural network compete pseudo-code}
     \label{alg:competition}
@@
 The compete function differs from the self-play function in such a way that we select the moves by a single pass through the neural network thereby only evaluating the neural networks ability to predict best actions.
-The pseudocode is available in Algorithm~\ref{alg:competition}.
+The pseudo-code shown as Algorithm~\ref{alg:competition}.
 
@@
 Once the neural network has learned to play the game of Breakthrough, we examine its internal state w.r.t HLC's that we understand.
-The first examined HLC is numbers advantage, that is, the number of pieces that a player has over his opponent.
-The higher-level idea to a human player would be that they are in a better position since they have more pieces.
+The first examined HLC is material advantage, that is, the number of pieces that a player has over the opponent.
+The higher-level idea to human players would be that they are in a better position since they have more pieces.
 
 To examine the higher-level component we take a look at the internal state of the neural network itself.
-To do this we take a state, we run the state through the neural network, and while the state is propagating through the network we select a layer to split the network, we call that layer $l_{split}$.
+To do this, we take a state, and run it through the neural network, and while the state is propagating through the network we select a layer to split the network, we call that layer $l_{split}$.
 At that point in time the state is a mutated vector $l_{split}(l_{split-1}(...l_0(state)))$ of the initial state.
 We save that vector as a point in the N-dimensional space that it represents, and once we've gathered enough points in that N-dimensional space we're able to train a linear classifier using the HLC as a label.
-We train a Stochastic Gradient Descent classifier, from the SKLearn library, to construct a hyperplane that splits the space into two binary states, one that contains the HLC and another that doesn't.
+We train a Stochastic Gradient Descent classifier, from the SKLearn library, to construct a hyper-plane that splits the space into two binary states, one that contains the HLC and another that doesn't.
 
@@
 
-An example of how this would look with only two dimensions is shown in Figure \ref{fig:scattersplit} Once we've successfully trained this linear classifier we can run another state $s$ through the neural network.
+An example of how this would look with only two dimensions is shown in Figure \ref{fig:scattersplit}.
+Once we've successfully trained this linear classifier we can run another state $s$ through the neural network.
 Once the prediction has finished, we view the selected action $a$ of $s$ from the policy vector $p$ from the neural network.
 We view the loss of that selected action $a$.
-Then we apply the backpropagation algorithm up to that same layer $l_{split}$ and evaluate the gradient there.
-If that gradient moves in the direction of the HLC we say that the state includes the HLC, and doesn't if it moves away from the HLC.
+We then apply the backpropagation algorithm up to that same layer $l_{split}$ and evaluate the gradient there.
+If that gradient moves in the direction of the HLC we say that the state includes the HLC, and that it doesn't if the gradient moves away from the HLC.
 
@@
 
-We show an example in Figure \ref{fig:scattersplitarrow}.
-where the arrow in the image represents the direction the gradient is moving, if it moves toward the HLC the state $s$ includes the HLC otherwise it does not.
+We show an example in Figure \ref{fig:scattersplitarrow}, where the arrow in the image represents the direction the gradient is moving, if it moves toward the HLC the state $s$ includes the HLC otherwise it does not.
 Using this method we evaluate the internal representation of the state within the neural network and can see whether it recognizes the HLC.
+
+\section{Summary}
+
+In this chapter we described the architecture of the nerual network, including how we model Breakthrough to be able to learn how to play the game.
+We described the processes we use to learn completely by self play.
+In the following chapter we will evaluate the neural network.

### 2020-11-12 latex/sections/results.tex
latexing

--- 2019-11-28 latex/sections/results.tex
@@
 
-\ifdraft{In this section you discuss any issues that came up while developing
-the system.  If you found something particularly interesting,
-difficult, or an important learning experience, put it here.  This is
-also a good place to put additional figures and data.}
-hello world
+\label{cha:results}
+
+In this chapter, we first evaluate the neural network against various agents to validate that it has learned how to play the game.
+Secondly, we describe each concept we intend to examine the neural network against.
+Thirdly, we take a look at the changes in the emphasis of the neural network during training.
+The main point of interest there being whether the neural network notices simple HLCs early then stops considering them as the network improves.
+
+\section{Evolution of the Neural Networks Win Rate}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.7\textwidth]{graphics/winratevsrandom.png}
+        \end{center}
+        \caption{Winrate vs random agent}
+        \label{fig:winratevsrandom}
+    \end{small}
+\end{figure}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.7\textwidth]{graphics/winratevsmcts.png}
+        \end{center}
+        \caption{Winrate vs MCTS agent}
+        \label{fig:winratevsmcts}
+    \end{small}
+\end{figure}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.7\textwidth]{graphics/winratevsneuralnetwork.png}
+        \end{center}
+        \caption{Winrate vs neural network trained for 150 generations}
+        \label{fig:winratevsneuralnetwork}
+    \end{small}
+\end{figure}
+
+Examining the neural network is only interesting if the neural network can effectively play the game.
+We first examine the history neural network playing against an agent that only takes random moves, on every $10$ generations.
+The win rate over generations can be seen in Figure \ref{fig:winratevsrandom}.
+It is clear that very early on in the training process, almost immediately after generation $10$ our agent performs nearly perfectly against the random agent.
+This implies that the agent does indeed learn some strategy in the game.
+The next agent we tried the neural network against, was an agent that does regular Monte-Carlo tree search on the game space with $100$ iterations of rollout for each move.
+A graph showing the win rate over generations is depicted in Figure \ref{fig:winratevsmcts}.
+Again, our agent quickly learns some strategy and can win often, although doesn't achieve a perfect win rate after $300$ generations.
+The last agent we tested against was the neural network itself, although only trained to $150$ generations.
+The graph in Figure \ref{fig:winratevsneuralnetwork} shows win rate.
+As expected the agent performs poorly until it has trained for more than $150$ generations.
+
+These results imply that our agent certainly understands how to play the game of breakthrough to a certain level.
+However, an intermediate level human player would most often be able to win the agent, based on the author's experience playing against it.
+Note, that the main focus of this was not to create an expert level agent.
+Rather to, gauge into the concepts the agent learns as described in the next section.
+
+\section{Evaluating the improvements of the Neural Network over generations}
+
+To test which HLC the neural network places its emphasis on during training we trained a neural network for $300$ generations, taking snapshots of the network every $10$ generations.
+We then had the neural network play against itself for $100$ games, collecting the states it encountered during play.
+These states were then examined by a concept activation vector representing these HLCs.
+
+We test four concepts \textit{Material Advantage}, \textit{Aggressiveness}, \textit{Unity}, and \textit{Lorentz-Horey}.
+
+\subsection{Description of concepts}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/dist_number_adv}
+        \end{center}
+    \end{small}
+    \caption{Distribution of piece amount difference}
+    \label{fig:distnumberadv}
+\end{figure}
+
+The first concept we examine is Material Advantage which conveys the idea of having more pawns than your opponent.
+When faced with a board it is difficult to argue for whether having a single pawn up on your opponent constitues as Material Advantage or three.
+In order to an appropriate value for this case we used the neural network that had been trained for $300$ generations to generate $20000$ unique states.
+The distribution of these states' difference of pawns can be seen in Figure \ref{fig:distnumberadv}.
+From examining the distribution and examining samples we selected the breakpoint of $\geq 2$, meaning that if you have $2$ pieces on your opponent you are in a state that has this concept.
+From the $20000$ states only $5.5\%$ of states have this concept.
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/dist_aggressiveness}
+        \end{center}
+        \caption{Distribution of difference of furthest pawns}
+        \label{fig:distaggressiveness}
+    \end{small}
+\end{figure}
+
+The next concept is Aggressiveness, which describes how much closer your furthest piece is to the opponents edge than your opponents furthest piece to your edge.
+This is then the minimum amount of how many moves it will take you to win the game.
+A distribution of these values can be seen in Figure \ref{fig:distaggressiveness}.
+We examined the distribution, and sampled states from various breaking points and decided on the value of $\geq 1$.
+That is if the value of a state's furthest piece difference is greater than or equal to $1$ that state has the concept of Aggressiveness.
+From the dataset $18.3\%$ of states have this concept.
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/dist_unity}
+        \end{center}
+        \caption{Difference of distance from center}
+        \label{fig:distunity}
+    \end{small}
+\end{figure}
+
+Our third concept is Unity.
+This concept represents the absolute average distance of your pawns from the center row of your pawns, and is calculated as follows: the row of your furthest pawn from the starting row $r_{far}$, the row of your nearest pawn from the starting row $r_{near}$.
+Finding the middle row is then $\frac{r_{far} + r_{near}}{2}$, we then take the absolute of the average distance from all of the players pawns to that row.
+The distribution of the average distance values is shown in Figure \ref{fig:distunity}.
+From sampling states from several points we decided on a breakpoint of $\le 0.25$ for identifying states as states where the Unity concept is precent.
+The value of $0.25$ generally allows your states to have two rows that have the majority of the pawns and one or two pawns one row away from the group.
+From the dataset of 20,000 states $20\%$ of states have this concept.
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/dist_lorentz}
+        \end{center}
+        \caption{Difference of Lorentz-Horey Score}
+        \label{fig:distlorentz}
+    \end{small}
+\end{figure}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|}
+        \hline
+        5  & 15 & 5  & 5  & 15 & 5  \\\hline
+        2  & 3  & 3  & 3  & 3  & 2  \\\hline
+        4  & 6  & 6  & 6  & 6  & 4  \\\hline
+        7  & 10 & 10 & 10 & 10 & 7  \\\hline
+        11 & 15 & 15 & 15 & 15 & 11 \\\hline
+        21 & 21 & 21 & 21 & 21 & 21 \\\hline
+    \end{tabular}
+    \caption{Lorentz-Horey cell values, from white players point of view}
+    \label{table:lorentzcell}
+\end{table}
+
+The last concept we examined was the Lorentz-Horey.
+This concept is a popular heuristic in Breakthrough defined in the paper by Lorentz \& Horey \cite{lorentz:heuristic}.
+For this heuristic, we give each cell on the board a point value.
+For a given player the heuristic is then the sum of the cell values where they have pawns.
+The cell values are shown in Table \ref{table:lorentzcell}.
+The matrix shown is flipped from the black player's point of view.
+The values are selected in such a way that as your pawn approach the opponent's side their value increases, having pieces on the edge isn't optimal, as such a pawn will only be able to capture in one direction and can not escape capture as easily.
+To select the breakpoint for Lorentz-Horey heuristic, we generate a distribution of the difference in Lorentz-Horey Score over our dataset.
+The resulting distribution is shown in Figure \ref{fig:distlorentz}.
+Our examination lead us to select the value of $5$, as a breakpoint indicating that a state has the concept of Lorentz-Horey.
+Out of our dataset of 20,000 states $29.4\%$ of states have this concept.
+
+\subsection{Material Advantage}
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/number_pawns_trend}
+    \caption{Percentage of selected states containing the HLC Material Advantage}
+    \label{fig:numberadvantage}
+\end{figure}
+
+The Figure \ref{fig:numberadvantage}.
+shows that throughout training the Material Advantage HLC is only ever a slight factor in the selection of states and we can say that the neural network doesn't consider number advantage in its selection process.
+
+\subsection{Agressiveness}
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/most_advanced_trend.png}
+    \caption{Percentage of selected states containing the HLC aggressiveness}
+    \label{fig:aggressiveness}
+\end{figure}
+
+From Figure \ref{fig:aggressiveness}.
+we can see that the aggressiveness HLC is a growing factor over as the neural network is trained.
+As a piece can always move forward, if your opponent does not capture that piece, this concept relates to how many moves it will take for you to win.
+Generally, when your opponent is not playing well this concept does well, and quickly becomes a bad heuristic in classical search algorithms.
+
+\subsection{Unity}
+
+\begin{figure}[]
+    \centering
+    \includegraphics[width=0.7\textwidth]{graphics/unity_trend.png}
+    \caption{Percentage of selected states containing the HLC unity}
+    \label{fig:unity}
+\end{figure}
+
+\begin{figure}
+    \begin{small}
+        \begin{center}
+            \includegraphics[width=0.95\textwidth]{graphics/lorentz_horey_trend.png}
+        \end{center}
+        \caption{Percentage of selected states containing the HLC Lorentz-Horey}
+        \label{fig:lorentzheuristic}
+    \end{small}
+\end{figure}
+
+The literature of Breakthrough implies that being patient and waiting until your opponent makes a mistake is generally the preferred strategy for winning.\cite{lorentz:heuristic} Examining Figure \ref{fig:lorentzheuristic}, we can see that the trend of using this HLC increases over time.
+As this heuristic is the one most often cited as a successful strategy in Breakthrough, it is interesting to see that our Breakthrough agent is gradually learning a similar concept.
+
+\section{Summary}
+
+In this section we discussed some concepts that we believed would be representative of concepts that a nerual network would learn over time playing Breakthrough.
+An important note is that there are an infinite amount of concepts, and there could exist some that the neural network uses extensively that we overlooked.
+We examined the neural network against these concepts.
+In the following chapter we will discuss the next steps for this research.

### 2020-11-16 latex/sections/background.tex
latexing

@@
 
-The algorithms that are used in traditional state-space searches are for instance Depth-first search (DFS), Breadth-first Search (BFS), Alpha-Beta Pruning Search (AB-Search), and Monte-Carlo Tree Search (MCTS).
+The algorithms that are used in traditional state-space searches are for instance Depth-first search (DFS), Breadth-first Search (BFS), Alpha-Beta Pruning Search (AB-Search)\cite{abpruning:dj}, and Monte-Carlo Tree Search (MCTS).
 
@@
 
-In the algorithm Monte-Carlo Tree Search, there is an agent within some environment.
+In the algorithm Monte-Carlo Tree Search described by R.
+Coulom \cite{mcts:coulom}, there is an agent within some environment.
 Where each node in the environment represents a state-action pair of the environment, by state-action pair what is meant it is some state and the action that brought the agent to that state.
@@
 
-Many algorithms are popular in reinforcement learning, for instance Q-learning(CITE ME), CARLA (CITE ME),
-and many others.
+Many algorithms are popular in reinforcement learning, for instance Q-learning\cite{qlearning:watkins}, and many others.
 
@@
 $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.
-
-Neural networks are machine learning models that take in as input anything numerical and they apply continuous differentiable functions to that input s.t.
-they end with some output.
-This output is then compared to the expected output the difference between these outputs is called a loss.
-A loss is backpropagated through the neural network, and each function is derived to find its slope.
-We can then modify the weights of the neural network in the direction of the correct output.
-Leading to a function approximation of this function $f_{neuralnetwork}(X_{input}) = O_{expectedoutput}$
 
@@
 based on the value of the input value. It is therefore easy to see how to examine the tree
-to explain the output. We just follow the branches in the tree. That being said the branches
-are created by algorithms like ID3 which construct branches based on the initial dataset used
-to construct the tree, but again ID3 follows simple statistics and can be explained properly.
+to explain the output. We just follow the branches in the tree.
 

### 2020-12-09 latex/sections/background.tex
latex updates

@@
 
-In this chapter, we discuss traditional artificial intelligence in the context of search the algorithms we used, what some other similar methods exist as well as the field in general.
+In this chapter, we discuss traditional artificial intelligence in the context of search, the algorithms we used, what some other similar methods exist as well as the field in general.
 We allocate a large portion of the section to Breakthrough as that will be the testbed for future sections.
-We discuss the different classes of machine learning as well as discuss neural networks in more depht.
+We discuss the different classes of machine learning as well as neural networks in more depth.
 
@@
 
-With classical Artificial Intelligence Game Environments are commonly used to validate a method, game environments can be games like Tic-Tac-Toe, Breakthrough, and driving simulators.
-Game environments are a suitable place to apply AI as they serve an important function as an abstraction of the real world, for instance, a self-driving car agent who is verified to avoid driving into walls in a simulation is possibly safer than one who is not.
+Classical Artificial Intelligence Game Environments are commonly used to validate a method, e.g game environments can be games like Tic-Tac-Toe, Breakthrough, and driving simulators.
+Game environments are a suitable place to apply AI as they serve as an abstraction of the real world, for instance, a self-driving car agent who is verified to avoid driving into walls in a simulation is possibly safer than one who is not.
 
@@
 
-The game breakthrough is a simplified version of chess, the game is set up on a $MxN$ board with cells like in chess, and each player starts with two rows of pawns at either end.
+The game Breakthrough is a simplified version of chess; the game is set up on a $MxN$ board with squares like in chess, and each player starts with two rows of pawns at opposite ends.
 The objective of the game is for a player to move one of their pawns to the opposite end of the board.
 A player wins if either they have reached the opposite end of the board or have captured all of his opponents' pawns.
-The pawns differ from chess pawns in such a way that they can not move two squares on the first move and, they can move diagonally as well as forward, this leads to the game being impossible to draw as pieces are always able to move.
-An example of an initial board in breakthrough can be seen in Figure~\ref{fig:initbtboard}
+The pawns differ from chess pawns in such a way that they can not move two squares on the first move and, they can move diagonally as well as forward.
+This leads to the game being impossible to draw as pieces are always able to move.
+An example of an initial board in Breakthrough can be seen in Figure~\ref{fig:initbtboard}
 
@@
 We end up with the description shown in Table~\ref{tab:breakthrough_cat}.
-These characteristics are identical to that of Tic-Tac-Toe, and Chess.
+These characteristics are identical to that of Tic-Tac-Toe, and chess.
 This categorization is the most common in board games where two-player compete.
@@
 
-To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance a very simple heuristic would be a players Material Advantage.
-Material Advantage is the amount of pieces the player has minus the amount of pieces the opponent has.
+To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance a very simple heuristic would be a players material advantage.
+\textit{Material Advantage} is the amount of pieces the player has minus the amount of pieces the opponent has.
 This heuristics gives us some insight into how well the game is progressing, but obviously, there are cases where this doesn't tell us much, as in cases when your opponent has a single piece left that is on the row immediately before the row needed for him to win.
@@
 
-A different heuristic would be the distance of your most advanced pawn minus your opponents most advanced pawn, this heuristic could give you insight into how close you are to winning the game or how close your opponent is.
-As a higher-level concept, we can call this concept your aggressiveness, as it closely resembles how aggressive you are going for in for the win.
-Generally, in Breakthrough, it is favorable to move your whole team as a unit and play more defensively, and additional heuristics for Breakthrough will be discussed in detail later in this research.
+A different heuristic would be the distance of your most advanced pawn minus your opponent's most advanced pawn; this heuristic could give you insight into how close you are to winning the game or how close your opponent is.
+As a higher-level concept, we can call this concept your aggressiveness, as it closely resembles how aggressive you are going for the win.
+Generally, in Breakthrough, it is favorable to move your whole team as a unit and play more defensively.
+Additional heuristics for Breakthrough will be discussed in detail later in this research.
 
@@
 
-Traditionally methods for playing games were searching through the environment using a heuristic to guide the search.
-A simple way of doing a heuristic-based search would be to give all states a $0$ as a score and give states where you win a positive score.
-We say that that search algorithm is not guided, and the algorithm will probably have to evaluate every state in the state-space.
-This method of searching is generally extremely inefficient as the state-spaces of game environments are often extremely large, and sometimes infinite.
-For instance, an upper-bound estimate of the state-space for breakthrough is $3^{(M-1)*(N-1)}+2*N$ where $M$ is the height of the board $N$ is the width of the board.
+Traditionally, methods for playing games search through the environment using a heuristic to guide the search.
+A simple way of doing a heuristic-based search would be to give all non-terminal states a $0$ score and terminal states positive or negative scores based on whether it is a inw or a loss, repesctively.
+We say that that a search algorithm is not guided, and the algorithm will probably have to evaluate a large portion of the state-space.
+This method of searching is generally extremely inefficient as the state-spaces of game environments are often extremely large, even infinite.
+For instance, an upper-bound estimate of the state-space for Breakthrough is $3^{(M-1)*(N-1)}+2*N$ where $M$ is the height of the board $N$ is the width of the board.
 The $3^{(M-1)*(N-1)}$ represents each position of the board having either a white, or black piece, or being empty.
+And, the $2*N$ component represents each square the final piece to move could have moved to.
 So for a small board, $5x4$ the upper-bound estimation of the state-space is $531,449$ states.
@@
 In the algorithm Monte-Carlo Tree Search described by R.
-Coulom \cite{mcts:coulom}, there is an agent within some environment.
+Coulom\cite{mcts:coulom}, there is an agent within some environment.
 Where each node in the environment represents a state-action pair of the environment, by state-action pair what is meant it is some state and the action that brought the agent to that state.
@@
 ML is a sub-field of Artificial Intelligence.
-ML algorithms come in the form of two classes \textit{Classification}, where the algorithm should find a class representing the data it is given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.
+ML algorithms try to achieve one of two classes, \textit{Classification}, where the algorithm should find a class representing the data it is given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.
 
@@
 
-In RL the algorithm is given only some input, and then the algorithm tries some outcome generally actions in some environment.
+% In RL the algorithm is given only some input, and then the algorithm tries some outcome generally actions in some environment.
 Then over an episode~\footnote{a series of outcomes / a timespan} once the episode is finished some reward is given.
@@
 
-In supervised learning, the data and their corresponding outcomes are already available to us. Examples of the data we can use in supervised learning would be labeled images.
+In Supervised Learning, the ML algorithms attempt to build a model from a data set of labelled examples.
+The labelled examples are a set of input values and their corresponding output value.
+The ML algorithm then uses this data set to construct a model that is as accurate as possible at outputting the correct output value given the input value.
+In supervised learning many techniques are applied to maintain the generality of the model s.t.
+it does not just represent the data set it is given but also has high accuracy on a possible future data set it has not yet encountered.
+
+Some examples of algorithms that are popular for Supervised learning would be, Decision Trees, Support Vector Machines, or Naive Bayes.
 
@@
 
-Reinforcement Learning is where a model learns from experience.
-That is, the notion of input/output values changes, the model uses itself to generate input values by acting in an environment.
-The environment then returns some result, that could be losing in a game, making a correct prediction, or any number of outcomes.
-The result is then propagated through the model allowing it to improve with this new information.
-% When doing moving it's parameters in the direction of the recieved output
-% from the environment.
+Reinforcement Learning focuses on the idea of trial and error for an ML algorithm, where the algorithm directly interacts with the environment it operates in, and from operating in the environment it is provided with either positive or negative feedback for it's actions.
+Typically, the environment needs to be modeled as a Markov Decision Process.
+That is, the selection of actions in a state requires only the knowledge of being in that state, not the actions it took to get to that state.
 
 Many algorithms are popular in reinforcement learning, for instance Q-learning\cite{qlearning:watkins}, and many others.
+
+\subsection{Unsupervised Learning}
+
+In Unsupervised Learning, the machine learning algorithms attempt to build a model from a data set of values that don't have a corresponding output value.
+These algorithms then generally attempt to find pattern within the data set, to which we could then later label upon examination of the patterns.
+Importantly, in Unsupervised Learning, all columns of values in the data set should be normalized to the same range, and should be standardized s.t.
+the mean of the values is $0$ and it's standard deviation is $1$.
+This is done in order for one value not to dominate the patterns in the data set.
+
+Common algorithms in Unsupervised Learning are, K-Nearest Neighbours (KNN), K-Means\cite{lloyd:kmeans}, and DBScan\cite{ling:dbscan}.
 
@@
 There we have neurons that when they get some input, generally a vector of numbers.
-The neuron takes the sum of that vector, weighs the sum by a constant, then applies a non-linear activation function to it.
+The neuron takes the sum of that vector, weighs the sum by a constant, then applies an activation function to it.
 The result of doing this is then passed on to the next neuron.
@@
 It can then be said that a neural network is doing a function approximation of the input to some value.
-$f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.
+And, would be mathematically stated as $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.
 
@@
 
-Model Interpretability is the more common approach of XAI, as it is
+Model Interpretability is the more common approach of XAI, mainly because it is less constrained than model explainability.
+In order to achieve model interpretability, we must be able to answer the question of what prediction will the model return on a given input, with a high accuracy.
+Simple examples would be, given a trained neural network and a picture of a dog, if that model is highly interpretable, we can say with high certainty that the predicted value will be dog.
 
-\subsubsection{Saliency Maps}
+\subsection{Model Explainability}
+
+Model explainability within the context of neural networks isn't possible today. Model
+explainability referrs to firstly considering some input and output from a model. Then
+afterwards the model is examined to determine exactly what led to the predicted output.
+This concept is simple when we're working with Decision Trees. A decision tree is a tree
+whose nodes are representative of an input value and at every node a branch is selected
+based on the value of the input value. It is therefore easy to see how to examine the tree
+to explain the output, by following the branches in the tree we can exactly explain why the model predicted the output.
+
+When we talk about neural networks this process is much more difficult, the underlying nodes
+are generally in the millions, the different layers of the neural network vary in the operations they apply to the input.
+During this process, the value is modified such that becomes far removed from the initial input value.
+That being said, while the possibility of completely monitoring the training process and completely monitoring the evaluation process is truly possible it is not feasible.
+
+\subsection{Saliency Maps}
 
 Within XAI many methods have been developed to try to evaluate
-ANN's, these methods are often referred as model interpretability methods.
-In the field image recognition there has been a lot of work examining which
-pixels of an image the model deems important. Arguably the most popular
-method for this is Saliency Maps\cite{Koch:saliency}. There the pixel values the model
+ANN's. In the field image recognition there has been a lot of work examining which
+pixels of an image the model deems important. One such method is, Saliency Maps\cite{Koch:saliency}. There the pixel values the model
 deems important are colored in s.t. a human can examine the image and get a
@@
 a classification of a dog can be seen in Figure \ref{fig:dog_saliency}.
+This method is understandable to a human when the saliency map lines up to what we would focus on.
+However, this achieves no explanation on the prediction if the saliency map doesn't line up to human intuition or the prediction, for example, if in the same Figure\ref{fig:dog_saliency} we had the same saliency map but the prediction would be dragon, or if the salience map was on the trees and the prediction was dog.
 
@@
 
-While this method is understandable in the context of image recognition it
-lacks severely when your input is not an image.
-
-\subsubsection{Shapley Values}
+\subsection{Shapley Values}
 
@@
 
-\subsubsection{Concept Activation Vectors}
+\subsection{Concept Activation Vectors}
 
-A recent paper by Been Kim Et. al \cite{Keem:TCAV}, shows a method for examining
+A recent paper by Been Kim et al.\cite{Keem:TCAV}, shows a method for examining
 a neural network giving a much more human insight into a prediction. Using Concept
@@
 The construction of a CAV requires a method of labelling the values in your dataset with
-the corresponding concept in order to create a simple binary classifier on the data points.
-
-\subsection{Model Explainability}
-
-Model explainability within the context of neural networks isn't possible today. Model
-explainability referrs to firstly considering some input and output from a model. Then
-afterwards the model is examined to determine exactly what led to the predicted output.
-This concept is simple when we're working with Decision Trees. A decision tree is a tree
-whose nodes are representative of an input value and at every node a branch is selected
-based on the value of the input value. It is therefore easy to see how to examine the tree
-to explain the output. We just follow the branches in the tree.
-
-When we talk about neural networks this process is much more difficult, the underlying nodes
-are generally in the thousands, the different layers of the neural network varies in the operations
-it applies to the input value and such while travelling through the neural network the modified
-value becomes far removed from the initial input value to the eyes of the reader. That being said,
-while the possibility of completely monitoring the training process and completely monitoring the
-evaluation process is truly possible it is not feasible. And secondly the process of seeing an
-input and it's corresponding output will not be of any value if one were to consider the process
-of prediction.
+the corresponding concept in order to create a binary classifier on data.
+The binary classifier is constructed on the internal representation of the data points within the neural network.
+After training the neural network, and constructing the classifier, when we run a new datapoint through the neural network, and we examine the directional derivative of that datapoint.
+If the direction is in the direction of the binary classifier we say that the datapoint contains the concept.

### 2020-12-09 latex/sections/introduction.tex
latex updates

@@
 
-The world of deep learning (DL) is exciting, we have models that can examine
-images and very reliably be able to identify it's content. Deep learning models have
-beaten Ophthalmologists in identifying diabetic retinopathy, they've identified
+The world of deep learning (DL) is exciting. We have models that can examine
+images and reliably identify their content. Deep learning models have
+outperformed Ophthalmologists in identifying diabetic retinopathy, and identified
 cancer cells where others have not. Deep learning models are used for the bleeding
-edge of protein folding, to gain further understanding of the underlying structure of organisms, such as, viruses. \cite{deepmind:alphafold}
+edge of protein folding, to gain further understanding of the underlying structure of organisms, such as, viruses.\cite{deepmind:alphafold}
 These models have done these things extremely accurately, cheap, and fast.
 
-DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search was used to defeat the standing world champion Lee Sedol (CITE ALPHAGO ZERO) in the incredibly expansive game of Go.
+DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search was used to defeat the standing world champion Lee Sedol\cite{silver:alphagozero} in the incredibly expansive game of Go.
 This was done in the year 2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself.
@@
 is still new. This is the field of Explainable Artificial Intelligence (XAI).
-If we wish to continue using artificial intelligence (AI) and machine learning (ML),
-to improve our lives we must examine how they work, this is not only to improve
-the models themselves but could also augment our ability in many cases.
+If we wish to continue using artificial intelligence (AI) and machine learning (ML)
+to improve our lives we must examine how they work, not only to improve
+the models themselves but also to augment our own ability in many cases.
 Furthermore, these explanations are a legal requirement now in many areas\cite{legalexplanation:goodman}, namely, legal, and
@@
 \begin{enumerate}
-  \item Given a NN that plays Breakthrough very well, can we evaluate if that NN recognizes the
+      \item Given a NN that plays Breakthrough, can we evaluate if that NN recognizes the
             HLCs that human players use.

### 2020-12-16 latex/sections/abstract/english.tex
yngvi+james comments fixed

@@
-this is my abstract written in the language of English I am testing thought
-
-alas  I knew
-
-ok so now I'm trying something new
-
-asdf
-
-test
+Deep neural networks have exceeded expectations in many areas, they can generate powerful and accurate predictions of the real world.
+This is going to be a powerful tool for the future, and we require a method of examining them to understand why and how they come to a specific prediction to alleviate trust issues.
+This paper examines the strategies of a deep neural network that plays the game Breakthrough uses.
+We explain the process of creating a neural network that can achieve success in playing board games only from playing against itself, without any human input.
+We outline the process of creating a model to examine the internal representation of the neural network of the game state to evaluate whether it both recognizes and uses strategies applied by humans that play the game.
+These strategies are compared against each other to infer which is best, and which the neural network uses the most.
+We show that the neural network emphasizes strategies that are considered the most successful handcrafted strategies for the game Breakthrough.
+Lastly, we examine a trained neural network with a suite of simple to complex strategies to understand which is used most.

### 2020-12-16 latex/sections/abstract/icelandic.tex
yngvi+james comments fixed

@@
-this is my abstract written in the language of íslensku þæööasdf
+Djúptauganet hafa farið fram úr væntingum á mörgum sviðum, þau geta framkallað öflugar og nákvæmar spár um raunvöruleikann.
+Þetta mun verða öflugt tól í framtíðinni og við munum þurfa aðferðir til að geta rýnt í þessar spár til að styrkja traust okkar á þeim.
+Þessi ritgerð skoðar hvaða herkænsku aðferðir djúptauganet sem spilar leikinn Breakthrough nýtir sér til að taka ákvarðanir.
+Við útskýrum ferlið að smíða tauganet sem getur náð árangri í borðspilum með því einungis að spila gegn sjálfu sér, án inngrips frá mönnum.
+Við útskýrum ferlið að búa til líkan til að skoða innri framsetningu tauganets á leikstöðu til að meta hvort tauganetið þekki eða nýtir sér aðferðir sem menn hafa notað til að spila leikinn vel.
+Þessar aðferðir eru bornar saman til að álykta hver sé best og hver tauganetið notar mest.
+Við sýnum að tauganetið leggur áherslu á aðferðir sem eru taldar bestar fyrir leikinn Breakthrough.
+Að lokum skoðum við þjálfað tauganet með hóp af auðveldum sem og flóknum aðferðum til að rýna í hver herkænsku aðferð er mest notuð af tauganetinu.

### 2020-12-16 latex/sections/background.tex
yngvi+james comments fixed

@@
 
-In this chapter, we discuss traditional artificial intelligence in the context of search, the algorithms we used, what some other similar methods exist as well as the field in general.
-We allocate a large portion of the section to Breakthrough as that will be the testbed for future sections.
-We discuss the different classes of machine learning as well as neural networks in more depth.
+In this chapter, we discuss traditional artificial intelligence in the context of search, and explain the algorithms we used.
+Additionally, we discuss similar methods for searching state-spaces.
+We allocate a large portion of the section to Breakthrough as that will be our testbed.
+We discuss the different classes of machine learning as well as neural networks.
 
@@
 
-When researching Artificial Intelligence it is important to select an environment that is a suitable abstraction for the task at hand.
-Environments vary significantly and can be identified by their characteristic.
-The characteristics that are generally used to describe environments can be seen in Table~\ref{tab:env_characteristics}.
-This categorization of environments is described in the book Artificial Intelligence by Norvig \& Russell.
-\cite{Russell:AIModern}
+When researching artificial intelligence it is important to select an environment that is a suitable abstraction for the task at hand.
+Environments vary significantly and can be identified by their characteristics.
+The characteristics that are commonly used to describe environments can be seen in Table~\ref{tab:env_characteristics}\cite{Russell:AIModern}.
 
@@
 	\centering
+	\caption{Characteristics of environments}
 	\begin{tabular}{|c|c|p{6cm}|}
@@
 		\hline
-    Deterministic           & Deterministic, Stochastic    & Do the actions your agent do deterministicly impact the environment.                 \\
+		Deterministic           & Deterministic, Stochastic    & Do the actions your agents execute deterministicly impact the environment.                 \\
 		\hline
@@
 		\hline
-    Static                  & Static, Semi-Static, Dynamic & Does the environment without agent input, or does it wait until agents take actions. \\
+		Static                  & Static, Semi-Static, Dynamic & Does the environment change without agent input, or does it wait until agents take actions. \\
 		\hline
@@
 	\end{tabular}
-  \caption{Characteristics of environments}
 	\label{tab:env_characteristics}
@@
 
-Categorizing environments like this gives you the power to find an environment in which a method works and know it can be applied to different environments with the same characteristics.
+Categorizing environments using Table \ref{tab:env_characteristics} gives one the power to find an environment in which a method works and know it can be applied to different environments with the same characteristics.
 Additionally, it allowes us to talk about agents in the context of environments as the entities that act within the environment.
@@
 
-Classical Artificial Intelligence Game Environments are commonly used to validate a method, e.g game environments can be games like Tic-Tac-Toe, Breakthrough, and driving simulators.
+Classical artificial intelligence game environments are commonly used to validate a method, e.g game environments can be games like Tic-Tac-Toe or Breakthrough, or simluation environments like driving simulators.
 Game environments are a suitable place to apply AI as they serve as an abstraction of the real world, for instance, a self-driving car agent who is verified to avoid driving into walls in a simulation is possibly safer than one who is not.
@@
 
-The game Breakthrough is a simplified version of chess; the game is set up on a $MxN$ board with squares like in chess, and each player starts with two rows of pawns at opposite ends.
+The game Breakthrough is a simplified version of chess; the game is set up on a $M\times N$ board with squares like in chess, and each player starts with two rows of pawns at opposite ends.
 The objective of the game is for a player to move one of their pawns to the opposite end of the board.
 A player wins if either they have reached the opposite end of the board or have captured all of his opponents' pawns.
-The pawns differ from chess pawns in such a way that they can not move two squares on the first move and, they can move diagonally as well as forward.
-This leads to the game being impossible to draw as pieces are always able to move.
-An example of an initial board in Breakthrough can be seen in Figure~\ref{fig:initbtboard}
+The pawns differ from chess pawns in such a way that they can not move two squares on the first move and they can both move and capture diagonally.
+This leads to the game being impossible to draw as pieces are always able to move or capture.
+An example of an initial board in Breakthrough can be seen in Figure~\ref{fig:initbtboard}.
 
@@
 
-  \caption{Initial breakthrough board}
+	\caption{Initial Breakthrough board}
 	\label{fig:initbtboard}
@@
 	\centering
+	\caption{Categorization of Breakthrough}
 	\begin{tabular}{|c|c|}\hline
 		\textbf{Characteristic} & \textbf{Value} \\\hline
-    Observable              & Fully          \\
-    Agents                  & Multi          \\
-    Deterministic           & Deterministic  \\
-    Episodic                & Sequential     \\
-    Static                  & Static         \\
+		Observable              & Fully          \\\hline
+		Agents                  & Multi          \\\hline
+		Deterministic           & Deterministic  \\\hline
+		Episodic                & Sequential     \\\hline
+		Static                  & Static         \\\hline
 		Discrete                & Discrete       \\\hline
 	\end{tabular}
-  \caption{Categorization of Breakthrough}
 	\label{tab:breakthrough_cat}
@@
 
-Categorizing Breakthrough with the characteristics described in Section~\ref{sec:environments}.
-We end up with the description shown in Table~\ref{tab:breakthrough_cat}.
+Categorizing Breakthrough with the characteristics described in Section~\ref{sec:environments} we end up with the description shown in Table~\ref{tab:breakthrough_cat}.
 These characteristics are identical to that of Tic-Tac-Toe, and chess.
-This categorization is the most common in board games where two-player compete.
+This categorization is the most common in board games where two players compete.
 
@@
 
-To evaluate the game of Breakthrough we can consider many heuristics (higher-level concepts) for instance a very simple heuristic would be a players material advantage.
+To evaluate the game of Breakthrough we may consider many heuristics (higher-level concepts), for instance a very simple heuristic would be a players material advantage.
 \textit{Material Advantage} is the amount of pieces the player has minus the amount of pieces the opponent has.
-This heuristics gives us some insight into how well the game is progressing, but obviously, there are cases where this doesn't tell us much, as in cases when your opponent has a single piece left that is on the row immediately before the row needed for him to win.
+This heuristics gives us some insight into how well the game is progressing, but obviously, there are cases where this does not tell us much, for example when your opponent has a single piece left that is on the row immediately before the row needed for them to win.
 No matter how many pieces you have left, this state is bad for you if you're not able to capture that piece.
@@
 
-A different heuristic would be the distance of your most advanced pawn minus your opponent's most advanced pawn; this heuristic could give you insight into how close you are to winning the game or how close your opponent is.
-As a higher-level concept, we can call this concept your aggressiveness, as it closely resembles how aggressive you are going for the win.
-Generally, in Breakthrough, it is favorable to move your whole team as a unit and play more defensively.
-Additional heuristics for Breakthrough will be discussed in detail later in this research.
+A different heuristic would be the distance of your most advanced pawn minus your opponent's most advanced pawn; this could give you insight into how close you are to winning the game or how close your opponent is.
+As a higher-level concept, we can call this concept your \textit{aggressiveness}, as it closely resembles how aggressive you are going for the win.
+Generally, in Breakthrough, it is favorable to move your whole team as a unit and play more defensively\cite{lorentz:heuristic}.
+More advanced heuristics for Breakthrough will be discussed in detail later in a later chapter.
 
@@
 Traditionally, methods for playing games search through the environment using a heuristic to guide the search.
-A simple way of doing a heuristic-based search would be to give all non-terminal states a $0$ score and terminal states positive or negative scores based on whether it is a inw or a loss, repesctively.
+A simple way of doing a heuristic-based search would be to give all non-terminal states a $0$ score and terminal states positive or negative scores based on whether it is a win or a loss, repesctively.
 We say that that a search algorithm is not guided, and the algorithm will probably have to evaluate a large portion of the state-space.
@@
 And, the $2*N$ component represents each square the final piece to move could have moved to.
-So for a small board, $5x4$ the upper-bound estimation of the state-space is $531,449$ states.
+So for a small board, $5\times 4$ the upper-bound estimation of the state-space is $531{,}449$ states.
 
-This is why a good heuristic is very valuable because we can disregard all following states that result from doing a move some in a previous state as they will only lead to worse outcomes.
+This is why a good heuristic is needed because we can identify paths in the game tree that will lead to a bad outcome early, allowing us to disregard a significant amount of states, reducing the time required to search.
 
-The algorithms that are used in traditional state-space searches are for instance Depth-first search (DFS), Breadth-first Search (BFS), Alpha-Beta Pruning Search (AB-Search)\cite{abpruning:dj}, and Monte-Carlo Tree Search (MCTS).
+The algorithms that are used in traditional state-space search are for instance Depth-first search (DFS), Breadth-first Search (BFS), Alpha-Beta Search (AB-Search)\cite{abpruning:dj}, and Monte-Carlo Tree Search (MCTS).
 
-More modernly, these search methods have been amplified by Machine Learning, in such a way that we do not need to figure out a good heuristic for a given state, but rather, we apply a machine learning model to learn a function that takes in a state and returns an evaluation of that state.\cite{neuralnetworksgames:michulke} This can lead to a significant time reduction as we do not need to simulate a whole game from a state to receive its evaluation we rather receive the evaluation from the model.
+More modernly, these search methods have been augmented by Machine Learning, in such a way that we do not need to figure out a good heuristic for a given state, but rather, we apply a machine learning model to learn a function that takes in a state and returns an evaluation of that state\cite{neuralnetworksgames:michulke}.
+This can lead to a significant time reduction as we do not need to simulate a whole game from a state to receive its evaluation but rather receive the evaluation from the model.
 
@@
 
-In the algorithm Monte-Carlo Tree Search described by R.
-Coulom\cite{mcts:coulom}, there is an agent within some environment.
-Where each node in the environment represents a state-action pair of the environment, by state-action pair what is meant it is some state and the action that brought the agent to that state.
-This pair should be unique within the environment.
+In the algorithm Monte-Carlo Tree Search, described by R.
+Coulom\cite{mcts:coulom}, there is an agent within an environment.
+Each node in the environment represents a state of the environment, and each edge from the node represents a possible action from that state $s$ to a child state $s'$.
+Often, these concepts are considered state-action pairs, which should be unique within the environment.
+And as such, they create a tree where we are able to search the state-space.
 
-MCTS is a method of exploring an environment in a randomized manner (Monte Carlo is the term implying randomness).
-In MCTS there are four stages.
-Selection, Expansion, Simulation, and Back-propagation.
+MCTS is a method of exploring an environment in a best-first manner.
+In MCTS there are four stages: Selection, Expansion, Simulation, and Back-propagation.
 They happen sequentially and repeatedly.
 MCTS is initialized with a tree consisting of the unexpanded initial state of the environment.
-
-In MCTS there is a tree representing the game-environment.
-This tree consists of nodes $n_i$ where $i$ represents the point in time of the node, for example, $N_0$ in chess is the initial position and $N_x$ is some position in the middle of the game and $N_e$ is one of the states representing a position where there is either a draw or one player has won the match.
+This tree consists of nodes $n_i$ where $i$ represents the point in time of game state, for example, $n_0$ in chess is the initial position and $n_x$ is some position in the middle of the game and $n_e$ is one of the states representing a position where there is either a draw or one player has won the match.
 Each of the nodes has $4$ values, $s$, $a$, $Q$, and $N$.
@@
 
-The MCTS algorithms four phases
-\begin{enumerate}
-  \item Selection
-  \item Expansion
-  \item Simulation/Rollout
-  \item Backpropagation
-\end{enumerate}
+We describe the MCTS algorithm's four phases: Selection, Expansion, Simulation/Rollout, Backpropagation, in more detail below.
 
 \begin{equation} \label{UCT_formula}
-  \text{Child UCT value} = \frac{Q_{(s',a')}}{N_{(s',a')}} + c_{uct} * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
+	\text{Child UCT value} = Q_{(s',a')} + c_{uct} * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
 \end{equation}
@@
 \subsection*{Selection}
-During the selection phase, a node $(s,a)$ within the tree which has not yet been expanded is found.
-This process uses Upper Confidence Bound on Trees (UCT) to find that node $(s,a)$, the formula is described in Equation \ref{UCT_formula}. For a parent node $(s,a)$
-(initially the root of the tree) we select the child with the highest UCT value. Repeatedly until an unexpanded
+During the selection phase, a node $s$ within the tree which has not yet been expanded is found.
+This process uses Upper Confidence Bound on Trees (UCT)\cite{kocsis:uct} to find that node $s$, the formula is described in Equation \ref{UCT_formula}. For a parent node $s_{parent}$
+(initially the root of the tree) we select the child with the highest UCT value repeatedly until an unexpanded
 node is found. This process is done to balance the amount of exploration vs exploitation of nodes in the
@@
 \subsection*{Expansion}
-Then the expansion phase expands the node generating all of $(s,a)$'s children, $(s',a')$ are generated by applying all actions $a'$ in $(s,a)$.
+Then the expansion phase expands the node generating all $s'$ nodes, the nodes are generated by applying all actions $a$ in $s$.
 
 \subsection*{Rollout}
-Next during rollout, actions from $(s,a)$ are randomly selected to move to $(s',a')$, then repeated to go to $(s'',a'')$, until a terminal node within the environment is reached.
+Next during rollout, actions $a$ from $s$ are randomly selected to move to $s'$, then repeated to go to $s''$, until a terminal node within the environment is reached.
 By terminal, we mean a state in which the game is finished.
@@
 
+\subsection{Neural Network based UCT}
+
 When training a neural network the UCT formula is modified slightly to prefer selecting nodes
 that the neural network values highly by introducing a second scalar to the formula $f((s,a)) = (p,v)$, where $f$ is the neural network, $p$ is the policy vector returned by the neural network and $v$ is the predicted value from the neural network.
-The resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT.
-Secondly, the backpropagation process
-is modified to instead of doing rollout/simulation to receive a reward the predicted value $v$ from the neural network is used instead.
+The resulting formula is described in Equation \ref{PUCT_formula}, and is called PUCT\cite{rosin:puct}.
+Secondly, the backpropagation process is modified to instead of doing rollout/simulation to receive a reward the predicted value $v$ from the neural network is used.
 
 \begin{equation} \label{PUCT_formula}
-  \text{Child PUCT value} = \frac{Q_{(s',a')}}{N_{(s',a')}} + c_{uct} * p_{(s,a)} * \frac{\sqrt{log(N_{(s,a)})}}{N_{(s',a')}}
+	\text{Child PUCT value} = Q(s,a) + c_{puct} * p(s,a) * \frac{\sqrt{\sum_b{N(s,b)}}}{1+N(s,a)}
 \end{equation}
@@
 
-Machine Learning (ML) is a research field in which machines apply statistical functions on data to achieve a correct output, by \textit{correct} we mean the corresponding result which we expect.
+Machine Learning (ML) is a research field in which machines apply statistical functions on data to achieve a correct output, where by \textit{correct} we mean the corresponding result that we expect.
 Generally, this a repetitive process where we look at examples of the data, and the algorithm progressively gets closer to the underlying function of the data it is fed.
@@
 ML is a sub-field of Artificial Intelligence.
-ML algorithms try to achieve one of two classes, \textit{Classification}, where the algorithm should find a class representing the data it is given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.
 
-Typically ML can be viewed in three different groups, Supervised Learning, Reinforcement Learning (RL), and Unsupervised Learning.
-Where in Supervised Learning, the algorithm is given data examples and their corresponding outcome.
-For example, a supervised learning algorithm could be provided with data regarding the weather and the corresponding temperature, the algorithm should then find a pattern within the weather data and find the continuous function represented by the data.
-This would then be an example of a regression task.
-Flipping thing example around, if the algorithm would just be provided the temperature and it should tell us whether it is sunny outside or not, that would be a classification task.
-
-% In RL the algorithm is given only some input, and then the algorithm tries some outcome generally actions in some environment.
-Then over an episode~\footnote{a series of outcomes / a timespan} once the episode is finished some reward is given.
-The algorithm will then learn whether the actions were good actions from the reward.
-Examples of this are agents playing a game like Flappy Bird, where the data they are given is the state of the game, and they try to either jump or not jump.
-
-%Machine learning (ML) focuses on the using of data and a corresponding outcome w.r.t that data to 
-%extrapolate some underlying function of that data. Examples of how we use machine learning is for instance the ability to predicting the 
-%rise and fall of some stock, predicting what the weather will be in a week, and whether an image 
-%is an image of a dog or a cat. Many data structures and algorithms are used to achieve this 
-%goal, but recently the field of ANN/DNN's has been the standard for achieving the best
-%results.
-
-%As a general notion we can split machine learning into three different sects those are Supervised Learning, Reinforcement Learning, and Unsupervised Learning.
+We can categorize machine learning into three different sub-categories: Supervised Learning, Reinforcement Learning, and Unsupervised Learning.
 
@@
 
-Reinforcement Learning focuses on the idea of trial and error for an ML algorithm, where the algorithm directly interacts with the environment it operates in, and from operating in the environment it is provided with either positive or negative feedback for it's actions.
+Reinforcement Learning focuses on the idea of trial and error for an ML algorithm, where the algorithm directly interacts with the environment it operates in, and from operating in the environment it is provided with either positive or negative feedback for its actions.
 Typically, the environment needs to be modeled as a Markov Decision Process.
@@
 
-Many algorithms are popular in reinforcement learning, for instance Q-learning\cite{qlearning:watkins}, and many others.
+Many algorithms are popular in reinforcement learning, for instance Q-learning\cite{qlearning:watkins}, and TD-Lambda\cite{sutton:tdlambda}.
 
@@
 
-In Unsupervised Learning, the machine learning algorithms attempt to build a model from a data set of values that don't have a corresponding output value.
+In Unsupervised Learning, the machine learning algorithms attempt to build a model from a data set of values that do not have a corresponding output value.
 These algorithms then generally attempt to find pattern within the data set, to which we could then later label upon examination of the patterns.
@@
 
-Neural networks (NN) are popular methods within a sub-field of ML which is called Deep Learning (DL).
+Neural networks (NN) are popular methods within a sub-field of ML known as Deep Learning (DL).
+Neural networks are generally used on a labelled dataset.
 NN's are created to resemble how the human brain functions.
-In the brain, we have neurons which when they get a signal they apply some function to them and if the resulting signal is high enough, they fire to the next neuron.
+In the brain, we have neurons which when they receive a signal they apply some function to them and if the resulting signal is high enough, they fire to the next neuron.
 This is how it is done in the neural network model as well.
-There we have neurons that when they get some input, generally a vector of numbers.
-The neuron takes the sum of that vector, weighs the sum by a constant, then applies an activation function to it.
+In NNs we have a layer of neurons $f_i$, initial layer being $f_0$ and the last layer being $f_n$.
+These layers behave in such a way that when they get some input, generally a vector of numbers, each neuron in the layer takes the sum of that vector, weighs the sum by a learned weighing factor, then applies an activation function to it.
 The result of doing this is then passed on to the next neuron.
 Until a final layer of neurons is reached.
-At that point, we have a value that the neural network corresponds to the input value.
-This value can be a binary classification (cat or dog image), a regression value (the value of a property), or any number of outputs.
+At that point, we have reached a layer that corresponds to the prediction of the network.
+This value can be a binary classification (cat or dog image), a regression value (the value of a property), or even the neural networks representation of the input, e.g where a neural network attempts to create an image from an example image\cite{karras:stylegan}.
 It can then be said that a neural network is doing a function approximation of the input to some value.
@@
 
+The core of the learning in neural networks comes from optimizing the weighing factor.
+That weighting factor is gradually made to move in the direction of the actual correct output by algorithms, e.g the backpropagation algorithm\cite{rumelhard:backprop}.
+
+%\section{Traditional Learning in Board Games}
+
 \section{Explainable Artificial Intelligence (XAI)}
 
-The field of XAI research is still very far behind its counterpart AI research, within XAI two fields are the largest, those are Model Interpretability and Model Explainability.
+The field of XAI research is still far behind its counterpart AI research\cite{tjoa:surveyxai}. Within XAI two fields are the largest: Model Interpretability and Model Explainability.
 
@@
 Model Interpretability is the more common approach of XAI, mainly because it is less constrained than model explainability.
-In order to achieve model interpretability, we must be able to answer the question of what prediction will the model return on a given input, with a high accuracy.
+In order to achieve model interpretability, we must be able to answer the question of what prediction the model will return on a given input, with a high accuracy.
 Simple examples would be, given a trained neural network and a picture of a dog, if that model is highly interpretable, we can say with high certainty that the predicted value will be dog.
@@
 
-Model explainability within the context of neural networks isn't possible today. Model
-explainability referrs to firstly considering some input and output from a model. Then
+Model explainability within the context of neural networks is not possible today. Model
+explainability refers to firstly considering some input and output from a model. Then
 afterwards the model is examined to determine exactly what led to the predicted output.
@@
 based on the value of the input value. It is therefore easy to see how to examine the tree
-to explain the output, by following the branches in the tree we can exactly explain why the model predicted the output.
+to explain the output. By following the branches in the tree we can exactly explain why the model predicted the output.
 
-When we talk about neural networks this process is much more difficult, the underlying nodes
-are generally in the millions, the different layers of the neural network vary in the operations they apply to the input.
-During this process, the value is modified such that becomes far removed from the initial input value.
-That being said, while the possibility of completely monitoring the training process and completely monitoring the evaluation process is truly possible it is not feasible.
+When we talk about neural networks this process is much more difficult; the underlying nodes
+are generally in the millions and the different layers of the neural network vary in the operations they apply to the input.
+During this process, the value is modified such that it becomes far removed from the initial input value.
+That being said, while the possibility of completely monitoring the training process and completely monitoring the evaluation process is possible it is not feasible.
 
@@
 
-Within XAI many methods have been developed to try to evaluate
-ANN's. In the field image recognition there has been a lot of work examining which
-pixels of an image the model deems important. One such method is, Saliency Maps\cite{Koch:saliency}. There the pixel values the model
-deems important are colored in s.t. a human can examine the image and get a
-sense of what portions of the image are important to the model, an example of
+Within XAI many methods have been developed to try to explain
+ANN's. In the field of image recognition there has been a lot of work examining which
+pixels of an image the model deems important. One such method is Saliency Maps\cite{Koch:saliency}. There the pixel values the model
+deems important are colored s.t. a human can examine the image and get a
+sense of what portions of the image are important to the model. An example of
 a classification of a dog can be seen in Figure \ref{fig:dog_saliency}.
 This method is understandable to a human when the saliency map lines up to what we would focus on.
-However, this achieves no explanation on the prediction if the saliency map doesn't line up to human intuition or the prediction, for example, if in the same Figure\ref{fig:dog_saliency} we had the same saliency map but the prediction would be dragon, or if the salience map was on the trees and the prediction was dog.
+However, this achieves no explanation on the prediction if the saliency map doesn't line up to human intuition or the prediction, for example, if in the same Figure \ref{fig:dog_saliency} we had the same saliency map but the prediction would be dragon, or if the salience map was on the trees and the prediction was dog.
 
@@
 	\includegraphics[width=.5\textwidth]{graphics/dog_saliency}
-  \caption{Example of a models saliency map for an image of a dog}
+	\caption{Example of a model's saliency map for an image of a dog}
 	\label{fig:dog_saliency}
@@
 
-Methods for explaining models that aren't image recognition models include
-Shapley values, from the Lundberg \& Lee\cite{LundbergL:shapley}. There the input is examined against it's output, then iteratively
+Methods for explaining models that are not image recognition models include
+Shapley values, from Lundberg \& Lee\cite{LundbergL:shapley}. There the input is examined against its output, then iteratively
 input values are selected to be fixed. Then the other input values are varied and
-an average change in prediction is calculated. With this the shapley value can be
+an average change in prediction is calculated. With this the Shapley value can be
 estimated for the fixed input value. This is done to examine which input values have
 the strongest link to the output value. Shapley values on a dataset can give insights
-on which input values the model deems important.
+on which input values the model deems most important.
 
@@
 A recent paper by Been Kim et al.\cite{Keem:TCAV}, shows a method for examining
-a neural network giving a much more human insight into a prediction. Using Concept
+a neural network giving a much more humanlike insight into a prediction. Using Concept
 Activation Vectors (CAV) a directional derivative for a given input can be examined
@@
 concept, and if we're able to gather if a nerual network uses this strategy for prediction
-we have a deeper understanding of its underlying structure. Leading to an explanation of
+we have a deeper understanding of its underlying structure, leading to an explanation of
 the result.
@@
 The binary classifier is constructed on the internal representation of the data points within the neural network.
-After training the neural network, and constructing the classifier, when we run a new datapoint through the neural network, and we examine the directional derivative of that datapoint.
+After training the neural network, and constructing the classifier, we run a new datapoint through the neural network and we examine the directional derivative of that datapoint.
 If the direction is in the direction of the binary classifier we say that the datapoint contains the concept.
+
+\section{Summary}
+
+This section discussed some of the important algorithms related to the research done in this thesis, as well as giving a foundation in the readers understanding of the test bed we will be using in future sections.

### 2020-12-16 latex/sections/discussion.tex
latex updates
yngvi+james comments fixed

--- 2019-11-28 latex/sections/discussion.tex
@@
 
-\ifdraft{Here I will discuss the myriad of fields research like this can help
-for instance for health organizations, patients, taxpayers, and pharma}
-hey
+This chapter discusses the results and the future work that could span from this research. Additionally, we conclude the work.
 
@@
 
-\ifdraft{summarize the work}
-hey
+As seen in Chapter \ref{cha:results}, our trained agent does indeed quickly rise to using the higher-level concepts, thus answering our research question \ref{rq:1}.
+The results are promising, as we see a popular concepts like Lorentz-Horey score rise in emphasis for the neural network, and a simple but effective concept like Material Advantage rise slowly but not exessively.
+We find the Unity concept to be disappointing as stated earlier in this paper; maintaining closeness of your pawns tends to be preferential.
+And lastly, for a the concept Aggressiveness, the fact that it falls off early is what we expected.
+It would have been the preferred result if we saw a greater variability in the concepts, but this could be a result from not enough training iterations, or not a complex enough model.
+Considering that the most emphazised HLC is the Lorentz-Horey score, and that the least is Unity, we can answer our research question \ref{rq:2}.
+As we know from the literature, Lorentz-Horey is considered the leading heuristic for Breakthrough, and that HLC is highly emphazised by the neural network.
+Secondly, the Unity heuristic is not a validated heuristic, moreso, an attempt at creating a simple heuristic that closely resembles how tight-packed your pawns are.
+However, the fact that the resulting graphs are relatively steady the results are not as strong as we would have hoped.
+
+Importantly, these concepts are only a few possibilities of an infinite set of concepts, and the neural network could be learning a completely different concept then those that we tested.
+However, that was not the focus of this research, it was to examine if a neural network does move towards HLC's that we humans use in games.
+
+\section{Future Work}
+
+To iterate on this research, training a neural network with greater computing power, or with a larger neural network architecture, will allow the us to hopefully achieve a super-human neural network in Breakthrough.
+This would give greater confidence in the learned concepts, possibly allowing for pedagogical research on the topic regarding which concepts are optimal to train people in the game environment.
+
+The work done in this thesis was only a proof of concept of the process of explaining the internal concept a neural network uses while acting in a board game environment.
+Ideally, one would like to extend this work to a larger set of board games.
+
+Additionally applying this method to a more diverse set of environments would be an interesting field of research.
+We envision a self driving car agent being examined with respect to a higher level concept of aggressiveness, or a mortgage agent being examined with respect to a higher level concept of gender bias.
 
@@
 
-\ifdraft{conclude the work, discuss the significance of the work}
-hey
+The work done in this paper is only a first step in examining working agents in active environments with respect to human level concepts.
+If improved it could have a great impact in our trust on neural network that improve our daily lives.
+While the testbed for the research was a discrete game environment, there is a clear way forward to dynamic systems with discrete human level concepts.
+Our results show that for an agent that clearly doesn't achieve human level performance its actions are often selected with respect to human level concepts.

### 2020-12-16 latex/sections/introduction.tex
yngvi+james comments fixed

@@
 
-The world of deep learning (DL) is exciting. We have models that can examine
-images and reliably identify their content. Deep learning models have
-outperformed Ophthalmologists in identifying diabetic retinopathy, and identified
-cancer cells where others have not. Deep learning models are used for the bleeding
-edge of protein folding, to gain further understanding of the underlying structure of organisms, such as, viruses.\cite{deepmind:alphafold}
-These models have done these things extremely accurately, cheap, and fast.
+The world of deep learning (DL) is exciting.
+We have models that can examine images and reliably identify their content\cite{simonyan:imagerecognition}.
+Deep learning models have outperformed ophthalmologists in identifying diabetic retinopathy\cite{brown:retinopathy} and identified cancer cells where others have not\cite{kourou:cancer}.
+Deep learning models are used for the bleeding edge of protein folding, to gain further understanding of the underlying structure of organisms, such as viruses\cite{deepmind:alphafold}.
+These models have do these things accurately, cheap, and fast
 
-DL in conjunction with a randomized State-space Search Algorithm Monte-Carlo Tree Search was used to defeat the standing world champion Lee Sedol\cite{silver:alphagozero} in the incredibly expansive game of Go.
-This was done in the year 2017, the researchers used reinforcement learning (RL), where the agent learned by only playing against itself.
-These results imply that given enough time to learn the computer agent can gain a deep insight into how the game should be played.
+DL in conjunction with a randomized state-space search algorithm, Monte-Carlo Tree Search, was used to defeat the standing world champion Lee Sedol in the incredibly expansive game of Go\cite{silver:alphagozero}.
+The researchers used reinforcement learning (RL), where the agent learned by only playing against itself.
+These results imply that given enough time to learn the computer agent can gain a deep insight into how a game should be played.
 
-The field of examining deep learning models to gain a richer understanding
-of how they work, and what makes them so much better than humans at various tasks
-is still new. This is the field of Explainable Artificial Intelligence (XAI).
-If we wish to continue using artificial intelligence (AI) and machine learning (ML)
-to improve our lives we must examine how they work, not only to improve
-the models themselves but also to augment our own ability in many cases.
-Furthermore, these explanations are a legal requirement now in many areas\cite{legalexplanation:goodman}, namely, legal, and
-medical. Recently, XAI has generally been focused on Model Interpretability (MI), in
-particular, interpreting the model in such a way that we might predict what the model will
-do given a particular input. Another class of XAI would be Model Explainability (ME), where we can adequately explain how the internal mechanics of the model works.
-A simple example of when we're able to apply ME would be when we print out a decision tree
-that is generated from an ML algorithm, with the decision tree nodes and is conditions we can see exactly what the tree will do with a given input.
-While in MI we would have to predict the results, without going into details as to why.
+The field of examining deep learning models to gain a richer understanding of how they work, and what makes them so much better than humans at various tasks, is still new.
+This is the field of Explainable Artificial Intelligence (XAI).
+If we wish to continue using artificial intelligence (AI) and machine learning (ML) to improve our lives, we must examine how they work.
+Not only to improve the models themselves but also to augment our ability in many cases.
+Furthermore, these explanations are a legal requirement now in many areas\cite{legalexplanation:goodman}, namely, legal and medical.
+Recently, XAI has been focused on Model Interpretability (MI), in particular, interpreting the model in such a way that we can predict what the model will do given a particular input\cite{riberio:recentxai1}.
+Another class of XAI is Model Explainability (ME), where we attempt to adequately explain how the internal mechanics of the model works.
+A simple example of ME would be when we explore a decision tree generated by an ML algorithm.
+The decision tree nodes and conditions explain exactly what the tree will do with a given input.
+This differs from MI where we would have to predict the results, without going into details as to why.
 
-In this thesis, we take a look at how we can examine an artificial neural network (ANN)
-to understand which higher-level concepts (HLC) it deems important for a given state
-within games. These games can be simple like Tic-Tac-Toe or Breakthrough and extremely
-complicated like Chess or Go.
+In this thesis, we take a look at how we can examine an artificial neural network (ANN) to understand which higher-level concepts (HLC) it deems important for a given state within a game.
+Such a game can be simple like Tic-Tac-Toe or Breakthrough and complicated like Chess or Go.
 
-The method of examining HLC's within games has generally been done by examining the
-current state of the game by evaluating them concerning some heuristic. A heuristic within games are evaluations of the end reward for a state given only the state, for example,
-the number of pieces left within a game of chess. Intuitively, the amount of pieces
-left is a good estimate for a state in chess, this is a higher-level concept we
-use to evaluate a chess position. The piece amount can be considered as a lower-level concept
-than other concepts we use. For instance, grandmaster chess players evaluate a position w.r.t. states
-where the king is safe from attack or the structure of the pawn positions.
-This paper examines the evaluation of a neural network of a state regarding those higher-level concepts if human players evaluate the king's safety of a state
-low but the neural network highly values it, and the neural network plays better
-than the player. There could be an avenue for us to improve our game by considering king safety more thoroughly.
-
-We define the research questions:
+The method of examining HLC's within games has generally been done by examining the current state of the game by evaluating them using some heuristic.
+A heuristic within games are evaluations of the expected end reward for a state, for example, by using the number and type of pieces left within a game of chess.
+Intuitively, the material advantage is a higher-level concept we can use to evaluate a chess position.
+Of course there are many other relevant concepts to consider, for example whether the king is safe from attack, and the pawn structure.
+This thesis examines the evaluation of a neural network of a given state with respect to those higher-level concepts.
+More precisely, we define the following research questions:
 
 \begin{enumerate}
-      \item Given a NN that plays Breakthrough, can we evaluate if that NN recognizes the
-            HLCs that human players use.
-      \item Does a NN that trains itself using self-play learn some HLCs over time, and does it start to emphasize
-            the HLCs that are generally considered better, more as it trains more, and oppositely does it start to
+	\item \label{rq:1}Given a NN that plays Breakthrough, can we determine if that NN has learned the HLCs that human players use.
+	\item \label{rq:2}Does a NN that trains itself using self-play learn some HLCs over time, and does it start to emphasize
+	      the HLCs that are generally considered better the more it trains, and conversely, does it start to
 	      demphasize HLCs that are considered worse.
@@
 
-\subsection{Summary of the thesis}
+\subsection*{Summary}
 
 In this thesis, we take an example game of Breakthrough, train a neural network to play the game using only self-play.
-That is, the neural network is trained only by playing against itself with no outside input other than the rules of the game itself.
-And we examine the neural networks against popular Higher-level concepts (heuristics) that we use to play the game.
+That is, the neural network is trained only by playing against itself with no external input other than the rules of the game itself.
+And we examine the neural networks against popular higher-level concepts (heuristics) that we use to play the game.

### 2020-12-16 latex/sections/managerial.tex
yngvi+james comments fixed

@@
 
-\ifdraft{\coverchapter{Important!!!  Read the Instructions!!!} If you
-  have not already done so, \LaTeX{} the \path{instructions.tex} to
-  learn how to setup your document and use some of the features.  You
-  can see a (somewhat recent) rendered PDF of the instructions included in this folder at \path{instructions-publish.pdf}.
-  There is also more information on working with \LaTeX{} at
-  \url{http://samvinna.ru.is/project/htgaru/how-to-get-around-projects-publish.pdf}.
-  This includes common problems and fixes.
-
-  This page will disappear in anything other than draft mode.}{}
-
 %% Dedication is optional, comment out if it is absent
 \begin{dedications}
-  I dedicate this thesis to my family who while not understanding most of what I do
-  always support me. The friends that still want to talk to me even though my
-  schedule has rendered me unmeetupwithable. Lastly but certainly not least,
-
-  Don't Panic.
+  I dedicate this thesis to my family, who while not understanding most of what I do,
+  always support me and the friends that still want to talk to me even though my
+  schedule has rendered me unmeetupwithable. Lastly, my girlfriend Hulda, who always helps me get 
+  through the tough times, and makes the good times even better.
 \end{dedications}
@@
 \begin{quotation}
-So long, and thanks for all the fish.
+Throughout the writing of my thesis I always had support and assistance from the people around me.
+
+I would first like to thank my supervisor, Yngvi Björnsson, his knowledge in this field is vast and he is always ready to share invaluable wisdom when it's needed.
+Your critical feedback and consistent positivity has pushed me and my work to a higher level.
+
+Secondly, I would like to thank my colleagues, who provided stimulating conversations on the topic and a helpful eye from a different perspective every time I needed it.
+
 \end{quotation}
 \vspace{\baselineskip}
-
-\draftnote{Acknowledgements are optional; comment this chapter out if they are absent
-  Note that it is important to acknowledge any funding that helped in the work}
-
-This work was funded by \the\year~RANNIS grant ``Survey of man-eating Minke whales'' 1415550.
-Additional equipment was generously donated by the Icelandic Tourism Board.
-
-\coverchapter{Preface}
-
-This dissertation is original work by the author, Sigurður Helgason.
-
-\draftnote{The preface is an optional element
-  explaining a little who performed what work.  See
-  \url{https://www.grad.ubc.ca/sites/default/files/materials/thesis_sample_prefaces.pdf}
-  for suggestions.
-
-  List of publications as part of the preface is
-  optional unless elements of the work have already been published.
-  It should be a comprehensive list of all publications in which
-  material in the thesis has appeared, preferably with references to
-  sections as appropriate.  This is also a good place to state
-  contribution of student and contribution of others to the work
-  represented in the thesis.}
 
@@
 \end{tabular}
+\overfullrule=0pt

### 2020-12-16 latex/sections/methods.tex
latex updates
yngvi+james comments fixed

@@
-\chapter{Methods\label{cha:methods}}
+\chapter{Methods}
+
+\label{cha:methods}
 
@@
 The training process uses MCTS, as described in Section \ref{sec:mcts}, to guide its training.
-We outline the architecture used in the neural network.
+We outline the architecture we used for our neural network.
 
@@
 For a neural network to be able to use input, that input needs to be numeric.
-We model the board with a single $NxMx3$ array with two values $\{0,1\}$, where the first two indices on the $z$ axis represent the players, and the last layer on the $z$ axis represents which players turn it is.
-The first $Z$ index represents whites pawns, that is, each $x,y$ position on the board where white has a pawn that cell $x,y,0$ has the value $1$ otherwise it has $0$, and similarily on the second layer for the black player.
-The last $z$ layer, contains all $1$'s if it is the white players turn otherwise $0$.
+We model the board with a single $N\times M\times 3$ array with two values $\{0,1\}$, where the first two indices on the $z$ axis represent the players, and the last layer on the $z$ axis represents which players turn it is.
+The first $z$ index represents whites pawns, that is, each $(x,y)$ position on the board where white has a pawn has the value $1$ otherwise it has $0$, and similarily on the second layer for the black player.
+The last $z$ layer, contains all $1$'s if it is the white players turn otherwise all $0$'s.
 
-One of the outputs of the nerual network is the policy vector, the policy vector is a $3$-dimensional matrix where the lengths of the dimensions are $N * M * 6$ and each index of the matrix represents an action moving from cell $x,y$ moving to the direction $z$.
-The directions are as follows $0$ represents moving upwards and to the left diagonally on the board, $1$ upwards, and $2$ upwards and to the right diagonally, $4$ downwards and to the left diagonally, $5$ downwards, and lastly $6$ downwards and to the right diagonally.
-The cell values indicate the probability distribution of the neural network selecting that move.
+One of the outputs of the neural network is the policy vector.
+The policy vector is a $3$-dimensional matrix where the lengths of the dimensions are $N * M * 6$ and each index of the matrix represents an action moving from cell $x,y$ moving to the direction $z$.
+The directions are as follows: $0$ represents moving upwards and to the left diagonally on the board, $1$ upwards, $2$ upwards and to the right diagonally, $4$ downwards and to the left diagonally, $5$ downwards, and lastly $6$ downwards and to the right diagonally.
+The cell values indicate the probability distribution of the search selecting that move.
 Importantly, before we select a move from the neural networks prediction we zero out the illegal moves, and renormalize the array such that it sums to $1$.
@@
 
-The neural network architecture we opted to use was a single convolutional layer with a ReLU activation function, followed by 5 residual layers each containing two layers of convolution, batch normalization, and a ReLu activation.
+The neural network architecture we opted to use was a single convolutional layer with a ReLU\cite{hinton:relu} activation function, followed by 5 residual layers each containing two layers of convolution, batch normalization, and a ReLu activation.
 Lastly, a split policy/value head, where the output of the last residual layer is split into two different outputs.
@@
 The value head consists of three layers: first a convolutional layer, then a fully connected layer, and lastly a fully connected layer with a \textit{tanh} activation function.
-Figure~\ref{fig:nnarch}.
-depicts the architecture.
-Using this architecture we end up with $1737986$ trainable parameters.
+Figure~\ref{fig:nnarch} depicts the architecture.
+Using this architecture we end up with $1{,}737{,}986$ trainable parameters.
 
@@
 These residual layers stack convolutions on each other making the final convolution represent the locality of all the other localities, allowing the neural network to have a representation of the whole game state in its parameters.
-
-The selection of parameters was done for these reasons as well as to most closely resemble the architecture described in AlphaZero.
+The selection of parameters was done for these reasons as well as to most closely resemble the architecture described in AlphaZero\cite{Silver:alphazero}.
 
@@
 
-The residual layers as a concept remember the output of the previous layer and add together the results of the residual layer and the previous layer.
+The residual layers add the output of the previous layer and the results of the residual layer.
 On a higher level, this leads to the internal representation of the neural network to maintain the whole game boards in its representation, s.t.
@@
 
-Why we use a residual layer for applying a convolution layer multiple times is to gain locality of the whole game board.
-A residual layer applies a function like this $res(x) = x + l(x)$ where $l(x)$ is the function of the layer, commonly a convolutional layer.
+The reason for why we use a residual layer for applying a convolution layer multiple times is to gain locality of the whole game board.
+A residual layer applies a function like this $res(x) = x + l(x)$ where $x$ is the input and $l(x)$ is the function of the layer, commonly a convolutional layer.
 
@@
 
-The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board $h$ is the height of the board and $6$ represents the six cardinal direction pawns can move (straight, and diagonally both left and right for the first player, and backward for the second player).
+The policy head of the neural network returns a vector of the size of the action space $w * h * 6$ where $w$ is the width of the board, $h$ is the height of the board, and $6$ represents the six cardinal direction pawns can move (straight, and diagonally both left and right for the first player, and backward for the second player).
 The vector is then masked s.t.
@@
 
-\subsection{Implementation}
+\section{Implementation}
 
@@
         \FOR{1 \TO 100}
-        \WHILE{\textbf{not} game.is\_termina()}
+        \WHILE{\textbf{not} game.is\_terminal()}
         \STATE{nn1.make\_move()}
@@
 
-\section{Explainable state representations}
+\section{Explainable State Representations}
 
-In this section we describe the process of trainin a Concept Activation Vector in order to linearly separate each state into points in a space with the concept and ones without the concept.
+In this section we describe the process of training a Concept Activation Vector in order to linearly separate each state into points in a space with the concept and ones without the concept.
 
-\subsection{Testing With Concept Activation Vectors}
+\subsection{Testing with Concept Activation Vectors}
 
@@
 The higher-level idea to human players would be that they are in a better position since they have more pieces.
+
+\begin{algorithm}[t]
+    \caption{CAV creation pseudo-code}
+    \label{alg:CAVcreation}
+    \begin{algorithmic}[1]
+        \STATE{\textbf{Input:} neural network nn, column\_name, breakpoint, data\_set, sample\_size}
+        \STATE{\textbf{Output:} A concept activation vector for the column with \textit{column\_name}}
+        \STATE{positive\_set = \textbf{select\_positive\_samples}(data\_set, column\_name, sample\_size)}
+        \STATE{negative\_set = \textbf{select\_negative\_samples}(data\_set, column\_name, sample\_size)}
+        \STATE{labelled\_set = positive\_set + negative\_set}
+        \STATE{unlabelled\_set = dataset - unlabelled\_set}
+        \STATE{model = \textbf{StochasticGradientDescentClassifier}()}
+        \FOR{state \textbf{in} labelled\_set}
+        \STATE{//Modify labelled dataset to contain internal representations}
+        \STATE{state = nn.\textbf{get\_internal\_representation}(state)}
+        \ENDFOR
+        \STATE{train\_set, test\_set = \textbf{train\_test\_split}(labelled\_dataset)}
+        \STATE{model.\textbf{train}(train\_set)}
+        \STATE{prediction\_score, roc\_auc\_score = model.\textbf{predict}(test\_set)}
+        \STATE{\textbf{return} model}
+    \end{algorithmic}
+\end{algorithm}
 
@@
 We train a Stochastic Gradient Descent classifier, from the SKLearn library, to construct a hyper-plane that splits the space into two binary states, one that contains the HLC and another that doesn't.
+We show pseudo-code of this process in Algorithm \ref{alg:CAVcreation}.
 
@@
 Once the prediction has finished, we view the selected action $a$ of $s$ from the policy vector $p$ from the neural network.
-We view the loss of that selected action $a$.
 We then apply the backpropagation algorithm up to that same layer $l_{split}$ and evaluate the gradient there.
@@
 
+\begin{algorithm}[t]
+    \caption{Evaluate a neural network w.r.t concepts pseudo-code}
+    \label{alg:CAVapplication}
+    \begin{algorithmic}[1]
+        \STATE{\textbf{Input:} \textbf{neural network} nn, \textbf{concept activation vector} cav}
+        \STATE{\textbf{Output:} Portion of encountered states that contain the concept of the concept activation vector}
+        \STATE{state\_count = 0}
+        \STATE{has\_tcav = 0}
+        \FOR{1 \TO 100}
+        \STATE{current\_state = \textbf{Breakthrough}.initial\_state}
+        \WHILE{! current\_state.\textbf{terminal}()}
+        \STATE{current\_internal\_representation = \\ nn.\textbf{get\_internal\_representation}(current\_state)}
+        \STATE{output = nn.\textbf{finish\_propagating}(current\_internal\_representation)}
+        \STATE{gradient\_direction = \\ nn.\textbf{get\_gradient}(current\_internal\_representation, output)}
+        \STATE{tcav\_score = cav $ \cdot $ gradient\_direction}
+        \IF{\textbf{same\_direction}(tcav\_score, cav)}
+            \STATE{has\_tcav++}
+        \ENDIF
+        \STATE{state\_count++}
+        \STATE{current\_state = nn.\textbf{get\_next\_state}(current\_state)}
+        \ENDWHILE
+        \ENDFOR
+        \STATE{\textbf{return} has\_tcav / state\_count}
+    \end{algorithmic}
+\end{algorithm}
+
+In Algorithm \ref{alg:CAVapplication} we show pseudo-code depicting how we evaluate if a neural network recognizes a given concept.
+The resulting value from running the algorithm is, the proportion of states that the neural network encountered during playing against it self that contain the HLC.
+
 \section{Summary}
 
-In this chapter we described the architecture of the nerual network, including how we model Breakthrough to be able to learn how to play the game.
-We described the processes we use to learn completely by self play.
+In this chapter we described the architecture of the neural network, including how we model Breakthrough to be able to learn to play the game, we described the processes we use to learn from self play.
 In the following chapter we will evaluate the neural network.

### 2020-12-16 latex/sections/results.tex
latexing
yngvi+james comments fixed

@@
         \end{center}
-        \caption{Winrate vs random agent}
+        \caption{Winrate vs. random agent}
         \label{fig:winratevsrandom}
@@
         \end{center}
-        \caption{Winrate vs MCTS agent}
+        \caption{Winrate vs. MCTS agent}
         \label{fig:winratevsmcts}
@@
         \end{center}
-        \caption{Winrate vs neural network trained for 150 generations}
+        \caption{Winrate vs. neural network trained for 150 generations}
         \label{fig:winratevsneuralnetwork}
@@
 Examining the neural network is only interesting if the neural network can effectively play the game.
-We first examine the history neural network playing against an agent that only takes random moves, on every $10$ generations.
-The win rate over generations can be seen in Figure \ref{fig:winratevsrandom}.
-It is clear that very early on in the training process, almost immediately after generation $10$ our agent performs nearly perfectly against the random agent.
+We first examine the history of the neural network playing against an agent that only takes random moves.
+The win rate over every $10$ generations can be seen in Figure \ref{fig:winratevsrandom}.
+It is clear that very early on in the training process, almost immediately after generation $10$, our agent performs nearly perfectly against the random agent.
 This implies that the agent does indeed learn some strategy in the game.
-The next agent we tried the neural network against, was an agent that does regular Monte-Carlo tree search on the game space with $100$ iterations of rollout for each move.
+The next agent we tried the neural network against was an agent that does regular Monte-Carlo tree search on the game space with $100$ iterations of random UCT rollout for each move.
+The MCTS agent has a $c_{uct}$ value of $0.9$.
 A graph showing the win rate over generations is depicted in Figure \ref{fig:winratevsmcts}.
-Again, our agent quickly learns some strategy and can win often, although doesn't achieve a perfect win rate after $300$ generations.
+Again, our agent quickly learns some strategy and can win often, although not achieving a perfect win rate even after $300$ generations.
 The last agent we tested against was the neural network itself, although only trained to $150$ generations.
 The graph in Figure \ref{fig:winratevsneuralnetwork} shows win rate.
-As expected the agent performs poorly until it has trained for more than $150$ generations.
+As one would expect, the agent that has trained for less time performs poorly until it has trained for more than $150$ generations.
 
-These results imply that our agent certainly understands how to play the game of breakthrough to a certain level.
-However, an intermediate level human player would most often be able to win the agent, based on the author's experience playing against it.
-Note, that the main focus of this was not to create an expert level agent.
-Rather to, gauge into the concepts the agent learns as described in the next section.
+These results imply that our agent certainly understands how to play the game of Breakthrough to a certain level.
+However, an intermediate level human player would most often be able to beat the agent, based on the author's experience playing against it.
+Note that the main focus of this was not to create an expert level agent.
+Rather, to gauge into the concepts the agent learns as described in the next section.
 
-\section{Evaluating the improvements of the Neural Network over generations}
+\section{Evaluating the Improvements of the Neural Network over Generations}
 
@@
 
-We test four concepts \textit{Material Advantage}, \textit{Aggressiveness}, \textit{Unity}, and \textit{Lorentz-Horey}.
+We test four concepts \textit{Material Advantage}, \textit{Aggressiveness}, \textit{Unity}, and \textit{Lorentz-Horey score}.
 
-\subsection{Description of concepts}
+\subsection{Description of Concepts}
 
@@
 
-The first concept we examine is Material Advantage which conveys the idea of having more pawns than your opponent.
-When faced with a board it is difficult to argue for whether having a single pawn up on your opponent constitues as Material Advantage or three.
-In order to an appropriate value for this case we used the neural network that had been trained for $300$ generations to generate $20000$ unique states.
+The first concept we examine is Material Advantage which conveys the idea of having more pawns than the opponent.
+When faced with a board it is difficult to argue for whether having a single pawn or three up on your opponent constitues as significant Material Advantage.
+In order to select an appropriate value for this case we used the neural network that had been trained for $300$ generations to generate $20{,}000$ unique states.
 The distribution of these states' difference of pawns can be seen in Figure \ref{fig:distnumberadv}.
 From examining the distribution and examining samples we selected the breakpoint of $\geq 2$, meaning that if you have $2$ pieces on your opponent you are in a state that has this concept.
-From the $20000$ states only $5.5\%$ of states have this concept.
+From the $20{,}000$ states only $5.5\%$ of states have this concept.
 
@@
 
-The next concept is Aggressiveness, which describes how much closer your furthest piece is to the opponents edge than your opponents furthest piece to your edge.
-This is then the minimum amount of how many moves it will take you to win the game.
+The next concept is Aggressiveness, which describes how much closer your furthest piece is to the opponent's edge than your opponent's furthest piece to your edge.
+This is the minimum amount of how many moves it will take you to possibly win the game.
 A distribution of these values can be seen in Figure \ref{fig:distaggressiveness}.
 We examined the distribution, and sampled states from various breaking points and decided on the value of $\geq 1$.
-That is if the value of a state's furthest piece difference is greater than or equal to $1$ that state has the concept of Aggressiveness.
+That is, if the value of a state's furthest piece difference is greater than or equal to $1$ that state has the concept of Aggressiveness.
 From the dataset $18.3\%$ of states have this concept.
@@
     \centering
+    \caption{Lorentz-Horey cell values, from white players point of view}
     \begin{tabular}{|c|c|c|c|c|c|}
         \hline
+        21 & 21 & 21 & 21 & 21 & 21 \\\hline
+        11 & 15 & 15 & 15 & 15 & 11 \\\hline
+        7  & 10 & 10 & 10 & 10 & 7  \\\hline
+        4  & 6  & 6  & 6  & 6  & 4  \\\hline
+        2  & 3  & 3  & 3  & 3  & 2  \\\hline
         5  & 15 & 5  & 5  & 15 & 5  \\\hline
-        2  & 3  & 3  & 3  & 3  & 2  \\\hline
-        4  & 6  & 6  & 6  & 6  & 4  \\\hline
-        7  & 10 & 10 & 10 & 10 & 7  \\\hline
-        11 & 15 & 15 & 15 & 15 & 11 \\\hline
-        21 & 21 & 21 & 21 & 21 & 21 \\\hline
     \end{tabular}
-    \caption{Lorentz-Horey cell values, from white players point of view}
     \label{table:lorentzcell}
@@
 
-The last concept we examined was the Lorentz-Horey.
+\label{section:lorentzhoreyref}
+
+The last concept we examined was the Lorentz-Horey score.
 This concept is a popular heuristic in Breakthrough defined in the paper by Lorentz \& Horey \cite{lorentz:heuristic}.
@@
 The matrix shown is flipped from the black player's point of view.
-The values are selected in such a way that as your pawn approach the opponent's side their value increases, having pieces on the edge isn't optimal, as such a pawn will only be able to capture in one direction and can not escape capture as easily.
+The values are selected in such a way that as your pawns approach the opponent's side their value increases, having pieces on the edge isn't optimal, as such a pawn will only be able to capture in one direction and can not escape capture as easily.
 To select the breakpoint for Lorentz-Horey heuristic, we generate a distribution of the difference in Lorentz-Horey Score over our dataset.
@@
 
-\begin{figure}[]
+\begin{figure}[h]
     \centering
@@
 
-From Figure \ref{fig:aggressiveness}.
-we can see that the aggressiveness HLC is a growing factor over as the neural network is trained.
+From Figure \ref{fig:aggressiveness}, we can see that the aggressiveness HLC is a growing factor over as the neural network is trained.
 As a piece can always move forward, if your opponent does not capture that piece, this concept relates to how many moves it will take for you to win.
@@
 
-\begin{figure}
+Examining Figure \ref{fig:unity}, we see that the Unity HLC is steady, but then it drops as the network is trained.
+As this is considered a decent strategy in Breakthrough this is unexpected.
+However, this concept become increasingly unreliable as you lose pieces because it is the average of your remaining pieces distance to their center.
+
+\subsection{Lorentz-Horey Score}
+
+\begin{figure}[h]
     \begin{small}
@@
 
-The literature of Breakthrough implies that being patient and waiting until your opponent makes a mistake is generally the preferred strategy for winning.\cite{lorentz:heuristic} Examining Figure \ref{fig:lorentzheuristic}, we can see that the trend of using this HLC increases over time.
+Examining Figure \ref{fig:lorentzheuristic}, we can see a trend where as the neural network is trained its use of this HLC increases over time.
 As this heuristic is the one most often cited as a successful strategy in Breakthrough, it is interesting to see that our Breakthrough agent is gradually learning a similar concept.

### 2020-12-16 hlc_evaluator/gen_data.py

@@ -1,0 +1,70 @@
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from game_environments.breakthrough.breakthrough import BTBoard, config as BTconfig
+import random
+from tqdm import tqdm
+from PIL import Image
+import ast
+from itertools import permutations
+from monte_carlo.mcts import MCTS
+from monte_carlo.mctsnode import Node
+
+def from_np_array(array_string):
+    array_string = ','.join(array_string.replace('[ ', '[').split())
+    return np.array(ast.literal_eval(array_string))
+
+GAME = {
+    "breakthrough": BTBoard(np.zeros([6,6]), 1),
+}
+selected_game = GAME["breakthrough"]
+initial_state = selected_game.initial_state()
+seen_states = {}
+
+frontier = [initial_state]
+
+generate_new_data = True
+new_data_size = 1000000
+dfs = True
+
+if generate_new_data:
+    df = pd.DataFrame(columns=["id",
+                               "player",
+                               "state",
+                               "terminal", 
+                               "player_piece_amount", 
+                               "piece_difference", 
+                               "furthest_piece", 
+                               "furthest_piece_difference"], 
+                      dtype=np.int64
+                      )
+    state_id = 0
+
+    while frontier and state_id < new_data_size:
+        working_state = frontier.pop() if dfs else frontier.pop(0)
+        if working_state.__hash__() in seen_states:
+            continue
+
+
+        heuristics = working_state.get_heuristics()
+        df = df.append({
+            'id':state_id,
+            'state':working_state.encode_state(),
+            'player':min(working_state.player+1,1),
+            'terminal':working_state.is_terminal(),
+            'player_piece_amount':heuristics[0][1],
+            'piece_difference':heuristics[1][1],
+            'furthest_piece':heuristics[2][1],
+            'furthest_piece_difference':heuristics[3][1]
+        },ignore_index=True)
+        
+        state_id += 1
+        seen_states[working_state.__hash__()] = True
+        np.random.shuffle(working_state.legal_moves)
+        for lm in working_state.legal_moves:
+            frontier.append(working_state.execute_move(lm))
+    df.set_index('id') 
+
+    df.to_csv(f'./{"dfs" if dfs else "bfs"}_moves_heuristic_{new_data_size}.csv')
+else:
+    df = pd.read_csv(f'./{"dfs" if dfs else "bfs"}_moves_heuristic_{new_data_size}.csv', converters={'state': from_np_array}).set_index('id')

### 2020-12-16 hlc_evaluator/trainer.py

@@ -37,7 +37,7 @@ TEMP_THRESHOLD = 10000
 TRAINING_ITERS = 10
 VERIFICATION_GAMES = 20
 CUDA = torch.cuda.is_available()
-ITERATION = 69
+ITERATION = 369
 
 
 def generate_dataset(primary_nn: BreakthroughNN, game_example : GameNode, saved_monte_tree=None, verbose=False):
@@ -98,7 +98,7 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
 
   initial_node = Node(state_example.initial_state(), "START")
 
-  generation =  70
+  generation =  370
 
 
   while True:
@@ -123,7 +123,7 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
           # """
           pi,v = neural_network_1.safe_predict(curr_node.gamestate)
           if CUDA:
-            pi = pi.detach().cpu().numpy() 
+            pi = pi.detach().cpu().numpy()
             pi = pi.reshape(-1)
           else:
             pi = pi.view(-1)
@@ -166,7 +166,7 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
           # """
           pi,v = neural_network_2.safe_predict(curr_node.gamestate)
           if CUDA:
-            pi = pi.detach().cpu().numpy() 
+            pi = pi.detach().cpu().numpy()
             pi = pi.reshape(-1)
           else:
             pi = pi.view(-1)
@@ -215,7 +215,7 @@ def selfplay(first_network_path, first_network_name, second_network_path, second
         neural_network_1.savemodel(first_network_path,first_network_name)
         neural_network_2.savemodel(second_network_path,second_network_name)
     if generation % 10 == 0:
-        neural_network_1.savemodel("./trained_models", f"session_res10_gen{generation}.tar")
+        neural_network_1.savemodel("./trained_models", f"session2_res5_gen{generation}.tar")
 
 
     print("[trainer.py] STARTING TRAINING")
